### 2019-10-08 Progress Documentation.Rmd
added progress report

---
title: "Progress Documentation"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{itemize}
  \item 10 / 08 / 2019: Nimble crashes my Mac laptop every now and then, so I should do more simulation with that.
\end{itemize}

### 2019-10-09 Progress Documentation.Rmd

@@
   \item 10 / 08 / 2019: Nimble crashes my Mac laptop every now and then, so I should do more simulation with that.
+  \item 10 / 09 / 2019: Meeting with Dr.
+Kahle.
+I want to look into getting the WINE path set globally instead of specifying it each time.
+I also have been dealing with the problem of getting STAN compile time included or not included in the benchmark file.
+Currently I'm having two function calls to "mark", one that uses saved rds file and has STAN already compiled, and one that renames the rds file to force recompile.
+I could instead try to split STAN up into the compile function and the run function.
+  \item 10 / 09 / 2019: I am interested not just in time it takes for each model to run, but also the accuracy of each model.
+For models that can be calculated analytically, I can look at the MSE of each method in their posterior means compared to the true posterior mean.
+For those where I can't calculate analytically, I can run STAN for a super long time and get the posterior mean, and maybe check it with a Gibbs sampler as well, and that will be my "true" posterior mean when calculating the MSEs for each method in that model setup.
 \end{itemize}

### 2019-11-13 Progress Documentation.Rmd

@@
 
-\begin{itemize}
-  \item 10 / 08 / 2019: Nimble crashes my Mac laptop every now and then, so I should do more simulation with that.
-  \item 10 / 09 / 2019: Meeting with Dr.
+- 10 / 08 / 2019: 
+  - Nimble crashes my Mac laptop every now and then, so I should do more simulation with that.
+- 10 / 09 / 2019: 
+  - Meeting with Dr.
 Kahle.
@@
 I could instead try to split STAN up into the compile function and the run function.
-  \item 10 / 09 / 2019: I am interested not just in time it takes for each model to run, but also the accuracy of each model.
+- 10 / 09 / 2019: 
+  - I am interested not just in time it takes for each model to run, but also the accuracy of each model.
 For models that can be calculated analytically, I can look at the MSE of each method in their posterior means compared to the true posterior mean.
 For those where I can't calculate analytically, I can run STAN for a super long time and get the posterior mean, and maybe check it with a Gibbs sampler as well, and that will be my "true" posterior mean when calculating the MSEs for each method in that model setup.
-\end{itemize}
+- 10 / 10 / 2019: 
+  - How does each method do with specified number of iterations? 100? 1000? How close do they get to the truth?
+- 10 / 30 / 2019: 
+  - I have successfully made the benchmarking process more clean by having a "benchmark-function.R" file that includes all the code to benchmark the methods against each other, with a function argument to re-compile STAN or not, which can be sourced in each benchmark file within each folder.
+  - Nimble cannot save its compilation. It has to recompile each time, unlike STAN.
+  - We also want to clean up the benchmark files by sourcing data setup code from each script file instead of copying and pasting it every time into the benchmark file. 
+  - I am also going to start working on incorporating GRETA, which is R-based and uses Tensorflow to run HMC. 
+  - We want to test how methods for models we already have will respond to data given that is weird or highly correlated (or like data on a circle from Jerry's PPP).

### 2020-01-08 Paper.Rmd
code formatting update

---
title: "Paper"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Outline
- General Bayesian computational settings
  - Chains
  - Iterations
  - Warmup
  - Thinning
  - initial value specification
- Each sampler
  - BUGS
    - WINE
    - Debugging
  - Jags
  - Nimble
    - Crashes
  - STAN
    - RDS file
  - Greta
- Common coding techniques for each language
  - Truncated normal
  - Specifiying data and constants
  - Matrix multiplication
  - Initial value specification?
- Benchmarking
  - Run time
  - Memory allocation
  - Computational accuracy
  - Benchmarking files
- Folder design
  - General model type
    - Simple conjugate models
    - Regression models
    - Time Series models
    - Survival models
    - 8 Schools
  - Specific model type
  - Scripts for each method 
  - Benchmarking
- Here package
- Models Used

\newpage

## BUGS

OpenBugs is the oldest of the Bayesian samplers that we use. OpenBugs is designed for Windows machines, but there is a workaround to run it on MacOS and Linux, using WINE, a program designed to run Windows software on other operating systems. OpenBugs can be used through the actual click-and-point application, or through \texttt{R}, using the \texttt{R2OpenBugs} package, which allows for the user to send model scripts and data to OpenBugs without needing to leave the \texttt{R} interface. The \texttt{R} command to run OpenBugs is \texttt{bugs()}. Using this method causes OpenBugs to run invisibly in the background and return the results, unless an argument is specified to debug code in the actual application itself, which can be very useful. When using R to run OpenBugs, there is little way to know what went wrong if the code errors somewhere without interacting with the OpenBugs program itself. Using the \texttt{debug = T} argument within \texttt{bugs()} will cause the application to run in the forefront as if it were actually being used directly by the user. This way, any messages given by the application while running can be viewed directly. It is recommend to use \texttt{debug = T} only when debugging, as it does slow down the sampling process slightly. OpenBugs has many tools to visualize the performance the sampler which are easy to use. The biggest drawback of using the OpenBugs application directly is that specifiying the data can be tedious if the data is not very small in size or design. Using R to specify the data is much easier.

## JAGS

In terms of model specification, JAGS is very similar to OpenBugs, but works more seemlessly in R using the package \texttt{rjags}. The \texttt{runjags} is an optional package that allows for the JAGS model script to be inputted as a string, and run using the \texttt{run.jags()} function. Since JAGS runs inside \texttt{R} itself, it is very easy to work with. Helpful errors are given when there is an issue with the model code or sampling process. JAGS is often a very fast sampler, especially for less complicated models.

## Nimble

Nimble is an extension of Bugs that allows for bayesian computation within \texttt{R}, but uses C++ under the hood to compile models and algorithms for speed. Using the \texttt{nimble} package, Nimble can perform MCMC bayesian computations, but also allows for other types of algorithms to be loaded or custom-built. It can also compile other code without BUGS models into C++ for other uses. For the purpose of bayesian sampling using MCMC, Nimble is used very similarly to BUGS in practicality. The flexibility of Nimble is not needed in most cases when doing MCMC. The model specification language is almost identical to OpenBugs. A model can be both compiled and run using the function \texttt{nimbleMCMC()}. Because the model is compiled using C++ each time, this can drastically increase the total computation time as compared to other sampling methods. Occasionally, Nimble can also freeze up or fail, which can take a bit of time to fix.

## STAN

STAN is a language in-and-of itself used for statistcal modeling and data analysis. STAN can be used in several programming languages, including \texttt{R} and Python. While many bayesian samplers use Gibbs sampling, STAN uses Hamiltonian Monte Carlo under the hood, a more computationally efficient sampler that is based on an advanced physics simulation. This can greatly reduce the computational time for many types of statistical models, especially for more complicated ones. Like Nimble, STAN compiles each model, but one big advantage is that it can store each compiled model in a .Rds file, which allows for repeated use without the need to re-compile each time. In \texttt{R}, STAN is used through the \texttt{rstan} package, which allows for data to be loaded through normal \texttt{R} data structures. The biggest practical difference in using STAN compared to other bayesian computational methods is in the model specification process. STAN requires a model object in the form of a .stan file, which is created in a very unique language that is more formal than in BUGS. A .stan file, either created through a character string in R or in a seperate text file, can be used by the \texttt{stan()} function, which takes the file along with the data to perform the sampling. There is a very robust set of manuals for STAN users, compared to other methods.

## Greta

Greta, a very recent addition to the list of bayes computational methods, is made specifically for use in \texttt{R}, making the model and data specification process as easy as possible. It also uses Google TensorFlow to perform its calculations, and allows for more flexible use of CPUs and GPUs. Unfortunately, as of the writing of this, it is still in need of lots of development, and lacks an appropriate amount of documentation and support for users. Computationally, it is very slow compared to the other bayes methods, especially in simple cases. There are several types of semi-common models that Greta seems unable to properly handle in model specfication at the moment. Greta also attempts to make model specification very user-friendly, but uses some very common function names in the process that might override other often-used \texttt{R} functions.

### 2020-01-08 regression-models/hierarchical-logistic-regression/hierarchical-logistic-regression-bugs.R

@@ -12,31 +12,52 @@ library("here")
 ## generate/specify data
 ################################################################################
 
-theta <- 5 # poisson theta
+n <- 30L # sample size
+d <- 2L
+l <- 2L
+ll <- sapply(1:l, function(x) rep(x,n / l)) %>% as.numeric()
 
 set.seed(1)
-
-(y <- rpois(1, theta))
+(x <- matrix(rnorm(n * d), nrow = n, byrow = TRUE))
+(y <- rbinom(n, 1, 0.5))
 
 bugs_data <- list(
-  "y" = y
+  "N" = n,
+  "D" = d,
+  "L" = l,
+  "ll" = ll,
+  "y" = y, 
+  "x" = x
 )
+
 
 
 ## specify bugs model
 ################################################################################
 
 bugs_model <- function() {
-  y ~ dpois(theta)
-  theta ~ dgamma(3,1)
+  for (d in 1:D) {
+    mu[d] ~ dnorm(0, 0.0001)
+    tau[d] ~ dnorm(0,0.0001) %_% I(0,)
+    for (l in 1:L) {
+      beta[l,d] ~ dnorm(mu[d], tau[d])
+    }
+  }
+  for (n in 1:N) {
+    for (d in 1:D) {
+      temp[n,d] <- x[n,d] * beta[ll[n],d]
+    }
+    temp_sum[n] <- temp[n,1] + temp[n,2] 
+    logit(theta[n]) <- temp_sum[n]
+    y[n] ~ dbern(theta[n])
+  }
 }
-
 
 
 bugs.file <- file.path(tempdir(), "model.txt")
 write.model(bugs_model, bugs.file)
 
-bugs_monitor <- "theta"
+bugs_monitor <- c("mu", "beta", "theta", "tau")
 
 
 ## Specify path to WINE if using WINE 
@@ -53,6 +74,8 @@ if (getwd() == "/Users/evanmiyakawa/Git Projects/Bayes-Computational-Methods/Bay
 }
 
 
+
+
 ## configure model settings
 ################################################################################
 
@@ -61,7 +84,7 @@ n_iter <- 1e4L
 n_warmup <- 1e3L
 
 
-### fit model
+## fit model
 ################################################################################
 source(here("currently-benchmarking.R"))
 
@@ -70,7 +93,7 @@ if (!currently_benchmarking()) {
     "model.file" = bugs.file, "data" = bugs_data, "parameters.to.save" = bugs_monitor, 
[…]

### 2020-01-08 regression-models/hierarchical-logistic-regression/hierarchical-logistic-regression-jags.R

@@ -13,17 +13,25 @@ library("here")
 ## generate/specify data
 ################################################################################
 
-lambda <- 1/2 # exponential lambda
-n <- 10       # sample size
+n <- 30L # sample size
+d <- 2L
+l <- 2L
+ll <- sapply(1:l, function(x) rep(x,n / l)) %>% as.numeric()
 
 set.seed(1)
-
-(y <- rexp(n, lambda))
+(x <- matrix(rnorm(n * d), nrow = n, byrow = TRUE))
+(y <- rbinom(n, 1, 0.5))
 
 jags_data <- list(
-  "y" = y,
-  "N" = n
+  "N" = n,
+  "D" = d,
+  "L" = l,
+  "ll" = ll,
+  "y" = y, 
+  "x" = x
 )
+
+
 
 
 ## specify jags model
@@ -31,14 +39,21 @@ jags_data <- list(
 
 jags_model <- "
   model{
-    for (i in 1:N) {
-      y[i] ~ dexp(lambda)
+    for (d in 1:D) {
+      mu[d] ~ dnorm(0, 1 / (100 ^ 2))
+      tau[d] ~ dnorm(0,0.0001)  I(0, )
+      for (l in 1:L) {
+        beta[l,d] ~ dnorm(mu[d], tau[d])
+      }
     }
-    lambda ~ dgamma(1,1)
+    for (n in 1:N) {
+      logit(theta[n]) <- x[n,] %*% beta[ll[n],]
+      y[n] ~ dbern(theta[n])
+    }
   }
 "
 
-jags_monitor <- c("lambda")
+jags_monitor <- c("mu", "beta", "theta", "tau")
 
 
 ## configure model settings
@@ -47,6 +62,7 @@ jags_monitor <- c("lambda")
 n_chains <- 4L
 n_iter <- 1e4L
 n_warmup <- 1e3L
+
 
 
 ## fit model
@@ -73,7 +89,7 @@ if (!currently_benchmarking()) {
   dimnames(jags_fit_object) <- list(
     "iterations" = NULL, 
     "chains" = 1:n_chains, 
-    "parameters" = monitor
+    "parameters" = jags_monitor
   )
   
   

### 2020-01-08 regression-models/linear-regression/linear-regression-stan.R

@@ -80,17 +80,3 @@ stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)
 
 
 
-
-## benchmarking
-###################################################################################
-
-
-bench_results <- mark(
-  stan_fit <- stan(
-    "file" = stan_file, "data" = stan_data, 
-    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
-  ),
-  iterations = 3
-)
-bench_results[1,2:9]
-

### 2020-01-08 regression-models/logistic-regression/logistic-regression-bugs.R

@@ -5,6 +5,7 @@ library("tidyverse"); theme_set(theme_minimal())
 library("parallel"); options(mc.cores = detectCores())
 library("R2OpenBUGS")
 library("bench")
+library("here")
 
 
 
@@ -55,7 +56,7 @@ bugs_monitor <- c("alpha", "beta")
 ## Specify path to WINE if using WINE 
 ###################################################################################
 
-if (getwd() == "/Users/evanmiyakawa/hubiC/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
+if (getwd() == "/Users/evanmiyakawa/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
   WINE="/usr/local/Cellar/wine/4.0.1/bin/wine"
   WINEPATH="/usr/local/Cellar/wine/4.0.1/bin/winepath"
   OpenBUGS.pgm="/Users/evanmiyakawa/OpenBugs/OpenBUGS323/OpenBUGS.exe" #~
@@ -78,7 +79,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
   bugs_fit <- bugs(
     "model.file" = bugs.file, "data" = bugs_data, "parameters.to.save" = bugs_monitor, 
     "inits" = NULL, "n.chains" = n_chains, "n.iter" = n_iter, "n.burnin" = n_warmup,

### 2020-01-08 regression-models/logistic-regression/logistic-regression-jags.R

@@ -6,6 +6,7 @@ library("parallel"); options(mc.cores = detectCores())
 library("rjags"); library("runjags")
 library("bayesplot")
 library("bench")
+library("here")
 
 
 
@@ -59,7 +60,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
   
   
   jags_fit <- run.jags(

### 2020-01-08 regression-models/poisson-regression/poisson-regression-jags.R

@@ -6,6 +6,7 @@ library("parallel"); options(mc.cores = detectCores())
 library("rjags"); library("runjags")
 library("bayesplot")
 library("bench")
+library("here")
 
 
 

### 2020-01-08 simple models/continuous/exponential/gamma-prior/exponential-gamma-bugs.R

@@ -5,6 +5,7 @@ library("tidyverse"); theme_set(theme_minimal())
 library("parallel"); options(mc.cores = detectCores())
 library("R2OpenBUGS")
 library("bench")
+library("here")
 
 
 
@@ -63,7 +64,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
   bugs_fit <- bugs(
     "model.file" = bugs.file, "data" = bugs_data, "parameters.to.save" = bugs_monitor, 
     "inits" = NULL, "n.chains" = n_chains, "n.iter" = n_iter, "n.burnin" = n_warmup,

### 2020-01-08 simple models/continuous/normal/gamma-prior-tau/normal-gamma-tau-bugs.R

@@ -5,6 +5,7 @@ library("tidyverse"); theme_set(theme_minimal())
 library("parallel"); options(mc.cores = detectCores())
 library("R2OpenBUGS")
 library("bench")
+library("here")
 
 
 
@@ -46,7 +47,7 @@ bugs_monitor <- c("tau")
 ## Specify path to WINE if using WINE 
 ###################################################################################
 
-if (getwd() == "/Users/evanmiyakawa/hubiC/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
+if (getwd() == "/Users/evanmiyakawa/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
   WINE="/usr/local/Cellar/wine/4.0.1/bin/wine"
   WINEPATH="/usr/local/Cellar/wine/4.0.1/bin/winepath"
   OpenBUGS.pgm="/Users/evanmiyakawa/OpenBugs/OpenBUGS323/OpenBUGS.exe" #~
@@ -67,7 +68,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
   bugs_fit <- bugs(
     "model.file" = bugs.file, "data" = bugs_data, "parameters.to.save" = bugs_monitor, 
     "inits" = NULL, "n.chains" = n_chains, "n.iter" = n_iter, "n.burnin" = n_warmup,

### 2020-01-08 simple models/continuous/normal/gamma-prior-tau/normal-gamma-tau-jags.R

@@ -4,8 +4,9 @@
 library("tidyverse"); theme_set(theme_minimal())
 library("parallel"); options(mc.cores = detectCores())
 library("rjags"); library("runjags")
+library("bayesplot")
 library("bench")
-
+library("here")
 
 
 ## generate/specify data
@@ -51,7 +52,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
   
   
   jags_fit <- run.jags(

### 2020-01-08 simple models/continuous/normal/normal-prior/normal-normal-bugs.R

@@ -5,6 +5,7 @@ library("tidyverse"); theme_set(theme_minimal())
 library("parallel"); options(mc.cores = detectCores())
 library("R2OpenBUGS")
 library("bench")
+library("here")
 
 
 
@@ -47,7 +48,7 @@ bugs_monitor <- "mu"
 ## Specify path to WINE if using WINE 
 ###################################################################################
 
-if (getwd() == "/Users/evanmiyakawa/hubiC/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
+if (getwd() == "/Users/evanmiyakawa/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
   WINE="/usr/local/Cellar/wine/4.0.1/bin/wine"
   WINEPATH="/usr/local/Cellar/wine/4.0.1/bin/winepath"
   OpenBUGS.pgm="/Users/evanmiyakawa/OpenBugs/OpenBUGS323/OpenBUGS.exe" #~
@@ -68,7 +69,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
   bugs_fit <- bugs(
     "model.file" = bugs.file, "data" = bugs_data, "parameters.to.save" = bugs_monitor, 
     "inits" = NULL, "n.chains" = n_chains, "n.iter" = n_iter, "n.burnin" = n_warmup,

### 2020-01-08 simple models/continuous/normal/normal-prior/normal-normal-jags.R

@@ -4,7 +4,9 @@
 library("tidyverse"); theme_set(theme_minimal())
 library("parallel"); options(mc.cores = detectCores())
 library("rjags"); library("runjags")
+library("bayesplot")
 library("bench")
+library("here")
 
 
 

### 2020-01-08 simple models/discrete/bernoulli/beta-prior/beta-bernoulli-bugs.R

@@ -5,6 +5,7 @@ library("tidyverse"); theme_set(theme_minimal())
 library("parallel"); options(mc.cores = detectCores())
 library("R2OpenBUGS")
 library("bench")
+library("here")
 
 
 
@@ -61,7 +62,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
   bugs_fit <- bugs(
     "model.file" = bugs.file, "data" = bugs_data, "parameters.to.save" = bugs_monitor, 
     "inits" = NULL, "n.chains" = n_chains, "n.iter" = n_iter, "n.burnin" = n_warmup,

### 2020-01-08 simple models/discrete/bernoulli/beta-prior/beta-bernoulli-jags.R

@@ -6,6 +6,7 @@ library("parallel"); options(mc.cores = detectCores())
 library("rjags"); library("runjags")
 library("bayesplot")
 library("bench")
+library("here")
 
 
 
@@ -46,7 +47,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
   
   
   jags_fit <- run.jags(

### 2020-01-08 simple models/discrete/binomial/beta-prior/beta-binomial-bugs.R

@@ -5,6 +5,7 @@ library("tidyverse"); theme_set(theme_minimal())
 library("parallel"); options(mc.cores = detectCores())
 library("R2OpenBUGS")
 library("bench")
+library("here")
 
 
 
@@ -64,7 +65,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
   bugs_fit <- bugs(
     "model.file" = bugs.file, "data" = bugs_data, "parameters.to.save" = bugs_monitor, 
     "inits" = NULL, "n.chains" = n_chains, "n.iter" = n_iter, "n.burnin" = n_warmup,

### 2020-01-08 survival-models/survival.R

@@ -1,0 +1,107 @@
+library(tidyverse)
+library(survival)
+library(survminer)
+data("lung")
+head(lung)
+fit <- survfit(Surv(time, status) ~ sex, data = lung)
+print(fit)
+
+library(ranger)
+library(ggfortify)
+
+#------------
+data(veteran)
+head(veteran)
+
+km <- with(veteran, Surv(time, status))
+km_fit <- survfit(Surv(time, status) ~ 1, data=veteran)
+summary(km_fit, times = c(1,30,60,90*(1:10)))
+autoplot(km_fit)
+
+vet <- mutate(veteran, AG = ifelse((age < 60), "LT60", "OV60"),
+              AG = factor(AG),
+              trt = factor(trt,labels=c("standard","test")),
+              prior = factor(prior,labels=c("N0","Yes")))
+
+km_AG_fit <- survfit(Surv(time, status) ~ AG, data=vet)
+autoplot(km_AG_fit)
+
+cox <- coxph(Surv(time, status) ~ trt + celltype + karno + diagtime + age + prior , data = vet)
+summary(cox)
+
+?Surv
+Surv(vet$time, vet$status)
+
+vet <- filter(vet, status == 1)
+
+cox <- coxph(Surv(time, status) ~ trt, data = vet)
+summary(cox)
+
+
+####
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rjags"); library("runjags")
+library("bayesplot")
+library("bench")
+library("here")
+
+y <- vet$time
+trt <- vet$trt
+
+jags_data <- list(
+  "t" = y,
+  "x" = trt,
+  "N" = length(y)
+)
+
+jags_model <- "
+  model{
+    for (i in 1:N) {
+      lambda[i] <- exp(beta_0 + beta * x[i])
+      t[i] ~ dexp(lambda[i])
+    }
+    beta ~ dnorm(0,0.0001)
+    beta_0 ~ dnorm(0,0.0001)
+  }
+"
+
+# 
+# jags_model <- "
+#   model{
+#     for (i in 1:N) {
+#       y_hat[i] <- exp(beta * x[i])
+#       y[i] ~ dnorm(log(y_hat[i]), tau)
+#     }
+#     beta ~ dnorm(0,0.0001)
+#     tau ~ dnorm(0,0.0001)  I(0, )
+#   }
[…]

### 2020-01-08 time-series/ar-model/ar-model-bugs.R

@@ -5,6 +5,7 @@ library("tidyverse"); theme_set(theme_minimal())
 library("parallel"); options(mc.cores = detectCores())
 library("R2OpenBUGS")
 library("bench")
+library("here")
 
 
 
@@ -79,7 +80,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
   bugs_fit <- bugs(
     "model.file" = bugs.file, "data" = bugs_data, "parameters.to.save" = bugs_monitor, 
     "inits" = NULL, "n.chains" = n_chains, "n.iter" = n_iter, "n.burnin" = n_warmup,

### 2020-01-15 Paper.Rmd
updates

@@
   - Chains
+    - Greta
   - Iterations
@@
   - Initial value specification?
+  - sigma / tau
 - Benchmarking
@@
 Using R to specify the data is much easier.
+One unique feature in OpenBugs is the Doodle tool, which allows for model specification through the creation of a diagram of nodes representing the parameters.
+Though using Doodle can be helpful for illustration, it is not used often in practice.
 
@@
 Greta also attempts to make model specification very user-friendly, but uses some very common function names in the process that might override other often-used \texttt{R} functions.
+Greta does offer a model visualization tool that does not exist in any of the other Bayesian samplers, in that it can provide a graphical representation of the specified model through the use of nodes and arrows.
+This tool can be very useful in trying to explain how all the components of a model work together.

### 2020-01-17 Paper.Rmd
paper updates

@@
 \newpage
+
+# Bayesian Samplers
 
@@
 This tool can be very useful in trying to explain how all the components of a model work together.
+
+# Common Bayesian Computation Settings
+
+### Chains
+
+When using Markov Chain Monte Carlo methods to estimate Bayesian posteriors of parameters, the number of independent Markov chains to be run needs to be specified.
+It is recommended to use several chains to explore the space of the target distribution, as a single chain may not fully traverse the entire region on its own.
+For some simpler models that don't have irregular features in the target distribution, using one chain can suffice.
+Each sampler has an argument in its function call that specifies the number of chains.
+Some of the samplers have a default value if the number of chains isn't given.
+
+### Iterations
+
+The number of steps that each Markov chain will take is a very important setting to get right.
+If each chain is not set up to run long enough, the parameter space may not be fully explored and the estimates will be inaccurate.
+Checking convergence statistics can be helpful in determining if the number of iterations is sufficient.
+An argument specifiying the number of iterations for each chain can be inputted in the function call for each sampler.

### 2020-01-22 Paper.Rmd
paper updates

@@
   - Warmup
+    - STAN/Bugs uses warmup as a part of iterations, Jags/Greta/Nimble have it seperate.
   - Thinning
@@
 An argument specifiying the number of iterations for each chain can be inputted in the function call for each sampler.
+
+### Warmup
+
+Since our inferences are based on the assmumption that our draws are from the target distribution, it is wise to discard the samples at the beginning of each chain, since it takes time for the sampler to "settle down".
+The terms "warmup" and "burn-in" are often used to indicate the number of iterations that we will discard at the beginning of each chain when calculating statistics about our posterior distribution.
+We can choose this number in the function call for each sampler.
+In \texttt{OpenBUGS}, \texttt{Nimble}, and \texttt{Jags}, the argument is referred to as burnin, but in \texttt{STAN} and \texttt{Greta} it is called warmup.
+Some samplers (\texttt{Greta}, \texttt{Jags, Nimble}) treat this warmup argument as a number of iterations to run before starting to track the sampling results, but others (\texttt{STAN}, \texttt{BUGS}) have this be the actual number of iterations to discard out of the total number of specified iterations, so it is important to know which designation a sampler uses.
+
+### Thinning
+
+In some cases, there is high correlation between sequential samples in a given chain, which can bias the resulting posterior inference, which can be diagnosed using some graphical procedures.
+In this situation, it is wise to "thin" the results by only looking at every ith observation from a chain, such as every 10th observation.
+While each sampler defaults to including every simulated posterior draw, a user can specify a "thinning rate", which is the rate at which a sampler will keep memory of the observations.
+A thinning rate of 1 means that every observation will be kept, and a rate of 10 indicates that only every 10th observation will be used.
+
+### Initial Value Specification
+
+In order to start sampling, each Markov chain needs to have an initial value at which to start.
+These initial values can be either be specified, or can often be randomly generated based on the prior distributions for each parameter of interest.
+In many cases, it is easiest to have the initial values randomly generated by the sampler.
+However, in cases where the target distribution is very complex, it is possible that poorly chosen initial values can lead to a chain getting "stuck" in a difficult region, or diverging in the wrong direction.
+In these cases, a more strategic choice of initial values is advised.
+\texttt{Nimble} unfortunately requires for these values to be specified instead of randomly generated.
+Using \texttt{R} to randomly generate the values from the prior distributions can be used in this situation.

### 2020-01-23 Paper.Rmd
paper parameter update

@@
   - sigma / tau
+  - Parameter order
+    - Binomial Stan has (n,p), rest have (p,n)
 - Benchmarking

### 2020-01-24 Paper.Rmd
local change

--- 2020-01-22 Paper.Rmd
@@
   - sigma / tau
+  - Flat priors
 - Benchmarking

### 2020-01-24 Paper.Rmd

--- 2020-01-22 Paper.Rmd
@@
   - sigma / tau
+  - Flat priors
+  - Parameter order
+    - Binomial Stan has (n,p), rest have (p,n)
 - Benchmarking

### 2020-02-10 Paper.Rmd
paper additions

--- 2020-01-24 Paper.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(eval = FALSE)
 ```
@@
 ## Outline
+- Introduction to problem and history
+  - Bayesian paradigm
+  - How computing fits into bayesian paradigm
+  - Common terms that will be used in rest of paper
+  - Provide basic code example
 - General Bayesian computational settings
@@
   - Warmup
-    - STAN/Bugs uses warmup as a part of iterations, Jags/Greta/Nimble have it seperate.
+    - STAN/Bugs uses warmup as a part of iterations, Jags/greta/Nimble have it seperate.
   - Thinning
@@
 - Each sampler
-  - BUGS
+  - BUGS - 
     - WINE
@@
     - RDS file
-  - Greta
+  - greta
 - Common coding techniques for each language
+  - Model specification
     - Truncated normal
-  - Specifiying data and constants
-  - Matrix multiplication
-  - Initial value specification?
-  - sigma / tau
+    - matrix multiplication
     - Flat priors
+    - parameter order
+      - Binomial Stan has (n,p), rest have (p,n)
+    - Equations inside distribution parameters
+  - Data specification
+    - Nimble constants
+    - Nimble initial values
+    - initial values for each 
 - Benchmarking
@@
 Computationally, it is very slow compared to the other bayes methods, especially in simple cases.
-There are several types of semi-common models that Greta seems unable to properly handle in model specfication at the moment.
+There are several types of semi-common models that greta seems unable to properly handle in model specfication at the moment.
 Greta also attempts to make model specification very user-friendly, but uses some very common function names in the process that might override other often-used \texttt{R} functions.
@@
 We can choose this number in the function call for each sampler.
-In \texttt{OpenBUGS}, \texttt{Nimble}, and \texttt{Jags}, the argument is referred to as burnin, but in \texttt{STAN} and \texttt{Greta} it is called warmup.
-Some samplers (\texttt{Greta}, \texttt{Jags, Nimble}) treat this warmup argument as a number of iterations to run before starting to track the sampling results, but others (\texttt{STAN}, \texttt{BUGS}) have this be the actual number of iterations to discard out of the total number of specified iterations, so it is important to know which designation a sampler uses.
+In \texttt{OpenBUGS}, \texttt{Nimble}, and \texttt{Jags}, the argument is referred to as burnin, but in \texttt{STAN} and \texttt{greta} it is called warmup.
+Some samplers (\texttt{greta}, \texttt{Jags, Nimble}) treat this warmup argument as a number of iterations to run before starting to track the sampling results, but others (\texttt{STAN}, \texttt{BUGS}) have this be the actual number of iterations to discard out of the total number of specified iterations, so it is important to know which designation a sampler uses.
 
@@
 \texttt{Nimble} unfortunately requires for these values to be specified instead of randomly generated.
-Using \texttt{R} to randomly generate the values from the prior distributions can be used in this situation.
+Using \texttt{R} to randomly generate the values from the prior distributions can be used in this situation:
+
+```{r}
+nimble_inits <- list(
+  "alpha" = rnorm(1,0,1000),
+  "beta" = rnorm(1,0,1000),
+  "sigma" = runif(1,0,10)
+)
+
+nimbleMCMC("inits" = nimble_inits, ... )
+```
+
+If multiple chains are being used, a list containing a number of lists equalling number of chains can be used:
+
+```{r}
+nimble_inits <- list(
+  list(
+    "alpha" = rnorm(1,0,1000),
+    "beta" = rnorm(1,0,1000),
+    "sigma" = runif(1,0,10)
+  ),
+    list(
+    "alpha" = rnorm(1,0,1000),
+    "beta" = rnorm(1,0,1000),
+    "sigma" = runif(1,0,10)
+  )
+)
+```
+
+Alternatively, if a list containing the initial values for one chain is used, these values will be recycled for the remaining chains.
+
+# Common Tricks for Model and Data Specification
+
+When describing a bayesian model through code, there are instances were slightly different implementations are needed based on the sampler being used.
+Both in model and data specification, the user needs to understand the intricacies of the specific sampling tool being used in order to get the analysis to run properly.
+This section provides some helpful tips for implementing models in each of the different samplers being examined.
+
+## Model Specification
+
+The most important distinction between each of the samplers is the way that each allows for model specification text to be inputted.
+
+##### OpenBUGS
+In \texttt{OpenBUGS}, the earliest method created of the lot, the model specification code must be wrapped inside the expression \texttt{model\{\}}:
+
+```{r}
+model{
+  ...model code here...
+}
+```
+
+Since we are using the \texttt{R2OpenBUGS} library to run \texttt{OpenBUGS} through an \texttt{R} session, a text file with the model code is passed to the \texttt{bugs()} function.
+We can specify the model in \texttt{R} and create the appropriate text file:
+
+```{r}
+bugs_model <- function() {
+  ...model code here...
+}
+
+bugs.file <- file.path(tempdir(), "model.txt")
+write.model(bugs_model, bugs.file)
+
+bugs("model" = bugs.file, ... )
+```
+
+##### JAGS
+The \texttt{JAGS} sampler was intended to allow for model specification code to be almost identical to \texttt{BUGS}, so the implementation is very similar.
+There are two common approaches, one using the \texttt{rjags} package, and one using the \texttt{runjags} package.
+We use the latter, which allows for the model code to be specified in a multi-line character string:
+
+```{r}
+jags_model <- "
+  model{
+  ...model code here...
+  }
+"
+
+run.jags("model" = jags_model, ... )
+```
+
+##### Nimble
+Model specification in \texttt{Nimble} is extremely similar to \texttt{BUGS} as well.
+Model code is wrapped inside the \texttt{nimble} package function \texttt{nimbleCode()}, and then fed into \texttt{nimbleMCMC()} sampler function:
+
+```{r}
+nimble_model <- nimbleCode({
+  ...model code here...
+})
+
+nimbleMCMC("code" = nimble_model, ... )
+```
+
+##### STAN
+The STAN model specification language is significantly different than other samplers, in that a .stan file is fed to the \texttt{C++} compiler.
+A .stan file has a very unique model specification format, which divides code into "data", "parameter", and "model" blocks at a minimum.
+When needed, "transformed data" and "transformed parameters" blocks are also used.
+The basic format for model specification is provided below:
+
+```{r}
+data {
+  ...data specified here...        
+}
+
+parameters {
+  ...parameters specified here...                
+}
+
+model {
+  ...model code here...
+}
+```
+
+Each data variable and and parameter variable needs to be initialized in the appropriate block, along with associated domain limits.
+For example, if one were initializing a rate parameter for an exponential distribution called "lambda", it would be initialized with a lower bound of 0:
+
+```{r}
+parameters {
+  <lower=0> lambda;               
+}
+```
+
+A character string specifying the path to the .stan file is then passed to \texttt{stan()}.
+
+```{r}
+stan_file <- "file.stan"
+stan("file" = stan_file, ... )
+```
+
+## Data Specification and Input
+
+Specifying data for the bayesian samplers in \texttt{R} is very straightforward.
+In \texttt{OpenBUGS}, \texttt{JAGS}, and \texttt{STAN}, a list containing each data object or constant can be passed to the function that calls the sampler.
+For example, in \texttt{JAGS}, the data might constructed like this:
+
+```{r}
+n <-  10 # binomial n
+y <- rbinom(1, n, p)
+
+jags_data <- list(
+  "y" = y,
+  "n" = n
+)
+
+run.jags("data" = jags_data, ... )
+```
+
+In \texttt{Nimble}, the constants need to be specified in a separate list from the data.
+
+```{r}
+n <-  10 # binomial n
+y <- rbinom(1, n, p)
+
+nimble_data <- list(
+  "y" = y
+)
+
+nimble_constants <- list(
+  "n" = n
+)
+
+nimbleMCMC("data" = nimble_data, "constants" = nimble_constants, ... )
+```
+
+## Monitoring Parameters
+
+In \texttt{OpenBUGS}, \texttt{JAGS}, \texttt{Nimble}, and \texttt{STAN}, specifying parameters for the sampler to monitor is as easy as providing a character vector with the names of the parameters to monitor:
+
+```{r}
+jags_monitors <- jags_monitor <- c("alpha", "beta")
+
+run.jags("monitor" = jags_monitor, ... )
+```
+
+## Greta's Unique Specification Language
+
+Implementing models in \texttt{greta} is very unique in that each data variable or parameter is initialized seperately with its own line of \texttt{R} code.
+Each data object in \texttt{R} is redefined using the \texttt{as\_data()} function in the \texttt{greta} package:
+```{r}
+y <- as_data(rpois(5, theta))
+```
+
+Each parameter is defined using a function indicating the prior distribution and the chosen prior parameters.
+
+```{r}
+theta <- gamma(3,1)
+```
+
+Once all the variables have been defined, the relationship between the data and the unknown parameters is established using the \texttt{distribution()} function:
+
+```{r}
+distribution(y) <- poisson(theta)
+```
+
+Finally, the greta model is constructed using \texttt{model()} function, which takes each of the monitored parameters as arguments:
+
+```{r}
+greta_model <- model(theta)
+
+mcmc("model" = greta_model, ... )
+```
+
+# How to Specify Tricky Models
+
+Each sampler has its own unique methods for using certain bayesian model details. We provide illustrations for some of the most common ones.
+
+##### Truncated Distribution
+
+Using a truncated distribution is a common practice in bayesian modeling, especially when specifying priors for parameters with positive domains.
+The syntax for specifying a truncated distribution is slightly different for each sampling method.
+Below we are using a truncated normal distribution with a lower bound of zero and no upper bound:
+
+```{r}
+## Truncated normal distribution in OpenBUGS
+tau ~ dnorm(0,1) I(0, )
+
+## Model Text for OpenBUGS using R2OpenBUGS
+tau ~ dnorm(0,1) %_% I(0, )
+
+## JAGS
+tau ~ dnorm(0,1) I(0, )
+
+## Nimble
+tau ~ T(dnorm(0,1), 0,)
+
+## Greta
+tau <- normal(0,1, truncation = c(0,Inf))
+
+## STAN
+# Parameter block initialization
+real<lower=0> tau;
+
+# Model block specification
+tau ~ normal(0,1);
+
+```
+
+##### Flat Priors
+
+Sometimes using a flat, diffuse prior is appropriate.
+Most of the samplers allow for improper flat priors, although \texttt{JAGS} does not.
+In \texttt{STAN}, a parameter has is given a flat, improper prior by default if it doesn't have a distribution assigned.
+Simply declaring the parameter will do the job.
+```{r}
+## Flat prior in OpenBUGS
+theta ~ dflat()
+
+## Nimble
+theta ~ dflat()
+
+## Greta
+theta <- variable()
+
+## STAN
+# Parameter block initialization
+real theta;
+
+```
+
+##### Distribution Parameter Order
+
+Some samplers use different parameter orders for certain distributions.
+
+Some use the $\sigma$ parameterization of the normal distribution, where $\sigma$ is the standard deviation, and others use $\tau$ instead, where $\tau = 1 / \sigma^2$.
+Among bayesian samplers, \texttt{JAGS}, \texttt{OpenBUGS}, and \texttt{Nimble} use the $\tau$ parameteriztion, while \texttt{STAN} and \texttt{greta} use $\sigma$.
+
+\textit{\textbf{Normal}}: The $N(\mu,\sigma)$ parameterization is used by \texttt{STAN} and \texttt{greta}, where $\sigma$ is the standard deviation.
+The $N(\mu,\tau)$ parameterization is used by \texttt{JAGS}, \texttt{OpenBUGS}, and \texttt{Nimble}, which uses the precision parameter $\tau = 1 / \sigma^2$.
+The same parameterization guidelines are the same for the student-t distribution as well.
+
+\textit{\textbf{Binomial}}: The $Bin(n,p)$ parameterization is used by \texttt{OpenBUGS}, \texttt{JAGS}, and \texttt{Nimble}, while the $Bin(p,n)$ parameterization is used by \texttt{STAN} and \texttt{greta}.
+
+\textit{\textbf{Gamma}}: The $Gam(\alpha,\lambda)$ parameterization, where $\alpha$ is a shape parameter and $\lambda$ is a rate parameter, is used by \texttt{OpenBUGS}, \texttt{JAGS}, \texttt{Nimble}, and \texttt{greta}.
+The $Gam(\alpha,\beta)$ parameterization is used by \texttt{STAN}, using the scale parameter $\beta = 1 / \lambda$.
+
+\textit{\textbf{Weibull}}: The $Weibull(v,\lambda)$ parameterization is used by \texttt{OpenBUGS}, \texttt{JAGS}, \texttt{Nimble}, and \texttt{greta}, where $\lambda$ is a shape parameter.
+The $Weibull(v,\sigma)$ parameterization is used by \texttt{STAN}, where $\sigma = 1 / \lambda$.
+
+##### Equations inside distribution parameters
+
+In order to utilized calculated parameters inside \texttt{OpenBUGS}, a variable must be defined using the assignment operator that is the result of the calculation:
+
+```{r}
+y_hat <- alpha + beta * x
+y ~ dnorm(y_hat, tau)
+```
+
+Fortunately, the other samplers allow for variable calculations inside the actual parameter definition of a distribution function. Here is an example in \texttt{JAGS}:
+
+```{r}
+y ~ dnorm(alpha + beta * x, tau)
+```

### 2020-02-11 Paper.Rmd
paper update no outline

@@
 ---
-title: "Paper"
+title: "Bayes Computational Method"
 output: pdf_document
@@
 
-## Outline
-- Introduction to problem and history
-  - Bayesian paradigm
-  - How computing fits into bayesian paradigm
-  - Common terms that will be used in rest of paper
-  - Provide basic code example
-- General Bayesian computational settings
-  - Chains
-    - Greta
-  - Iterations
-  - Warmup
-    - STAN/Bugs uses warmup as a part of iterations, Jags/greta/Nimble have it seperate.
-  - Thinning
-  - initial value specification
-- Each sampler
-  - BUGS - 
-    - WINE
-    - Debugging
-  - Jags
-  - Nimble
-    - Crashes
-  - STAN
-    - RDS file
-  - greta
-- Common coding techniques for each language
-  - Model specification
-    - Truncated normal
-    - matrix multiplication
-    - Flat priors
-    - parameter order
-      - Binomial Stan has (n,p), rest have (p,n)
-    - Equations inside distribution parameters
-  - Data specification
-    - Nimble constants
-    - Nimble initial values
-    - initial values for each 
-- Benchmarking
-  - Run time
-  - Memory allocation
-  - Computational accuracy
-  - Benchmarking files
-- Folder design
-  - General model type
-    - Simple conjugate models
-    - Regression models
-    - Time Series models
-    - Survival models
-    - 8 Schools
-  - Specific model type
-  - Scripts for each method 
-  - Benchmarking
-- Here package
-- Models Used
-
-\newpage
-
 # Bayesian Samplers
 
-## BUGS
+## OpenBUGS
 

### 2020-02-24 regression-models/hierarchical-logistic-regression/hierarchical-logistic-regression-benchmark.R
hierarchical log reg greta changes

@@ -1,0 +1,46 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("bench")
+library("here")
+library("rjags"); library("runjags")
+library("R2OpenBUGS")
+library("nimble")
+library("rstan"); rstan_options(auto_write = FALSE)
+options("bayes_benchmark" = TRUE)
+
+## Source code
+################################################################################
+
+source(here("regression-models","hierarchical-logistic-regression","hierarchical-logistic-regression-jags.R"))
+source(here("regression-models","hierarchical-logistic-regression","hierarchical-logistic-regression-bugs.R"))
+source(here("regression-models","hierarchical-logistic-regression","hierarchical-logistic-regression-nimble.R"))
+source(here("regression-models","hierarchical-logistic-regression","hierarchical-logistic-regression-stan.R"))
+# source(here("regression-models","hierarchical-logistic-regression","hierarchical-logistic-regression-greta.R"))
+
+options("bayes_benchmark" = FALSE)
+
+## Set Parameters
+################################################################################
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+## benchmarking
+###################################################################################
+
+num_iterations <- 1 # Number of times you want to benchmark each model\
+
+rds_file_location <- here("simple models", "hierarchical-regression-models","hierarchical-logistic-regression","hierarchical-logistic-regression.rds")
+
+source(here("benchmark-function.R"))
+
+# choose to include stan compilation time with argument "stan_compile = TRUE"
+run_benchmark(rds_file_location)
+run_benchmark(rds_file_location, stan_compile = TRUE)
+
+
+
+

### 2020-02-24 regression-models/hierarchical-logistic-regression/hierarchical-logistic-regression-greta.R

@@ -13,7 +13,7 @@ library("here")
 ################################################################################
 n <- 30L # sample size
 d <- 2L
-l <- 2L
+l <- 3L
 ll <- sapply(1:l, function(x) rep(x,n / l)) %>% as.numeric()
 
 set.seed(1)
@@ -39,17 +39,27 @@ y <- as_data(y)
 
 
 #### NOT COMPLETE ############
-mu <- normal(0, 1000, dim = c(d,1))
+mu <- normal(0, 1000, dim = c(1,d))
 sigma <- normal(0,1000, dim = c(d,1), truncation = c(0,Inf))
-beta_0 <- normal(rep(mu, l), rep(sigma,l), dim = c(d,l))
-beta <- rep(beta, l)
+mu_matrix <- zeros(3,2)
+mu_matrix[1,] <- mu
+mu_matrix[2,] <- mu
+mu_matrix[3,] <- mu
+sigma_matrix <- zeros(3,2)
+sigma_matrix[1,] <- sigma
+sigma_matrix[2,] <- sigma
+sigma_matrix[3,] <- sigma
+beta <- normal(mu_matrix, sigma_matrix)
 
-beta[,1] <- normal(mu, sigma, dim = c(d,1))
-beta_0 <- normal(mu, sigma, dim = c(d,1))
+theta <- zeros(n,1)
+i <- 1
+for (i in 1:n) {
+  theta[i,] <- x[i,] %*% t(beta[ll[i],])
+}
 
-distribution(y) <- normal(alpha + beta * x, sigma)
+distribution(y) <- bernoulli(ilogit(theta))
 
-greta_model <- model(alpha, beta, sigma)
+greta_model <- model(mu, beta, sigma, theta)
 
 plot(greta_model)
 

### 2020-02-24 regression-models/hierarchical-logistic-regression/hierarchical-logistic-regression-jags.R

@@ -15,7 +15,7 @@ library("here")
 
 n <- 30L # sample size
 d <- 2L
-l <- 2L
+l <- 3L
 ll <- sapply(1:l, function(x) rep(x,n / l)) %>% as.numeric()
 
 set.seed(1)

### 2020-02-24 regression-models/hierarchical-logistic-regression/hierarchical-logistic-regression-nimble.R

@@ -13,7 +13,7 @@ library("bench")
 
 n <- 30L # sample size
 d <- 2L
-l <- 2L
+l <- 3L
 ll <- sapply(1:l, function(x) rep(x,n / l)) %>% as.numeric()
 
 set.seed(1)
@@ -64,10 +64,10 @@ n_iter <- 1e4L
 n_warmup <- 1e3L
 
 nimble_inits <- list(
-  "mu" = rnorm(d,0,1000),
+  "mu" = rnorm(d,0,10),
   "theta" = rbeta(n,1,1),
-  "tau" = rep(1 / (abs(rnorm(1,0,1000)) ^ 2), d),
-  "beta" = matrix(rnorm(l * d,0,1000), nrow = l, ncol = d)
+  "tau" = rep(1 / (abs(rnorm(1,0,10)) ^ 2), d),
+  "beta" = matrix(rnorm(l * d,0,10), nrow = l, ncol = d)
 )
 
 

### 2020-02-24 regression-models/hierarchical-logistic-regression/hierarchical-logistic-regression-stan.R

@@ -15,7 +15,7 @@ library("bench")
 
 n <- 30L # sample size
 d <- 2L
-l <- 2L
+l <- 3L
 ll <- sapply(1:l, function(x) rep(x,n / l)) %>% as.numeric()
 
 set.seed(1)
@@ -42,54 +42,47 @@ stan_file <- here("regression-models", "hierarchical-logistic-regression", "hier
 
 
 
-## fit model
+## configure model settings
 ################################################################################
 
 n_chains <- 4L
 n_iter <- 1e4L
 n_warmup <- 1e3L
 
-stan_fit <- stan(
-  "file" = stan_file, "data" = stan_data, 
-  "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
-)
 
-
-
-
-## assess fit
+## fit model
 ################################################################################
-
-str(stan_fit, 2)
-
-summary(stan_fit)$summary
-get_posterior_mean(stan_fit)
-stan_dens(stan_fit) + theme_bw()
-stan_fit %>% as.array() %>% bayesplot::mcmc_dens()
-
-
-## assess convergence issues 
-###################################################################################
-
-stan_fit %>% as.array() %>% mcmc_acf_bar()
-stan_fit %>% as.array() %>% mcmc_pairs()
-stan_fit %>% as.array() %>% mcmc_trace()
-
-# see each chain
-stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)
-
-
-
-## benchmarking
-###################################################################################
-
-
-bench_results <- mark(
+if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+  
   stan_fit <- stan(
     "file" = stan_file, "data" = stan_data, 
     "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
-  ),
-  iterations = 3
-)
-bench_results[1,2:9]
+  )
+  
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(stan_fit)$summary
+  get_posterior_mean(stan_fit)
+  stan_dens(stan_fit) + theme_bw()
+  stan_fit %>% as.array() %>% bayesplot::mcmc_dens()
+  
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  stan_fit %>% as.array() %>% mcmc_acf_bar()
[…]

### 2020-02-26 survival-models/cox-exponential-benchmark.R
model code updates

@@ -19,7 +19,7 @@ source(here("survival-models", "cox-exponential-jags.R"))
 source(here("survival-models", "cox-exponential-bugs.R"))
 source(here("survival-models", "cox-exponential-nimble.R"))
 source(here("survival-models", "cox-exponential-stan.R"))
-source(here("survival-models", "cox-exponential-greta.R"))
+# source(here("survival-models", "cox-exponential-greta.R"))
 
 options("bayes_benchmark" = FALSE)
 

### 2020-02-26 time-series/ma-model/ma-model-jags.R

@@ -72,7 +72,6 @@ if (!currently_benchmarking()) {
   ) 
   
   
-  
   ## assess fit
   ################################################################################
   

### 2020-02-26 time-series/ma-model/ma-model-nimble.R

@@ -59,9 +59,9 @@ n_iter <- 1e4L
 n_warmup <- 1e3L
 
 nimble_inits <- list(
-  "mu" = rnorm(1,0,1000),
-  "theta" = rnorm(1,0,1000),
-  "tau" = rnorm(1,0,0.001) %>% abs()
+  "mu" = rnorm(1,0,10),
+  "theta" = rnorm(1,0,10),
+  "tau" = rnorm(1,0,0.01) %>% abs()
 )
 
 

### 2020-03-04 Paper.Rmd
hierarchical log reg greta changes

@@
 ---
-title: "Bayes Computational Method"
+title: "Bayes Computational Methods"
+author: "Evan Miyakawa"
 output: pdf_document
@@
 ```
+
+# Introduction
+
+## The Bayes Paradigm
+
+In Bayesian statistical inference, we are interested in finding the posterior distribution of a certain parameter, given a prior distribution and a likelihood for that parameter from a sample of data.
+Bayes Rule states that, given events $A$ and $B$, we can find the conditional probability of $A$ given $B$ using $$ P(A | B) = \frac{P(B|A)}{P(B)}.$$ More formally, we can derive the posterior distribution of a parameter $\theta$, $h(\theta | x)$, by $$h(\theta | x_1, x_2, \dots, x_n) = \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}},$$ where $\ell(x | \theta)$ is proportional to the likelihood of $\theta$, given the data, and $\pi(\theta)$ is the prior distribution of $\theta$.
+For example, suppose that $X_i \sim Bern(\theta)$, and we are interested in inference on $\theta$.
+A reasonable prior to assign to $\theta$ would be a $Beta(\alpha, \beta)$, distribution, given that $\theta$ is constrained to be between 0 and 1, which is a property of the Beta distribution.
+Therefore, given that $\theta \sim Beta(\alpha, \beta)$ for some known values of $\alpha$ and $\beta$, then
+\begin{align*}
+  h(\theta | x) &= \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}} \\
+  &= \frac{\theta ^ {\sum_{i=1}^{n} x_i} (1 - \theta) ^ {n - \sum_{i=1}^{n} x_i} \frac{\Gamma (\alpha + \beta)}{\Gamma(\alpha) + \Gamma(\beta)} \theta^{\alpha - 1} (1 - \theta)^{\beta - 1}}{\int \theta ^ {\sum_{i=1}^{n} x_i} (1 - \theta) ^ {n - \sum_{i=1}^{n} x_i} \frac{\Gamma (\alpha + \beta)}{\Gamma(\alpha) + \Gamma(\beta)} \theta^{\alpha - 1} (1 - \theta)^{\beta - 1} d \theta} \\
+  &\propto \theta ^ {\sum (x_i) + \alpha - 1} (1 - \theta) ^ {n - \sum (x_i) + \beta - 1},
+\end{align*}
+which is the kernel of a $Beta\left(\sum (x_i) + \alpha, n - \sum (x_i) + \beta \right)$ distribution.
+Therefore, we can analytically calculate the posterior distribution of $\theta$ as $$h(\theta | x) \sim Beta(\sum (x_i) + \alpha, n - \sum (x_i) + \beta).$$
+
+In more complicated scenarios, it is not easy to calculate the posterior distribution of a parameter by hand, and in many cases it is impossible. This is where computing can help us.
 

### 2020-03-04 Paper.Rmd
MCMC paper updates

@@
 
-In more complicated scenarios, it is not easy to calculate the posterior distribution of a parameter by hand, and in many cases it is impossible. This is where computing can help us.
+In more complicated scenarios, it is not easy to calculate the posterior distribution of a parameter by hand, and in many cases it is impossible.
+This is where computing can help us.
+If we could somehow computationally grab samples from the posterior distribution we are interested in, then we could obtain most of the statistics that we are interested in.
+Markov Chain Monte Carlo (referred to as MCMC), is a common approach that allows for drawing samples of a parameter from a distribution that is approximately equivalent to the target posterior distribution of that parameter.
+An essential property of any Markov chain is that the distribution of each draw only depends on the most recent value drawn before it.
+As we keep obtaining draws, each draw is probabilisticly more likely to have come from the target posterior distribution.
+Based on the convergence properties of MCMC, we can expect each draw to be coming from the real distribution of the parameter we are interested, once we have let the Markov chain run long enough.
 

### 2020-03-04 Paper.Rmd
adding outlines

--- 2020-02-11 Paper.Rmd
@@
 ```
+
+# Introduction
+
+## The Bayes Paradigm
+
+In Bayesian statistical inference, we are interested in finding the posterior distribution of a certain parameter, given a prior distribution and a likelihood for that parameter from a sample of data.
+Bayes Rule states that, given a parameter of interest $\theta$ and some known data $x$, we can find the conditional probability of $\theta$ given $y$ using $$ p(\theta | x) = \frac{p(x | \theta) p(\theta)}{p(x)},$$ where in the discrete case $p(x) = \sum_\theta p(\theta) p(x|\theta),$ and in the continuous case, $p(x) = \int p(\theta) p(y|\theta) d\theta$.
+More formally, we can derive the posterior distribution of a parameter $\theta$, $p(\theta | x)$, by $$p(\theta | x_1, x_2, \dots, x_n) = \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}},$$ where $\ell(x | \theta)$ is proportional to the likelihood of $\theta$, conditioned on the data, and $\pi(\theta)$ is the prior distribution of $\theta$.
+For example, suppose that $X_i \sim \text{Bern}(\theta)$, and we are interested in inference on $\theta$.
+A reasonable prior to assign to $\theta$ would be a $\text{Beta}(\alpha, \beta)$, distribution, given that $\theta$ is constrained to be between 0 and 1, which is a property of the Beta distribution.
+Therefore, given that $\theta \sim \text{Beta}(\alpha, \beta)$ for some known values of $\alpha$ and $\beta$, then
+\begin{align*}
+  p(\theta | x) &= \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}} \\
+  &= \frac{\theta ^ {\sum_{i=1}^{n} x_i} (1 - \theta) ^ {n - \sum_{i=1}^{n} x_i} \frac{\Gamma (\alpha + \beta)}{\Gamma(\alpha) + \Gamma(\beta)} \theta^{\alpha - 1} (1 - \theta)^{\beta - 1}}{\int \theta ^ {\sum_{i=1}^{n} x_i} (1 - \theta) ^ {n - \sum_{i=1}^{n} x_i} \frac{\Gamma (\alpha + \beta)}{\Gamma(\alpha) + \Gamma(\beta)} \theta^{\alpha - 1} (1 - \theta)^{\beta - 1} d \theta} \\
+  &\propto \theta ^ {\sum (x_i) + \alpha - 1} (1 - \theta) ^ {n - \sum (x_i) + \beta - 1},
+\end{align*}
+which is the kernel of a $\text{Beta}\left(\sum (x_i) + \alpha, n - \sum (x_i) + \beta \right)$ distribution.
+Therefore, we can analytically calculate the posterior distribution of $\theta$ as $$p(\theta | x) \sim \text{Beta}\left(\sum (x_i) + \alpha, n - \sum (x_i) + \beta\right).$$
+
+In more complicated scenarios, it is not easy to calculate the posterior distribution of a parameter by hand, and in many cases it is impossible. This is where computing can help us.
 

### 2020-03-04 sim-save.R
updates

@@ -1,6 +1,8 @@
 x <- run_benchmark(rds_file_location)
-ma_model_results <- x
-saveRDS(ma_model_results, file = "ma-model-results.rds")
+hierarchical_logistic_regression_results <- x
+saveRDS(hierarchical_logistic_regression_results, 
+        file = "hierarchical-logistic-regression-results.rds")
+
 x
 
 # readRDS("exponential-gamma-results.rds")

### 2020-03-05 survival-models/cox-exponential-bugs.R
survival updates

@@ -18,18 +18,23 @@ data(veteran) # load spatial data set
 t <- veteran$time # observed / censored event times
 x <- veteran$trt # single covariate
 not_censored <- veteran$status # indicator variable for those who weren't censored
-is_censored <- 1 - status # indicator variable for those who were censored
+is_censored <- 1 - not_censored # indicator variable for those who were censored
+N <- length(t)
 
 # Here we create variables that will allow us to handle censored data in the bayes model
-t_censor <- t + status
-t[status == 0] <- NA
-
+num_uncensored <- sum(not_censored)
+start_censored <- num_uncensored + 1
+t <- t[order(is_censored)]
+x <- x[order(is_censored)]
+t_censor <- t
+t[start_censored:N] <- NA
 
 bugs_data <- list(
   "t" = t,
-  "x" = x,
   "t_censor" = t_censor,
-  "is_censored" = is_censored,
+  "x" = x,
+  "num_uncensored" = num_uncensored,
+  "start_censored" = num_uncensored + 1,
   "N" = length(t)
 )
 
@@ -48,9 +53,13 @@ bugs_model <- function() {
 }
 
 bugs_model <- function() {
-  for (i in 1:N) {
+  for (i in 1:num_uncensored) {
     lambda[i] <- exp(beta * x[i])
-    t[i] ~ dexp(lambda[i]) %_% I(t_censor[i], )
+    t[i] ~ dexp(lambda[i])
+  }
+  for (i in start_censored:N) {
+    lambda[i] <- exp(beta * x[i])
+    t[i] ~ dexp(lambda[i]) %_% I(t_censor[i],)
   }
   beta ~ dnorm(0,0.0001)
 }
@@ -79,13 +88,13 @@ if (getwd() == "/Users/evanmiyakawa/hubiC/Git Projects/Bayes-Computational-Metho
 ## configure model settings
 ################################################################################
 
-n_chains <- 1L
-n_iter <- 5000
-n_warmup <- 1000
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
 
-t_inits <- veteran$time + 1
-t_inits[status == 1] <- NA
-bugs_inits <- replicate(n_chains, list("t" = t_inits), simplify = FALSE)
+t_inits <- t_censor + 1
+t_inits[1:num_uncensored] <- NA
[…]

### 2020-03-05 survival-models/cox-exponential-jags.R

@@ -16,23 +16,24 @@ library("survival")
 
 data(veteran) # load spatial data set
 
-t <- veteran$time # observed / censored event times
-x <- veteran$trt # single covariate
-not_censored <- veteran$status # indicator variable for those who weren't censored
-is_censored <- 1 - not_censored # indicator variable for those who were censored
-
-# Here we create variables that will allow us to handle censored data in the bayes model
-t_censor <- t + not_censored
-t[not_censored == 0] <- NA
-
-
-jags_data <- list(
-  "t" = t,
-  "x" = x,
-  "t_censor" = t_censor,
-  "is_censored" = is_censored,
-  "N" = length(t)
-)
+# t <- veteran$time # observed / censored event times
+# x <- veteran$trt # single covariate
+# not_censored <- veteran$status # indicator variable for those who weren't censored
+# not_censored <- rep(1, length(not_censored))
+# is_censored <- 1 - not_censored # indicator variable for those who were censored
+# 
+# # Here we create variables that will allow us to handle censored data in the bayes model
+# t_censor <- t + not_censored
+# t[not_censored == 0] <- NA
+# 
+# 
+# jags_data <- list(
+#   "t" = t,
+#   "x" = x,
+#   "t_censor" = t_censor,
+#   "is_censored" = is_censored,
+#   "N" = length(t)
+# )
 
 t <- veteran$time # observed / censored event times
 x <- veteran$trt # single covariate
@@ -62,16 +63,16 @@ jags_data <- list(
 ## specify jags model
 ################################################################################
 
-jags_model <- "
-  model{
-    for (i in 1:N) {
-      is_censored[i] ~ dinterval(t[i], t_censor[i])
-      lambda[i] <- exp(beta * x[i])
-      t[i] ~ dexp(lambda[i])
-    }
-    beta ~ dnorm(0,0.0001)
-  }
-"
+# jags_model <- "
+#   model{
+#     for (i in 1:N) {
+#       is_censored[i] ~ dinterval(t[i], t_censor[i])
+#       lambda[i] <- exp(beta * x[i])
+#       t[i] ~ dexp(lambda[i])
+#     }
+#     beta ~ dnorm(0,0.0001)
+#   }
+# "
 
 jags_model <- "
   model{

### 2020-03-05 survival-models/cox-exponential-nimble.R

@@ -12,23 +12,51 @@ library("survival")
 ## generate/specify data
 ################################################################################
 
-data(veteran)
-vet <- mutate(veteran, AG = ifelse((age < 60), "LT60", "OV60"),
-              AG = factor(AG),
-              trt = factor(trt,labels=c("standard","test")),
-              prior = factor(prior,labels=c("N0","Yes")))
-vet <- filter(vet, status == 1)
+data(veteran) # load spatial data set
 
-y <- vet$time
-trt <- vet$trt %>% as.numeric() - 1
+# t <- veteran$time # observed / censored event times
+# x <- veteran$trt # single covariate
+# not_censored <- veteran$status # indicator variable for those who weren't censored
+# not_censored <- rep(1, length(not_censored))
+# is_censored <- 1 - not_censored # indicator variable for those who were censored
+# 
+# # Here we create variables that will allow us to handle censored data in the bayes model
+# t_censor <- t + not_censored
+# t[not_censored == 0] <- NA
+# 
+# 
+# jags_data <- list(
+#   "t" = t,
+#   "x" = x,
+#   "t_censor" = t_censor,
+#   "is_censored" = is_censored,
+#   "N" = length(t)
+# )
+
+t <- veteran$time # observed / censored event times
+x <- veteran$trt # single covariate
+not_censored <- veteran$status # indicator variable for those who weren't censored
+is_censored <- 1 - not_censored # indicator variable for those who were censored
+N <- length(t)
+
+# Here we create variables that will allow us to handle censored data in the bayes model
+num_uncensored <- sum(not_censored)
+start_censored <- num_uncensored + 1
+t <- t[order(is_censored)]
+x <- x[order(is_censored)]
+t_censor <- t
+t[start_censored:N] <- NA
 
 nimble_data <- list(
-  "t" = y,
-  "x" = trt
+  "t" = t,
+  "t_censor" = t_censor,
+  "x" = x
 )
 
 nimble_constants <- list(
-  "N" = length(y)
+  "num_uncensored" = num_uncensored,
+  "start_censored" = num_uncensored + 1,
+  "N" = length(t)
 )
 
 
@@ -36,9 +64,13 @@ nimble_constants <- list(
 ################################################################################
[…]

### 2020-03-05 survival-models/cox-exponential-stan.R

@@ -14,6 +14,8 @@ library("survival")
 
 data(veteran) # load spatial data set
 
+
+
 t <- veteran$time # observed / censored event times
 x <- veteran$trt # single covariate
 not_censored <- veteran$status # indicator variable for those who weren't censored
@@ -22,24 +24,29 @@ N <- length(t)
 
 # Here we create variables that will allow us to handle censored data in the bayes model
 num_uncensored <- sum(not_censored)
+num_censored <- length(t) - num_uncensored
 start_censored <- num_uncensored + 1
-num_censored <- sum(is_censored)
 t <- t[order(is_censored)]
 x <- x[order(is_censored)]
 t_censor <- t
-
-t_censor <- t[start_censored:N]
-t <- t[1:num_uncensored]
-x_c <- x[start_censored:N]
-x <- x[1:num_uncensored]
+t[start_censored:N] <- NA
 
 stan_data <- list(
   "t" = t,
   "t_censor" = t_censor,
   "x" = x,
-  "x_c" = x_c,
   "num_uncensored" = num_uncensored,
-  "num_censored" = num_censored
+  "start_censored" = num_uncensored + 1,
+  "N" = length(t)
+)
+
+stan_data <- list(
+  "t" = t[1:num_uncensored],
+  "t_censor" = t_censor[(num_uncensored + 1):length(t)],
+  "x" = x,
+  "num_uncensored" = num_uncensored,
+  "num_censored" = length(t) - num_uncensored,
+  "N" = length(t)
 )
 
 
@@ -62,13 +69,9 @@ n_iter <- 1e4L
 n_warmup <- 1e3L
 
 
-t_inits <- t_censor + 1
-stan_inits <- list(
-  list("t_u" = t_inits),
-  list("t_u" = t_inits),
-  list("t_u" = t_inits),
-  list("t_u" = t_inits)
-)
+t_inits <- stan_data$t_censor + 1
+stan_inits <- replicate(4,list("t_c" = t_inits), simplify = FALSE)
+
 ## fit model
 ################################################################################
 source(here("currently-benchmarking.R"))

### 2020-03-24 Paper.Rmd
updates
survival updates
stan paper update

@@
 
+## Bayesian Computation
+
 In more complicated scenarios, it is not easy to calculate the posterior distribution of a parameter by hand, and in many cases it is impossible.
@@
 If we could somehow computationally grab samples from the posterior distribution we are interested in, then we could obtain most of the statistics that we are interested in.
+
+### History of Bayesian computation up to MCMC
+
+### Markov Chain Monte Carlo
+
 Markov Chain Monte Carlo (referred to as MCMC), is a common approach that allows for drawing samples of a parameter from a distribution that is approximately equivalent to the target posterior distribution of that parameter.
+It is an iterative algorithm, where each realization gets closer to the target density we are interested in.
 An essential property of any Markov chain is that the distribution of each draw only depends on the most recent value drawn before it.
@@
 Based on the convergence properties of MCMC, we can expect each draw to be coming from the real distribution of the parameter we are interested, once we have let the Markov chain run long enough.
+
+- MCMC is the only suitable method when we don't have a conjugate prior.
+
+### History up through BUGS (include slice sampling)
+
+- BUGS project began in 1989
+- BUGS language contains for a somewhat concise list of functions and distributions used for specifying relationships between 
+- The BUGS language only has a somewhat limited syntax, but is able to describe extremly complex models using this syntax.
+- While the programs that use BUGS-type language are continually evolving, the syntax has stayed very stable.
+- BUGS language is declarative, which means the order of statements doesn't matter, as opposed to a procedural language.
+- A main initial purpose for the creation of bugs was to allow for graphical models -- models that can be specified through click and point software that creates nodes, or expressing graphics of models that already exist for easier understanding.
+- Six steps in running BUGS models
+  1. Checking syntax of model specification
+  2. Loading data
+  3. Compiling the BUGS model
+  4. Starting the simulation at appropriate initial values
+  5. Simulate unknown parameters
+  6. Gather results and report summary statistics, along with other appropriate output
+- WinBUGS was the first widely used program to utilize the BUGS language.
+However, OpenBUGS is the more recently maintained version, which communicates better with other software and is more portable.
+- R2OpenBUGS is the R package that allows for R to interface with OpenBUGS
+-
+
+### Gibbs Sampler
+
+- One of most common algorithms for simulating MCMCs.
+- Special case of Metropolis-Hastings alg
+- Generates a multi-dimensional markov chain by splitting the vector of random unknown variables and sampling each subvector one by one, conditional on the most recent values of the other variables.
+- Algorithm: If we have a vector of unknown parameters $\theta$, where $\theta = (\theta_1, \theta_2, \dots, \theta_k)$,
+  1. Choose arbitrary starting values $\theta_1^{(0)}, \theta_2^{(0)}, \dots, \theta_k^{(0)}$ 
+  2. Complete one iteration of the sampler by getting new values for $\theta$ using the following:
+    - Sample a new value for $\theta_1$, from the full conditional distribution of $\theta_1$, given the most recent values of all other elements of $\theta$ and the data:
+    $$\theta_1^{(1)} \sim p(\theta_1 | \theta_2^{(0)}, \theta_3^{(0)}, \dots , \theta_k^{(0)}, y).$$
+    - Sample a new value $\theta_2^{(1)}$ for the second component of $\theta$, from the full conditional distribution $p(\theta_2 | \theta_1^{(1)}, \theta_3^{(0)}, \dots , \theta_k^{(0)}, y)$.
+    - Continue this process for the remainder of the $k$ elements of $\theta$.
+  3. Repeat step 2 for the total decided number of iterations that the sampler will run.
+
+- The Gibbs sampling strategy often allows for sampling from high-dimensional spaces to be simplified to drawing from one-dimensional distributions.
+- Weakness: because it uses a random-walk behavior, if two variables are highly correlated, traversing the space can take a very long time.
+
+### Metropolis
+
+### HMC
+
+Radford-Neal HMC Chapter:
+- Hamiltonian Monte Carlo uses Hamiltonian Dynamics
+  - Visualize the dynamics as a frictionless puck that slides over a surface of varying height
+    - The state of the system has three elements: position, momentum, and potential energy
+  - The position corresponds to variables of interest
+  - The potential energy will be minus the log of the probability density for these variables
+  - There is one momentum variable for each position variable
+- Using "Hamilton's equations", the movement of the puck is computed. 
+- For computational implementation, these equations must be approximated by discretizing time, using some small stepsize.
+- To use Hamiltonian dynamics to sample from a distribution using MCMC, we need to translate the density function for the distribution to a potential energy function.
+The variables of interest are now position variables
+- We simulate a markov chain where each iteration resamples the momentum variable associated with each position variable and then does a Metroplis update with the proposal found using Hamiltonian dynamics
+  - Two steps for each iteration:
+    1. New values for the momentum variables are randomly drawn from their Gaussian distribution, independent of the current values of the position variables.
+    2.
+A metroplis update is performed, using hamiltonian dynamics to propose a new state.
+The proposed state is either accepted or rejected based on the HMC probability function used.
+If the proposed state is rejected, the next state is the same as the current one.
+- The HMC algorithm typically will not be trapped in some subset of the state space, which allows for it to always asymptotically converge to the target distribution, and quickly.
+- Tuning Parameters for HMC:
+  - Stepsize - the leapfrog stepsize. Too large, and it will lead to a low acceptance rate for proposed states. Too small and it will be slow computationally.
+  - Trajectory length -
+
+### STAN
+STAN 2015
+- Based on a probabilistic programming language that is more flexible and expressive than BUGS/JAGS.
+- Based on HMC
+- Works in command line, Python, and R
 
@@
 In terms of model specification, JAGS is very similar to OpenBugs, but works more seemlessly in R using the package \texttt{rjags}.
+JAGS was developed independently of OpenBUGS, but uses the BUGS language for most of its model specification syntax.
 The \texttt{runjags} is an optional package that allows for the JAGS model script to be inputted as a string, and run using the \texttt{run.jags()} function.
@@
 
+# Alternative solution
+tau ~ normal(0,1) T[0,]
+
 ```

### 2020-04-08 HMC Talk Notes.Rmd
survival models fixed

---
title: "HMC Talk Notes"
author: "Evan Miyakawa"
date: "4/7/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- When you have more complexity than information (like in big data with more variables than observations), having a prior allows us to do inference that we couldn't do otherwise 
- Doing Bayes is hard because taking integrals is hard. We have to take integrals in Bayes equation
- Need to do numerical approximation
- We want to go where the integrand is big and focus our energy/computation there
- Problem is that, the higher in dimension we go, the most highest volume of our parameter space gets smaller and smaller relative to the rest of the space
- THe product of density and volume (which is probability mass) $\pi(q) dq$  is what contributes to an integral
- A _Typical Set_ is the center part of the probability mass around the mode
- One solution is Monte Carlo, which is drawing samples from the mass 
- Approximate integrating over prob mass by integrating over the typical set, and approx integrating over the typical set by integrating over the monte carlo samples of the typical set. Average over the MC samples
- Monte Carlo estimators average a given function over these samples to approximate the expectation
  - $\frac{1}{N} \sum_{n = 1}^N f(q_n) \sim N \left( E[f], \frac{Var[f]}{N}\right)$
- MC standard error is independent of the dimension of the data because we can quantify our MC error with the expression above
  - Problem is that drawing samples is hard. We haven't found the typical set yet
- Markov Chain allows us to get correlated samples
  - We can get transition density which preserves target distribution
    - $\pi(q) = \int dq'\pi(q') T(q|q')$, where $q'$ is the sample you just took and $T$ is transition?
  - Our transition density are going to concentrate towards the areas of high probability mass 
- If run long enough, our MCMC chains consistently estimate MC samples. Problem is that this is only true asymptotically. If you do it infinitely long, you will get the true expectation.
- How do you know how many initial samples to throw out? And how do you know if it's run long enough to get close enough to the true expectation?
- If we have geometries that are complex and small, our MCMC estimators will never explore that region, which leads to biased estimators
- Random Walk Metropolis explores posterior with "guided" diffusion
  - Add some random noise and have a proposed random step. Then accept or reject that point based on probability of density
  - $T(q|q') = N(q|q', \sigma^2) min \left(1, \frac{\pi(q)}{\pi(q')} \right)$
  - In high dimensional spaces, almost all of the potential locations you could jump to will end up being rejected because they fall outside the typical set. Very slow algorithm
  - This is an inefficient guess-and-check procedure

#### HMC:
- If we can create a vector field that will tell us where to go to explore the entire set, then we can just follow the arrows
- Each parameter $q$ has an associated momentum variable $p$.
- Construct prob distribution for momentum variables as well. 
  - $\pi(q,p) = \pi(q|p) \pi(p)$. This structure allows us at any time to through out momentum and recover just $\pi(p)$, which is what we are interested in.
- The definition of the Hamiltonian: 
\begin{align*}
  H(q,p) &= -\log \pi (p|q) \pi(q) \\
   &= -\log \pi (p|q) - log \pi(q) \\
   &= - K - V,\\
\end{align*}
where $K$ is potential energy and $V$ is kinetic energy.
- We now have a vector field for kinetic and potential. (meaning we have derivatives for each that point in a certain direction)
- (Can include some images of HMC visually from lecture / paper)
- HMC is very robust, hard to break
  - When it does break, it's super obvious
- "STAN is a platform for Bayesian inference that tries to abstract as much computation away as possible, so users can focus on building complex models"

### 2020-04-08 survival-models/cox-exponential-jags.R

@@ -47,7 +47,10 @@ start_censored <- num_uncensored + 1
 t <- t[order(is_censored)]
 x <- x[order(is_censored)]
 t_censor <- t
+t_censor[1:num_uncensored] <- 0
 t[start_censored:N] <- NA
+censored <- rep(0, length(t))
+censored[start_censored:N] <- 1
 
 jags_data <- list(
   "t" = t,
@@ -55,7 +58,8 @@ jags_data <- list(
   "x" = x,
   "num_uncensored" = num_uncensored,
   "start_censored" = num_uncensored + 1,
-  "N" = length(t)
+  "N" = length(t),
+  "censored" = censored
 )
 
 
@@ -83,6 +87,20 @@ jags_model <- "
     for (i in start_censored:N) {
       lambda[i] <- exp(beta * x[i])
       t[i] ~ dexp(lambda[i]) T(t_censor[i],)
+    }
+    beta ~ dnorm(0,0.0001)
+  }
+"
+jags_model <- "
+  model{
+    for (i in 1:num_uncensored) {
+      lambda[i] <- exp(beta * x[i])
+      t[i] ~ dexp(lambda[i])
+    }
+    for (i in start_censored:N) {
+      censored[i] ~ dinterval(t[i], t_censor[i])
+      lambda[i] <- exp(beta * x[i])
+      t[i] ~ dexp(lambda[i]) 
     }
     beta ~ dnorm(0,0.0001)
   }

### 2020-04-08 survival-models/cox-exponential-stan.R

@@ -78,7 +78,7 @@ source(here("currently-benchmarking.R"))
 
 if (!currently_benchmarking()) {
   stan_fit <- stan(
-    "file" = stan_file, "data" = stan_data, "init" = stan_inits,
+    "file" = stan_file, "data" = stan_data,# "init" = stan_inits,
     "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup, 
     "control" = list("adapt_delta" = 0.99)
   )

### 2020-04-08 Paper.Rmd
paper edits

@@
 This tool can be very useful in trying to explain how all the components of a model work together.
+
+# Example Model
+
+To illustrate how each bayesian computational method is used in practice, we will demonstrate how to construct a sampler for a simple Beta-Binomial model for each of the five methods we are studying.
+
+## Data Preparation
+
+First, we need to prepare the data that we will be using.
+
+```{r}
+p <- .25 # binomial p
+n <-  10 # binomial n
+
+set.seed(1)
+
+(y <- rbinom(1, n, p))
+
+model_data <- list(
+  "y" = y,
+  "N" = n
+)
+```
+
+## Model Configuration
+
+Next, we will pick some basic model parameters. Here, we will use four chains, a warmup of 1,000 iterations, and a total iteration count of 10,000.
+
+```{r}
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+```
+
+## OpenBUGS / R2OpenBUGS
+
+When using OpenBUGS through R, we will use the `R2OpenBUGS` library in order to pipe the data and model code into the program rather than using manual entry.
+This requires the creation of a temporary file that holds the model text that will be used.
+
+```{r}
+## load library
+library(R2OpenBUGS)
+
+## model script code
+bugs_model <- function() {
+  y ~ dbin(p,N)
+  p ~ dbeta(1,1)
+}
+
+## parameters to monitor
+bugs_monitor <- "p"
+
+## temporary model script text file creation
+bugs.file <- file.path(tempdir(), "model.txt")
+write.model(bugs_model, bugs.file)
+
+## run the model
+bugs_fit <- bugs(
+    "model.file" = bugs.file, "data" = model_data, "parameters.to.save" = bugs_monitor, 
+    "inits" = NULL, "n.chains" = n_chains, "n.iter" = n_iter, "n.burnin" = n_warmup
+)
+```
+
+## JAGS
+
+In JAGS, the model script simply needs to be inputted as a string to the `run.jags()` function.
+```{r}
+## load libraries
+library("rjags"); library("runjags")
+
+## model script code
+jags_model <- "
+  model{
+    y ~ dbin(p,n)
+    p ~ dbeta(1,1)
+  }
+"
+
+## parameters to monitor
+jags_monitor <- c("p")
+
+## run the model
+jags_fit <- run.jags(
+    "model" = jags_model, "data" = model_data, "monitor" = jags_monitor, 
+    "n.chains" = n_chains, "sample" = n_iter, "burnin" = n_warmup
+) 
+```
+
+## Nimble
+
+An additional step needed when using Nimble, which is to specify initial values for all parameters of interest, since the sampler cannot generate initial values on its own.
+```{r}
+## load library
+library(nimble)
+
+## model script code
+nimble_model <- nimbleCode({
+  y ~ dbin(p,n)
+  p ~ dbeta(1,1)
+})
+
+## parameters to monitor
+nimble_monitor <- c("p")
+
+## specify initial values of parameters
+nimble_inits <- list(
+  "p" = rbeta(1,1,1)
+)
+
+## run the model
+nimble_fit <- nimbleMCMC(
+    "code" = nimble_model, "data" = model_data, 
+    "inits" = nimble_inits, "monitors" = nimble_monitor, "nchains" = n_chains, 
+    "niter" = n_iter, "nburnin" = n_warmup, "summary" = TRUE
+)
+```
+
+## STAN
+
+Running models in STAN requires a .stan file which contains the model script code, written in the unique STAN syntax.
+This can either be a seperate file, or can be inputted as a character string.
+It is generally recommended to use a seperate .stan file, espcially when using RStudio, which can do syntax highlighting and point out errors in STAN code.
+
+Here is the .stan file for this example, which is saved as "stan_file.stan". Each STAN file contains a data block, parameters block, and model block at a minimum.
+```{r}
+data {
+  int<lower=0> y;
+  int<lower=0> n;
+}
+
+parameters {
+  real p;                
+}
+
+model {
+  y ~ binomial(n,p);
+  p ~ beta(1,1);
+}
+```
+
+The sampler can then be run simply using the `stan()` function call, with the .stan file location specified.
+
+```{r}
+## load library
+library("rstan")
+
+## run the model
+stan_fit <- stan(
+    "file" = "stan_file.stan", "data" = model_data, 
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup, 
+    "control" = list("adapt_delta" = 0.99)
+)
+```
+
+## Greta
+
+Each data variable needs to be converted to a greta array before it can be used in the model.
+```{r}
+## convert data into greta arrays
+y <- greta::as_data(y)
+n <- greta::as_data(n)
+```
+
+Next, the priors for each parameter of interested are defined using functions that describe the prior distribution paramter values.
+The distribution of the response variable given in the data is then described using the `distribution()` function.
+The model is compiled using the `model()` function, before being sampled by `mcmc()`.
+
+```{r}
+## load library
+library(greta)
+
+## specify prior for parameter(s) of interest
+p <- beta(1,1)
+
+## define distribution of data variable and relationship with parameter
+distribution(y) <- binomial(n,p)
+
+## compile model and run
+greta_model <- model(p)
+greta_fit <- mcmc(
+    "model" = greta_model, "n_samples" = n_iter,
+    "warmup" = n_warmup,"chains" = n_chains
+)
+```
 
@@
 
+## STAN Probabilty and Sampling Statements
+
+Inside the model block of a stan file, probability statements can be written in the same convention as in other bayesian sampling methods
+
+```{r}
+model{
+  y ~ binomial(n,p);
+  p ~ beta(1,1);
+}
+```
+
+This kind of probability statement does not actually perform any sampling, but is instead tranformed into a statement about the incremental log probability density of a target distribution.
+The log probability density function can be specified directly.
+
+```{r}
+model {
+  target += binomial_lpdf(y | n, p);
+  target += beta_lpdf(p | 1, 1);
+}
+```
+
+Either method yields the same result, but in some cases it is advantageous to use the incremental log probability density (or cumulative density) function to get exact log probability values for the model.
+In general, using the standard sampling statement syntax is preferred since it is easier to understand.
+
 ## Greta's Unique Specification Language

### 2020-04-09 Paper.Rmd
added paragraph on HMC

@@
 
-### HMC
+### Hamiltonian Monte Carlo
 
-Radford-Neal HMC Chapter:
-- Hamiltonian Monte Carlo uses Hamiltonian Dynamics
-  - Visualize the dynamics as a frictionless puck that slides over a surface of varying height
-    - The state of the system has three elements: position, momentum, and potential energy
-  - The position corresponds to variables of interest
-  - The potential energy will be minus the log of the probability density for these variables
-  - There is one momentum variable for each position variable
-- Using "Hamilton's equations", the movement of the puck is computed. 
-- For computational implementation, these equations must be approximated by discretizing time, using some small stepsize.
-- To use Hamiltonian dynamics to sample from a distribution using MCMC, we need to translate the density function for the distribution to a potential energy function.
-The variables of interest are now position variables
-- We simulate a markov chain where each iteration resamples the momentum variable associated with each position variable and then does a Metroplis update with the proposal found using Hamiltonian dynamics
-  - Two steps for each iteration:
-    1. New values for the momentum variables are randomly drawn from their Gaussian distribution, independent of the current values of the position variables.
+A main of objective of Hamiltonian Monte Carlo is to use the geometry of the typical set to help us know where to sample from, instead of blindly using a guess-and-check method (Betancourt HMC Youtube talk).
+This method utilizes Hamiltonian Dynamics, which is a physics system that has three elements: position, momentum, and potential energy variables.
+To use Hamiltonian Dynamics to traverse the typical set we are interested in, each parameters of interest is transformed into a position variable, which can be visualized as a frictionless puck sliding over a surface of varying height.
+Each position variable $p$ has an associated momentum variable $q$ that defines the direction and magnitude of the movement of the "puck".
+At any given time, the value of the parameter can be found as the position of $p$ in some dimension.
+
+We define the joint probability of the position $p$ and momentum $q$ as $\pi(q,p) = \pi(q|p) \pi(p)$.
+This structure allows us at anytime to throw out $q$ and recover the value of $p$, which iholds the position values of the parameter(s).
+The Hamiltonian is defined as
+
+\begin{align*}
+  H(q,p) &= -\log \pi (p|q) \pi(q) \\
+   &= -\log \pi (p|q) - log \pi(q) \\
+   &= - K - V,\\
+\end{align*}
+where $K$ is potential energy and $V$ is kinetic energy. Taking partial derivatives of the Hamiltonian yields Hamilton's equations, which is where we obtain our vector field.
+
+\begin{align*}
+  \frac{dq}{dt} &= \frac{\partial K}{\partial p} \\ 
+  \frac{dp}{dt} &= - \frac{\partial K}{\partial q} - \frac{\partial V}{\partial q}.
+\end{align*}
+
+Hamiltonian Dynamics can be used to sample from the target distribution by translating the density function of the distribution to a potential energy function, and obtaining position variables (and associated momentum variables) for parameters of interest.
+We can simulate a markov chain where each iteration consists of the following steps:
+
+1. New values for the momentum variables are randomly drawn from their Gaussian distribution, independent of the current values of the position variables
 2.
-A metroplis update is performed, using hamiltonian dynamics to propose a new state.
-The proposed state is either accepted or rejected based on the HMC probability function used.
+A metropolis update is performed, using hamiltonian dynamics to propose a new state for the position variables.
+The proposed state is either accepted or rejected based on the Hamiltonian Monte Carlo probability function used.
 If the proposed state is rejected, the next state is the same as the current one.
-- The HMC algorithm typically will not be trapped in some subset of the state space, which allows for it to always asymptotically converge to the target distribution, and quickly.
-- Tuning Parameters for HMC:
-  - Stepsize - the leapfrog stepsize. Too large, and it will lead to a low acceptance rate for proposed states. Too small and it will be slow computationally.
-  - Trajectory length -
+
+Hamiltonian Monte Carlo methods are very robust, in that the markov chains seldom get caught in complex regions of the posterior compared to other common MCMC methods.
+When the chains do get caught, the HMC chains break so obviously that errors can be easily spotted and reported to the user.
 
@@
 - Works in command line, Python, and R
+- "STAN is a platform for Bayesian inference that tries to abstract as much computation away as possible, so users can focus on building complex models"
 

### 2020-04-09 Radford Neal HMC notes.Rmd
added paragraph on HMC

---
title: "Radford Neal HMC notes"
author: "Evan Miyakawa"
date: "4/7/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Hamiltonian Monte Carlo uses Hamiltonian Dynamics
  - Visualize the dynamics as a frictionless puck that slides over a surface of varying height
    - The state of the system has three elements: position, momentum, and potential energy
  - The position corresponds to variables of interest
  - The potential energy will be minus the log of the probability density for these variables
  - There is one momentum variable for each position variable
- Using "Hamilton's equations", the movement of the puck is computed. 
- For computational implementation, these equations must be approximated by discretizing time, using some small stepsize.
- To use Hamiltonian dynamics to sample from a distribution using MCMC, we need to translate the density function for the distribution to a potential energy function. The variables of interest are now position variables
- We simulate a markov chain where each iteration resamples the momentum variable associated with each position variable and then does a Metroplis update with the proposal found using Hamiltonian dynamics
  - Two steps for each iteration:
    1. New values for the momentum variables are randomly drawn from their Gaussian distribution, independent of the current values of the position variables.
    2. A metropolis update is performed, using hamiltonian dynamics to propose a new state. The proposed state is either accepted or rejected based on the HMC probability function used. If the proposed state is rejected, the next state is the same as the current one.
- The HMC algorithm typically will not be trapped in some subset of the state space, which allows for it to always asymptotically converge to the target distribution, and quickly.
- Tuning Parameters for HMC:
  - Stepsize - the leapfrog stepsize. Too large, and it will lead to a low acceptance rate for proposed states. Too small and it will be slow computationally.
  - Trajectory length -

### 2020-04-15 Paper.Rmd
paper edits

@@
 
-### History up through BUGS (include slice sampling)
+### Metropolis-Hastings
 
-- BUGS project began in 1989
-- BUGS language contains for a somewhat concise list of functions and distributions used for specifying relationships between 
-- The BUGS language only has a somewhat limited syntax, but is able to describe extremly complex models using this syntax.
-- While the programs that use BUGS-type language are continually evolving, the syntax has stayed very stable.
-- BUGS language is declarative, which means the order of statements doesn't matter, as opposed to a procedural language.
-- A main initial purpose for the creation of bugs was to allow for graphical models -- models that can be specified through click and point software that creates nodes, or expressing graphics of models that already exist for easier understanding.
-- Six steps in running BUGS models
-  1. Checking syntax of model specification
-  2. Loading data
-  3. Compiling the BUGS model
-  4. Starting the simulation at appropriate initial values
-  5. Simulate unknown parameters
-  6. Gather results and report summary statistics, along with other appropriate output
-- WinBUGS was the first widely used program to utilize the BUGS language.
-However, OpenBUGS is the more recently maintained version, which communicates better with other software and is more portable.
-- R2OpenBUGS is the R package that allows for R to interface with OpenBUGS
--
+### Gibbs Sampling
 
-### Gibbs Sampler
+Gibbs Sampling is one of the most common Markov Chain Monte Carlo Algorithms used for Bayesian computation.
+A special case of the Metropolis-Hastings algorithm, a Gibbs sampler works by generating a multi-dimensional markov chain, with one dimension for each unknown variable of interest.
+By splitting the vector of random unknown variables, the algorithm samples each subvector one by one, conditional on the most recent values of the other variables.
 
-- One of most common algorithms for simulating MCMCs.
-- Special case of Metropolis-Hastings alg
-- Generates a multi-dimensional markov chain by splitting the vector of random unknown variables and sampling each subvector one by one, conditional on the most recent values of the other variables.
-- Algorithm: If we have a vector of unknown parameters $\theta$, where $\theta = (\theta_1, \theta_2, \dots, \theta_k)$,
+Lunn et. all outlines the Gibbs sampler algorithm as follows: If we have a vector of unknown parameters $\theta$, where $\theta = (\theta_1, \theta_2, \dots, \theta_k)$, then
+
 1. Choose arbitrary starting values $\theta_1^{(0)}, \theta_2^{(0)}, \dots, \theta_k^{(0)}$ 
@@
 
-- The Gibbs sampling strategy often allows for sampling from high-dimensional spaces to be simplified to drawing from one-dimensional distributions.
-- Weakness: because it uses a random-walk behavior, if two variables are highly correlated, traversing the space can take a very long time.
+The beauty of the Gibbs sampling strategy is that, no matter how complex the space of the posterior distribution is, sampling from these high-dimensional spaces can be broken down into more straightforward draws from one-dimensional distributions.
+The downside of this approach is that, in high-dimensional settings, it may take a very long time to sample from each conditional one-dimensional distribution in order to reach convergence.
+Furthermore, because the Gibbs sampler uses a random-walk behavior, if several variables are highly correlated, traversing the space can take a very high amount of iterations because the step-size needs to be small in order to not step outside the target distribution.
 
-### Metropolis
+### BUGS
+
+The BUGS project, which stands for Bayesian inference Using Gibbs Sampling, began in 1989 in an effort to create a program that utilized MCMC to approximate draws from the posterior distribution in a Bayesian framework.
+The BUGS language has functions and distributions that allow for specifying relationship between different "nodes" in a bayesian model.
+While the language itself has a somewhat limited syntax, it is able to describe a very diverse and complex set of models.
+The BUGS language is declarative, meaning that the order of statements in a model script does not matter since the entire script is evaluated at once, instead of line-by-line in a procedural language.
+
+Many different programs have been developed that utilize the BUGS syntax to accomplish bayesian sampling, and while these programs have continually evolved, the BUGS syntax has remained very consistent.
+The Windows program WinBUGS was created by the inventors of the BUGS language, and was a more popular choice for bayesian computation until it became deprecated in 2007.
+OpenBUGS is now more commonly used in place of WinBUGS, although it's last stable release was in 2014.
+OpenBUGS communicates better with other software than WinBUGS and is more portable.
+
+In general, there are six steps needed to run bayesian models in WinBUGS/OpenBUGS:
+
+1. Specify the model through BUGS syntax and check to make sure the model code is valid.
+2. Load the appropriate data.
+3. Compiling the BUGS model.
+4. Start the simulation at appropriate initial values for the Markov chains.
+5. Simulate draws from the posterior distributions of the unknown parameters of interest.
+6. Gather the results and report summary statistics, along with other appropriate output.
 

### 2020-04-27 PPP/Figures.Rmd
ppp update and adding accuracy bench

--- 2020-04-08 Paper.Rmd
@@
 ---
-title: "Bayes Computational Methods"
+title: "Figures"
 author: "Evan Miyakawa"
 output: pdf_document
+
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_chunk$set(eval = FALSE)
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+
 ```
 
-# Introduction
-
-## The Bayes Paradigm
-
-In Bayesian statistical inference, we are interested in finding the posterior distribution of a certain parameter, given a prior distribution and a likelihood for that parameter from a sample of data.
-Bayes Rule states that, given a parameter of interest $\theta$ and some known data $x$, we can find the conditional probability of $\theta$ given $y$ using $$ p(\theta | x) = \frac{p(x | \theta) p(\theta)}{p(x)},$$ where in the discrete case $p(x) = \sum_\theta p(\theta) p(x|\theta),$ and in the continuous case, $p(x) = \int p(\theta) p(y|\theta) d\theta$.
-More formally, we can derive the posterior distribution of a parameter $\theta$, $p(\theta | x)$, by $$p(\theta | x_1, x_2, \dots, x_n) = \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}},$$ where $\ell(x | \theta)$ is proportional to the likelihood of $\theta$, conditioned on the data, and $\pi(\theta)$ is the prior distribution of $\theta$.
-For example, suppose that $X_i \sim \text{Bern}(\theta)$, and we are interested in inference on $\theta$.
-A reasonable prior to assign to $\theta$ would be a $\text{Beta}(\alpha, \beta)$, distribution, given that $\theta$ is constrained to be between 0 and 1, which is a property of the Beta distribution.
-Therefore, given that $\theta \sim \text{Beta}(\alpha, \beta)$ for some known values of $\alpha$ and $\beta$, then
-\begin{align*}
-  p(\theta | x) &= \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}} \\
-  &= \frac{\theta ^ {\sum_{i=1}^{n} x_i} (1 - \theta) ^ {n - \sum_{i=1}^{n} x_i} \frac{\Gamma (\alpha + \beta)}{\Gamma(\alpha) + \Gamma(\beta)} \theta^{\alpha - 1} (1 - \theta)^{\beta - 1}}{\int \theta ^ {\sum_{i=1}^{n} x_i} (1 - \theta) ^ {n - \sum_{i=1}^{n} x_i} \frac{\Gamma (\alpha + \beta)}{\Gamma(\alpha) + \Gamma(\beta)} \theta^{\alpha - 1} (1 - \theta)^{\beta - 1} d \theta} \\
-  &\propto \theta ^ {\sum (x_i) + \alpha - 1} (1 - \theta) ^ {n - \sum (x_i) + \beta - 1},
-\end{align*}
-which is the kernel of a $\text{Beta}\left(\sum (x_i) + \alpha, n - \sum (x_i) + \beta \right)$ distribution.
-Therefore, we can analytically calculate the posterior distribution of $\theta$ as $$p(\theta | x) \sim \text{Beta}\left(\sum (x_i) + \alpha, n - \sum (x_i) + \beta\right).$$
-
-## Bayesian Computation
-
-In more complicated scenarios, it is not easy to calculate the posterior distribution of a parameter by hand, and in many cases it is impossible.
-This is where computing can help us.
-If we could somehow computationally grab samples from the posterior distribution we are interested in, then we could obtain most of the statistics that we are interested in.
-
-### History of Bayesian computation up to MCMC
-
-### Markov Chain Monte Carlo
-
-Markov Chain Monte Carlo (referred to as MCMC), is a common approach that allows for drawing samples of a parameter from a distribution that is approximately equivalent to the target posterior distribution of that parameter.
-It is an iterative algorithm, where each realization gets closer to the target density we are interested in.
-An essential property of any Markov chain is that the distribution of each draw only depends on the most recent value drawn before it.
-As we keep obtaining draws, each draw is probabilisticly more likely to have come from the target posterior distribution.
-Based on the convergence properties of MCMC, we can expect each draw to be coming from the real distribution of the parameter we are interested, once we have let the Markov chain run long enough.
-
-- MCMC is the only suitable method when we don't have a conjugate prior.
-
-### History up through BUGS (include slice sampling)
-
-- BUGS project began in 1989
-- BUGS language contains for a somewhat concise list of functions and distributions used for specifying relationships between 
-- The BUGS language only has a somewhat limited syntax, but is able to describe extremly complex models using this syntax.
-- While the programs that use BUGS-type language are continually evolving, the syntax has stayed very stable.
-- BUGS language is declarative, which means the order of statements doesn't matter, as opposed to a procedural language.
-- A main initial purpose for the creation of bugs was to allow for graphical models -- models that can be specified through click and point software that creates nodes, or expressing graphics of models that already exist for easier understanding.
-- Six steps in running BUGS models
-  1. Checking syntax of model specification
-  2. Loading data
-  3. Compiling the BUGS model
-  4. Starting the simulation at appropriate initial values
-  5. Simulate unknown parameters
-  6. Gather results and report summary statistics, along with other appropriate output
-- WinBUGS was the first widely used program to utilize the BUGS language.
-However, OpenBUGS is the more recently maintained version, which communicates better with other software and is more portable.
-- R2OpenBUGS is the R package that allows for R to interface with OpenBUGS
--
-
-### Gibbs Sampler
-
-- One of most common algorithms for simulating MCMCs.
-- Special case of Metropolis-Hastings alg
-- Generates a multi-dimensional markov chain by splitting the vector of random unknown variables and sampling each subvector one by one, conditional on the most recent values of the other variables.
-- Algorithm: If we have a vector of unknown parameters $\theta$, where $\theta = (\theta_1, \theta_2, \dots, \theta_k)$,
-  1. Choose arbitrary starting values $\theta_1^{(0)}, \theta_2^{(0)}, \dots, \theta_k^{(0)}$ 
-  2. Complete one iteration of the sampler by getting new values for $\theta$ using the following:
-    - Sample a new value for $\theta_1$, from the full conditional distribution of $\theta_1$, given the most recent values of all other elements of $\theta$ and the data:
-    $$\theta_1^{(1)} \sim p(\theta_1 | \theta_2^{(0)}, \theta_3^{(0)}, \dots , \theta_k^{(0)}, y).$$
-    - Sample a new value $\theta_2^{(1)}$ for the second component of $\theta$, from the full conditional distribution $p(\theta_2 | \theta_1^{(1)}, \theta_3^{(0)}, \dots , \theta_k^{(0)}, y)$.
-    - Continue this process for the remainder of the $k$ elements of $\theta$.
-  3. Repeat step 2 for the total decided number of iterations that the sampler will run.
-
-- The Gibbs sampling strategy often allows for sampling from high-dimensional spaces to be simplified to drawing from one-dimensional distributions.
-- Weakness: because it uses a random-walk behavior, if two variables are highly correlated, traversing the space can take a very long time.
-
-### Metropolis
-
-### HMC
-
-Radford-Neal HMC Chapter:
-- Hamiltonian Monte Carlo uses Hamiltonian Dynamics
-  - Visualize the dynamics as a frictionless puck that slides over a surface of varying height
-    - The state of the system has three elements: position, momentum, and potential energy
-  - The position corresponds to variables of interest
-  - The potential energy will be minus the log of the probability density for these variables
-  - There is one momentum variable for each position variable
-- Using "Hamilton's equations", the movement of the puck is computed. 
-- For computational implementation, these equations must be approximated by discretizing time, using some small stepsize.
-- To use Hamiltonian dynamics to sample from a distribution using MCMC, we need to translate the density function for the distribution to a potential energy function.
-The variables of interest are now position variables
-- We simulate a markov chain where each iteration resamples the momentum variable associated with each position variable and then does a Metroplis update with the proposal found using Hamiltonian dynamics
-  - Two steps for each iteration:
-    1. New values for the momentum variables are randomly drawn from their Gaussian distribution, independent of the current values of the position variables.
-    2.
-A metroplis update is performed, using hamiltonian dynamics to propose a new state.
-The proposed state is either accepted or rejected based on the HMC probability function used.
-If the proposed state is rejected, the next state is the same as the current one.
-- The HMC algorithm typically will not be trapped in some subset of the state space, which allows for it to always asymptotically converge to the target distribution, and quickly.
-- Tuning Parameters for HMC:
-  - Stepsize - the leapfrog stepsize. Too large, and it will lead to a low acceptance rate for proposed states. Too small and it will be slow computationally.
-  - Trajectory length -
-
-### STAN
-STAN 2015
-- Based on a probabilistic programming language that is more flexible and expressive than BUGS/JAGS.
-- Based on HMC
-- Works in command line, Python, and R
-
-# Bayesian Samplers
-
-## OpenBUGS
-
-OpenBugs is the oldest of the Bayesian samplers that we use.
-OpenBugs is designed for Windows machines, but there is a workaround to run it on MacOS and Linux, using WINE, a program designed to run Windows software on other operating systems.
-OpenBugs can be used through the actual click-and-point application, or through \texttt{R}, using the \texttt{R2OpenBugs} package, which allows for the user to send model scripts and data to OpenBugs without needing to leave the \texttt{R} interface.
-The \texttt{R} command to run OpenBugs is \texttt{bugs()}.
-Using this method causes OpenBugs to run invisibly in the background and return the results, unless an argument is specified to debug code in the actual application itself, which can be very useful.
-When using R to run OpenBugs, there is little way to know what went wrong if the code errors somewhere without interacting with the OpenBugs program itself.
-Using the \texttt{debug = T} argument within \texttt{bugs()} will cause the application to run in the forefront as if it were actually being used directly by the user.
-This way, any messages given by the application while running can be viewed directly.
-It is recommend to use \texttt{debug = T} only when debugging, as it does slow down the sampling process slightly.
-OpenBugs has many tools to visualize the performance the sampler which are easy to use.
-The biggest drawback of using the OpenBugs application directly is that specifiying the data can be tedious if the data is not very small in size or design.
-Using R to specify the data is much easier.
-One unique feature in OpenBugs is the Doodle tool, which allows for model specification through the creation of a diagram of nodes representing the parameters.
-Though using Doodle can be helpful for illustration, it is not used often in practice.
-
-## JAGS
-
-In terms of model specification, JAGS is very similar to OpenBugs, but works more seemlessly in R using the package \texttt{rjags}.
-JAGS was developed independently of OpenBUGS, but uses the BUGS language for most of its model specification syntax.
-The \texttt{runjags} is an optional package that allows for the JAGS model script to be inputted as a string, and run using the \texttt{run.jags()} function.
-Since JAGS runs inside \texttt{R} itself, it is very easy to work with.
-Helpful errors are given when there is an issue with the model code or sampling process.
-JAGS is often a very fast sampler, especially for less complicated models.
-
-## Nimble
-
-Nimble is an extension of Bugs that allows for bayesian computation within \texttt{R}, but uses C++ under the hood to compile models and algorithms for speed.
-Using the \texttt{nimble} package, Nimble can perform MCMC bayesian computations, but also allows for other types of algorithms to be loaded or custom-built.
-It can also compile other code without BUGS models into C++ for other uses.
-For the purpose of bayesian sampling using MCMC, Nimble is used very similarly to BUGS in practicality.
-The flexibility of Nimble is not needed in most cases when doing MCMC.
-The model specification language is almost identical to OpenBugs.
-A model can be both compiled and run using the function \texttt{nimbleMCMC()}.
-Because the model is compiled using C++ each time, this can drastically increase the total computation time as compared to other sampling methods.
-Occasionally, Nimble can also freeze up or fail, which can take a bit of time to fix.
-
-## STAN
-
-STAN is a language in-and-of itself used for statistcal modeling and data analysis.
-STAN can be used in several programming languages, including \texttt{R} and Python.
-While many bayesian samplers use Gibbs sampling, STAN uses Hamiltonian Monte Carlo under the hood, a more computationally efficient sampler that is based on an advanced physics simulation.
-This can greatly reduce the computational time for many types of statistical models, especially for more complicated ones.
-Like Nimble, STAN compiles each model, but one big advantage is that it can store each compiled model in a .Rds file, which allows for repeated use without the need to re-compile each time.
-In \texttt{R}, STAN is used through the \texttt{rstan} package, which allows for data to be loaded through normal \texttt{R} data structures.
-The biggest practical difference in using STAN compared to other bayesian computational methods is in the model specification process.
-STAN requires a model object in the form of a .stan file, which is created in a very unique language that is more formal than in BUGS.
-A .stan file, either created through a character string in R or in a seperate text file, can be used by the \texttt{stan()} function, which takes the file along with the data to perform the sampling.
-There is a very robust set of manuals for STAN users, compared to other methods.
-
-## Greta
-
-Greta, a very recent addition to the list of bayes computational methods, is made specifically for use in \texttt{R}, making the model and data specification process as easy as possible.
-It also uses Google TensorFlow to perform its calculations, and allows for more flexible use of CPUs and GPUs.
-Unfortunately, as of the writing of this, it is still in need of lots of development, and lacks an appropriate amount of documentation and support for users.
-Computationally, it is very slow compared to the other bayes methods, especially in simple cases.
-There are several types of semi-common models that greta seems unable to properly handle in model specfication at the moment.
-Greta also attempts to make model specification very user-friendly, but uses some very common function names in the process that might override other often-used \texttt{R} functions.
-Greta does offer a model visualization tool that does not exist in any of the other Bayesian samplers, in that it can provide a graphical representation of the specified model through the use of nodes and arrows.
-This tool can be very useful in trying to explain how all the components of a model work together.
-
-# Example Model
-
-To illustrate how each bayesian computational method is used in practice, we will demonstrate how to construct a sampler for a simple Beta-Binomial model for each of the five methods we are studying.
-
-## Data Preparation
-
-First, we need to prepare the data that we will be using.
+```{r}
+library(tidyverse); options(tibble.width = Inf)
+library(magrittr)
+library("greta")
+```
 
 ```{r}
+set.seed(2)
+
+df <- tibble(x = rnorm(750, mean = 50, sd = 25), y = rnorm(750, mean = x * 1.2, sd = 5))
+
+df %>% ggplot(aes(x,y)) +
+  geom_point(alpha = 0.7) + 
+  theme_minimal() + 
+  theme(
+    axis.title = element_blank(),
+    axis.text = element_blank()
+  )
+```
+
+```{r}
+set.seed(2)
+
+df <- tibble(x = rnorm(750, mean = 50, sd = 25), y = rnorm(750, mean = x * 1.2, sd = 5))
+
+df %>% ggplot(aes(x,y)) +
+  geom_point(alpha = 0.3) + 
+  theme_minimal() + 
+  theme(
+    axis.title = element_blank(),
+    axis.text = element_blank()
+  )
+```
+
+```{r, fig.height = 3}
+
+log_df <- tibble(x = seq(-4,4,length.out = 100), d_norm = - dnorm(x))
+
+log_df %>% ggplot(aes(x,d_norm)) + 
+  geom_line(size = 2) + 
+  theme_minimal() + 
+  theme(
+    axis.title = element_blank(),
+    axis.text = element_blank()
+  )
+
+```
+
+```{r}
+
+## generate/specify data
+################################################################################
+
+alpha <- 2 #beta distribution alpha
+beta <- 1 #beta distribution beta
+N <- 10 # sample size
+
+set.seed(1)
+
+n <- rbinom(N, 20, 0.5)
+theta <- rbeta(N, alpha, beta)
+y <- rbinom(N, n, theta)
+
+n <- as_data(n)
+y <- as_data(y)
+
+## specify greta model
+################################################################################
+
+theta <- beta(2,1, dim = c(N,1))
+distribution(y) <- binomial(n, theta, dim = c(N,1))
+
+greta_model <- model(theta)
+
+plot(greta_model)
+
+```
+
+```{r, eval = FALSE}
 p <- .25 # binomial p 
@@
 
-## Model Configuration
-
-Next, we will pick some basic model parameters. Here, we will use four chains, a warmup of 1,000 iterations, and a total iteration count of 10,000.
-
-```{r}
+```{r, eval = FALSE}
 n_chains <- 4L
@@
 
-## OpenBUGS / R2OpenBUGS
-
-When using OpenBUGS through R, we will use the `R2OpenBUGS` library in order to pipe the data and model code into the program rather than using manual entry.
-This requires the creation of a temporary file that holds the model text that will be used.
-
-```{r}
+```{r, eval = FALSE}
 ## load library
@@
 
-## JAGS
-
-In JAGS, the model script simply needs to be inputted as a string to the `run.jags()` function.
-```{r}
+```{r, eval = FALSE}
 ## load libraries
@@
 
-## Nimble
-
-An additional step needed when using Nimble, which is to specify initial values for all parameters of interest, since the sampler cannot generate initial values on its own.
-```{r}
+```{r, eval = FALSE}
 ## load library
@@
 
-## STAN
+```{r, eval = FALSE}
+## .stan file
 
-Running models in STAN requires a .stan file which contains the model script code, written in the unique STAN syntax.
-This can either be a seperate file, or can be inputted as a character string.
-It is generally recommended to use a seperate .stan file, espcially when using RStudio, which can do syntax highlighting and point out errors in STAN code.
-
-Here is the .stan file for this example, which is saved as "stan_file.stan". Each STAN file contains a data block, parameters block, and model block at a minimum.
-```{r}
 data {
@@
 
-The sampler can then be run simply using the `stan()` function call, with the .stan file location specified.
-
-```{r}
+```{r, eval = FALSE}
 ## load library
@@
     "file" = "stan_file.stan", "data" = model_data, 
-    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup, 
-    "control" = list("adapt_delta" = 0.99)
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
 )
+
 ```
 
-## Greta
-
-Each data variable needs to be converted to a greta array before it can be used in the model.
-```{r}
+```{r, eval = FALSE}
 ## convert data into greta arrays
@@
 ```
-
-Next, the priors for each parameter of interested are defined using functions that describe the prior distribution paramter values.
-The distribution of the response variable given in the data is then described using the `distribution()` function.
-The model is compiled using the `model()` function, before being sampled by `mcmc()`.
 
@@
 greta_model <- model(p)
+
+```
+
+```{r, eval = FALSE}
 greta_fit <- mcmc(
@@
 )
-```
-
-# Common Bayesian Computation Settings
-
-### Chains
-
-When using Markov Chain Monte Carlo methods to estimate Bayesian posteriors of parameters, the number of independent Markov chains to be run needs to be specified.
-It is recommended to use several chains to explore the space of the target distribution, as a single chain may not fully traverse the entire region on its own.
-For some simpler models that don't have irregular features in the target distribution, using one chain can suffice.
-Each sampler has an argument in its function call that specifies the number of chains.
-Some of the samplers have a default value if the number of chains isn't given.
-
-### Iterations
-
-The number of steps that each Markov chain will take is a very important setting to get right.
-If each chain is not set up to run long enough, the parameter space may not be fully explored and the estimates will be inaccurate.
-Checking convergence statistics can be helpful in determining if the number of iterations is sufficient.
-An argument specifiying the number of iterations for each chain can be inputted in the function call for each sampler.
-
-### Warmup
-
-Since our inferences are based on the assmumption that our draws are from the target distribution, it is wise to discard the samples at the beginning of each chain, since it takes time for the sampler to "settle down".
-The terms "warmup" and "burn-in" are often used to indicate the number of iterations that we will discard at the beginning of each chain when calculating statistics about our posterior distribution.
-We can choose this number in the function call for each sampler.
-In \texttt{OpenBUGS}, \texttt{Nimble}, and \texttt{Jags}, the argument is referred to as burnin, but in \texttt{STAN} and \texttt{greta} it is called warmup.
-Some samplers (\texttt{greta}, \texttt{Jags, Nimble}) treat this warmup argument as a number of iterations to run before starting to track the sampling results, but others (\texttt{STAN}, \texttt{BUGS}) have this be the actual number of iterations to discard out of the total number of specified iterations, so it is important to know which designation a sampler uses.
-
-### Thinning
-
-In some cases, there is high correlation between sequential samples in a given chain, which can bias the resulting posterior inference, which can be diagnosed using some graphical procedures.
-In this situation, it is wise to "thin" the results by only looking at every ith observation from a chain, such as every 10th observation.
-While each sampler defaults to including every simulated posterior draw, a user can specify a "thinning rate", which is the rate at which a sampler will keep memory of the observations.
-A thinning rate of 1 means that every observation will be kept, and a rate of 10 indicates that only every 10th observation will be used.
-
-### Initial Value Specification
-
-In order to start sampling, each Markov chain needs to have an initial value at which to start.
-These initial values can be either be specified, or can often be randomly generated based on the prior distributions for each parameter of interest.
-In many cases, it is easiest to have the initial values randomly generated by the sampler.
-However, in cases where the target distribution is very complex, it is possible that poorly chosen initial values can lead to a chain getting "stuck" in a difficult region, or diverging in the wrong direction.
-In these cases, a more strategic choice of initial values is advised.
-\texttt{Nimble} unfortunately requires for these values to be specified instead of randomly generated.
-Using \texttt{R} to randomly generate the values from the prior distributions can be used in this situation:
-
-```{r}
-nimble_inits <- list(
-  "alpha" = rnorm(1,0,1000),
-  "beta" = rnorm(1,0,1000),
-  "sigma" = runif(1,0,10)
-)
-
-nimbleMCMC("inits" = nimble_inits, ... )
-```
-
-If multiple chains are being used, a list containing a number of lists equalling number of chains can be used:
-
-```{r}
-nimble_inits <- list(
-  list(
-    "alpha" = rnorm(1,0,1000),
-    "beta" = rnorm(1,0,1000),
-    "sigma" = runif(1,0,10)
-  ),
-    list(
-    "alpha" = rnorm(1,0,1000),
-    "beta" = rnorm(1,0,1000),
-    "sigma" = runif(1,0,10)
-  )
-)
-```
-
-Alternatively, if a list containing the initial values for one chain is used, these values will be recycled for the remaining chains.
-
-# Common Tricks for Model and Data Specification
-
-When describing a bayesian model through code, there are instances were slightly different implementations are needed based on the sampler being used.
-Both in model and data specification, the user needs to understand the intricacies of the specific sampling tool being used in order to get the analysis to run properly.
-This section provides some helpful tips for implementing models in each of the different samplers being examined.
-
-## Model Specification
-
-The most important distinction between each of the samplers is the way that each allows for model specification text to be inputted.
-
-##### OpenBUGS
-In \texttt{OpenBUGS}, the earliest method created of the lot, the model specification code must be wrapped inside the expression \texttt{model\{\}}:
-
-```{r}
-model{
-  ...model code here...
-}
-```
-
-Since we are using the \texttt{R2OpenBUGS} library to run \texttt{OpenBUGS} through an \texttt{R} session, a text file with the model code is passed to the \texttt{bugs()} function.
-We can specify the model in \texttt{R} and create the appropriate text file:
-
-```{r}
-bugs_model <- function() {
-  ...model code here...
-}
-
-bugs.file <- file.path(tempdir(), "model.txt")
-write.model(bugs_model, bugs.file)
-
-bugs("model" = bugs.file, ... )
-```
-
-##### JAGS
-The \texttt{JAGS} sampler was intended to allow for model specification code to be almost identical to \texttt{BUGS}, so the implementation is very similar.
-There are two common approaches, one using the \texttt{rjags} package, and one using the \texttt{runjags} package.
-We use the latter, which allows for the model code to be specified in a multi-line character string:
-
-```{r}
-jags_model <- "
-  model{
-  ...model code here...
-  }
-"
-
-run.jags("model" = jags_model, ... )
-```
-
-##### Nimble
-Model specification in \texttt{Nimble} is extremely similar to \texttt{BUGS} as well.
-Model code is wrapped inside the \texttt{nimble} package function \texttt{nimbleCode()}, and then fed into \texttt{nimbleMCMC()} sampler function:
-
-```{r}
-nimble_model <- nimbleCode({
-  ...model code here...
-})
-
-nimbleMCMC("code" = nimble_model, ... )
-```
-
-##### STAN
-The STAN model specification language is significantly different than other samplers, in that a .stan file is fed to the \texttt{C++} compiler.
-A .stan file has a very unique model specification format, which divides code into "data", "parameter", and "model" blocks at a minimum.
-When needed, "transformed data" and "transformed parameters" blocks are also used.
-The basic format for model specification is provided below:
-
-```{r}
-data {
-  ...data specified here...        
-}
-
-parameters {
-  ...parameters specified here...                
-}
-
-model {
-  ...model code here...
-}
-```
-
-Each data variable and and parameter variable needs to be initialized in the appropriate block, along with associated domain limits.
-For example, if one were initializing a rate parameter for an exponential distribution called "lambda", it would be initialized with a lower bound of 0:
-
-```{r}
-parameters {
-  <lower=0> lambda;               
-}
-```
-
-A character string specifying the path to the .stan file is then passed to \texttt{stan()}.
-
-```{r}
-stan_file <- "file.stan"
-stan("file" = stan_file, ... )
-```
-
-## Data Specification and Input
-
-Specifying data for the bayesian samplers in \texttt{R} is very straightforward.
-In \texttt{OpenBUGS}, \texttt{JAGS}, and \texttt{STAN}, a list containing each data object or constant can be passed to the function that calls the sampler.
-For example, in \texttt{JAGS}, the data might constructed like this:
-
-```{r}
-n <-  10 # binomial n
-y <- rbinom(1, n, p)
-
-jags_data <- list(
-  "y" = y,
-  "n" = n
-)
-
-run.jags("data" = jags_data, ... )
-```
-
-In \texttt{Nimble}, the constants need to be specified in a separate list from the data.
-
-```{r}
-n <-  10 # binomial n
-y <- rbinom(1, n, p)
-
-nimble_data <- list(
-  "y" = y
-)
-
-nimble_constants <- list(
-  "n" = n
-)
-
-nimbleMCMC("data" = nimble_data, "constants" = nimble_constants, ... )
-```
-
-## Monitoring Parameters
-
-In \texttt{OpenBUGS}, \texttt{JAGS}, \texttt{Nimble}, and \texttt{STAN}, specifying parameters for the sampler to monitor is as easy as providing a character vector with the names of the parameters to monitor:
-
-```{r}
-jags_monitors <- jags_monitor <- c("alpha", "beta")
-
-run.jags("monitor" = jags_monitor, ... )
-```
-
-## STAN Probabilty and Sampling Statements
-
-Inside the model block of a stan file, probability statements can be written in the same convention as in other bayesian sampling methods
-
-```{r}
-model{
-  y ~ binomial(n,p);
-  p ~ beta(1,1);
-}
-```
-
-This kind of probability statement does not actually perform any sampling, but is instead tranformed into a statement about the incremental log probability density of a target distribution.
-The log probability density function can be specified directly.
-
-```{r}
-model {
-  target += binomial_lpdf(y | n, p);
-  target += beta_lpdf(p | 1, 1);
-}
-```
-
-Either method yields the same result, but in some cases it is advantageous to use the incremental log probability density (or cumulative density) function to get exact log probability values for the model.
-In general, using the standard sampling statement syntax is preferred since it is easier to understand.
-
-## Greta's Unique Specification Language
-
-Implementing models in \texttt{greta} is very unique in that each data variable or parameter is initialized seperately with its own line of \texttt{R} code.
-Each data object in \texttt{R} is redefined using the \texttt{as\_data()} function in the \texttt{greta} package:
-```{r}
-y <- as_data(rpois(5, theta))
-```
-
-Each parameter is defined using a function indicating the prior distribution and the chosen prior parameters.
-
-```{r}
-theta <- gamma(3,1)
-```
-
-Once all the variables have been defined, the relationship between the data and the unknown parameters is established using the \texttt{distribution()} function:
-
-```{r}
-distribution(y) <- poisson(theta)
-```
-
-Finally, the greta model is constructed using \texttt{model()} function, which takes each of the monitored parameters as arguments:
-
-```{r}
-greta_model <- model(theta)
-
-mcmc("model" = greta_model, ... )
-```
-
-# How to Specify Tricky Models
-
-Each sampler has its own unique methods for using certain bayesian model details. We provide illustrations for some of the most common ones.
-
-##### Truncated Distribution
-
-Using a truncated distribution is a common practice in bayesian modeling, especially when specifying priors for parameters with positive domains.
-The syntax for specifying a truncated distribution is slightly different for each sampling method.
-Below we are using a truncated normal distribution with a lower bound of zero and no upper bound:
-
-```{r}
-## Truncated normal distribution in OpenBUGS
-tau ~ dnorm(0,1) I(0, )
-
-## Model Text for OpenBUGS using R2OpenBUGS
-tau ~ dnorm(0,1) %_% I(0, )
-
-## JAGS
-tau ~ dnorm(0,1) I(0, )
-
-## Nimble
-tau ~ T(dnorm(0,1), 0,)
-
-## Greta
-tau <- normal(0,1, truncation = c(0,Inf))
-
-## STAN
-# Parameter block initialization
-real<lower=0> tau;
-
-# Model block specification
-tau ~ normal(0,1);
-
-# Alternative solution
-tau ~ normal(0,1) T[0,]
 
 ```
-
-##### Flat Priors
-
-Sometimes using a flat, diffuse prior is appropriate.
-Most of the samplers allow for improper flat priors, although \texttt{JAGS} does not.
-In \texttt{STAN}, a parameter has is given a flat, improper prior by default if it doesn't have a distribution assigned.
-Simply declaring the parameter will do the job.
-```{r}
-## Flat prior in OpenBUGS
-theta ~ dflat()
-
-## Nimble
-theta ~ dflat()
-
-## Greta
-theta <- variable()
-
-## STAN
-# Parameter block initialization
-real theta;
-
-```
-
-##### Distribution Parameter Order
-
-Some samplers use different parameter orders for certain distributions.
-
-Some use the $\sigma$ parameterization of the normal distribution, where $\sigma$ is the standard deviation, and others use $\tau$ instead, where $\tau = 1 / \sigma^2$.
-Among bayesian samplers, \texttt{JAGS}, \texttt{OpenBUGS}, and \texttt{Nimble} use the $\tau$ parameteriztion, while \texttt{STAN} and \texttt{greta} use $\sigma$.
-
-\textit{\textbf{Normal}}: The $N(\mu,\sigma)$ parameterization is used by \texttt{STAN} and \texttt{greta}, where $\sigma$ is the standard deviation.
-The $N(\mu,\tau)$ parameterization is used by \texttt{JAGS}, \texttt{OpenBUGS}, and \texttt{Nimble}, which uses the precision parameter $\tau = 1 / \sigma^2$.
-The same parameterization guidelines are the same for the student-t distribution as well.
-
-\textit{\textbf{Binomial}}: The $Bin(n,p)$ parameterization is used by \texttt{OpenBUGS}, \texttt{JAGS}, and \texttt{Nimble}, while the $Bin(p,n)$ parameterization is used by \texttt{STAN} and \texttt{greta}.
-
-\textit{\textbf{Gamma}}: The $Gam(\alpha,\lambda)$ parameterization, where $\alpha$ is a shape parameter and $\lambda$ is a rate parameter, is used by \texttt{OpenBUGS}, \texttt{JAGS}, \texttt{Nimble}, and \texttt{greta}.
-The $Gam(\alpha,\beta)$ parameterization is used by \texttt{STAN}, using the scale parameter $\beta = 1 / \lambda$.
-
-\textit{\textbf{Weibull}}: The $Weibull(v,\lambda)$ parameterization is used by \texttt{OpenBUGS}, \texttt{JAGS}, \texttt{Nimble}, and \texttt{greta}, where $\lambda$ is a shape parameter.
-The $Weibull(v,\sigma)$ parameterization is used by \texttt{STAN}, where $\sigma = 1 / \lambda$.
-
-##### Equations inside distribution parameters
-
-In order to utilized calculated parameters inside \texttt{OpenBUGS}, a variable must be defined using the assignment operator that is the result of the calculation:
-
-```{r}
-y_hat <- alpha + beta * x
-y ~ dnorm(y_hat, tau)
-```
-
-Fortunately, the other samplers allow for variable calculations inside the actual parameter definition of a distribution function. Here is an example in \texttt{JAGS}:
-
-```{r}
-y ~ dnorm(alpha + beta * x, tau)
-```

### 2020-04-27 Paper.Rmd
ppp update and adding accuracy bench

@@
     - Sample a new value for $\theta_1$, from the full conditional distribution of $\theta_1$, given the most recent values of all other elements of $\theta$ and the data:
+
     $$\theta_1^{(1)} \sim p(\theta_1 | \theta_2^{(0)}, \theta_3^{(0)}, \dots , \theta_k^{(0)}, y).$$
@@
 We define the joint probability of the position $p$ and momentum $q$ as $\pi(q,p) = \pi(q|p) \pi(p)$.
-This structure allows us at anytime to throw out $q$ and recover the value of $p$, which iholds the position values of the parameter(s).
+This structure allows us at anytime to throw out $q$ and recover the value of $p$, which holds the position values of the parameter(s).
 The Hamiltonian is defined as
@@
   H(q,p) &= -\log \pi (p|q) \pi(q) \\
-   &= -\log \pi (p|q) - log \pi(q) \\
+   &= -\log \pi (p|q) - \log \pi(q) \\
    &= - K - V,\\
@@
 Greta, a very recent addition to the list of bayes computational methods, is made specifically for use in \texttt{R}, making the model and data specification process as easy as possible.
-It also uses Google TensorFlow to perform its calculations, and allows for more flexible use of CPUs and GPUs.
+Not only does it utilize Hamiltonian Monte Carlo methods for sampling, but it also uses Google TensorFlow to perform its calculations, and allows for more flexible use of CPUs and GPUs.
 Unfortunately, as of the writing of this, it is still in need of lots of development, and lacks an appropriate amount of documentation and support for users.
@@
     "file" = "stan_file.stan", "data" = model_data, 
-    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup, 
-    "control" = list("adapt_delta" = 0.99)
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
 )

### 2020-04-27 simple-models/discrete/binomial/beta-prior/beta-binomial-benchmark.R

@@ -7,8 +7,8 @@ library("bench")
 library("here")
 library("rjags"); library("runjags")
 library("R2OpenBUGS")
-library("nimble")
 library("rstan"); rstan_options(auto_write = FALSE)
+library("nimble")
 options("bayes_benchmark" = TRUE)
 
 ## Source code
@@ -18,7 +18,7 @@ source(here("simple-models", "discrete", "binomial", "beta-prior", "beta-binomia
 source(here("simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-bugs.R"))
 source(here("simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-nimble.R"))
 source(here("simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-stan.R"))
-# source(here("simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-greta.R"))
+source(here("simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-greta.R"))
 
 options("bayes_benchmark" = FALSE)
 
@@ -28,7 +28,7 @@ n_chains <- 4L
 n_iter <- 1e4L
 n_warmup <- 1e3L
 
-## benchmarking
+## benchmarking computation time
 ###################################################################################
 
 num_iterations <- 1 # Number of times you want to benchmark each model\
@@ -41,6 +41,15 @@ source(here("benchmark-function.R"))
 run_benchmark(rds_file_location)
 # run_benchmark(rds_file_location, stan_compile = TRUE)
 
+## benchmarking accuracy
+###################################################################################
 
+num_iterations <- 10 # Number of times you want to benchmark each model\
+
+source(here("benchmark-accuracy-function.R"))
+
+true_dist <- distr::Beta(3,9)
+
+run_accuracy_benchmark(true_dist,num_iterations, n_iter, n_warmup)
 
 

### 2020-04-30 PPP/Figures.Rmd

@@
 
-```{r}
+```{r, eval = FALSE}
 ## load library
@@
 ```
+
+```{r}
+typical_df <- tibble(x = seq(-4,4,length.out = 100), d_norm =  dnorm(x))
+
+typical_df %>% ggplot(aes(x,d_norm)) + 
+  annotate("rect", xmin=-1.7, xmax=1.7, ymin=0, ymax=Inf, alpha = 0.5) +
+  # geom_rect(aes(xmin=-1.7, xmax=1.7, ymin=0, ymax=Inf), alpha = 0.3, fill = "gray") +
+  geom_line(size = 2) + 
+  theme_minimal() + 
+  theme(
+    axis.title = element_blank(),
+    axis.text = element_blank()
+  ) + 
+  xlab(expression(pi*"("*theta*")"*" "*d*theta))
+
+```

### 2020-04-30 Simulation Results/Simulation Results.Rmd
add large version of hierch full beta binom

---
title: "Simulation Results"
author: "Evan Miyakawa"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
library(tidyverse); options(tibble.width = Inf)
library(magrittr)
library(here)
```

```{r}
exp_gamma_bench_files <- c(
  here("Simulation Results", "simple-models", "continuous", "exponential", "gamma-prior", "exponential-gamma-results.rds"),
  here("Simulation Results", "simple-models", "continuous", "exponential", "gamma-prior", "exponential-gamma-results-greta.rds"),
  here("Simulation Results", "simple-models", "continuous", "exponential", "gamma-prior", "exponential-gamma-results-stan.rds")
)

normal_gamma_tau_bench_files <- c(
  here("Simulation Results", "simple-models", "continuous", "normal", "gamma-prior-tau", "normal-gamma-tau-results.rds"),
  here("Simulation Results", "simple-models", "continuous", "normal", "gamma-prior-tau", "normal-gamma-tau-results-greta.rds"),
  here("Simulation Results", "simple-models", "continuous", "normal", "gamma-prior-tau", "normal-gamma-tau-results-stan.rds")
)

normal_normal_bench_files <- c(
  here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-results.rds"),
  here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-results-greta.rds"),
  here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-results-stan.rds")
)

beta_bernoulli_bench_files <- c(
  here("Simulation Results", "simple-models", "discrete", "bernoulli", "beta-prior", "beta-bernoulli-results.rds"),
  here("Simulation Results", "simple-models", "discrete", "bernoulli", "beta-prior", "beta-bernoulli-results-greta.rds"),
  here("Simulation Results", "simple-models", "discrete", "bernoulli", "beta-prior", "beta-bernoulli-results-stan.rds")
)

beta_binomial_bench_files <- c(
  here("Simulation Results", "simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-results.rds"),
  here("Simulation Results", "simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-results-greta.rds"),
  here("Simulation Results", "simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-results-stan.rds")
)

poisson_gamma_bench_files <- c(
  here("Simulation Results", "simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-results.rds"),
  here("Simulation Results", "simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-results-greta.rds"),
  here("Simulation Results", "simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-results-stan.rds")
)

combine_bench <- function(files, model_name) {
  combined_results <- files %>% map_dfr(~ readRDS(.))
  combined_results

  expression_string <- combined_results$expression[[7]]
  expression_string
  get_method_name <- function(expression_string) {
    string <- as.character(expression_string)[1]
    string
    if (string == "run.jags") {
      method <- "JAGS"
    } else if (string == "bugs") {
      method <- "OpenBUGS"
    } else if (string == "nimbleMCMC") {
      method <- "Nimble"
    } else if (string == "stan") {
      method <- "STAN"
    } else if (string == "mcmc") {
      method <- "Greta"
    } else if (string == "stan_model") {
      method <- "STAN Compile"
    }
    method
  }

  method_names <- combined_results$expression %>% map_chr(get_method_name)

  method_mean <- combined_results$time %>% map_dbl(~ mean(.))
  method_sd <- combined_results$time %>% map_dbl(~ sd(.))

  results <- combined_results %>% 
    mutate(method = method_names, mean = method_mean, sd = method_sd, 
           model_name = model_name) %>% 
    select(method, mean, median, min, sd, n_itr, model_name)

  rows_to_keep <- results$method %>% unique() %>% 
    map_dbl(~ which(results$method == .)[1])

  results <- results[rows_to_keep,] %>% 
    arrange(method)

}

exp_gamma_results <- exp_gamma_bench_files %>% 
  combine_bench("exponential_gamma")
normal_gamma_tau_results <- normal_gamma_tau_bench_files %>%
  combine_bench("normal_gamma_tau")
normal_normal_results <- normal_normal_bench_files %>%
  combine_bench("normal_normal")
beta_bernoulli_results <- beta_bernoulli_bench_files %>%
  combine_bench("beta_bernoulli")
beta_binomial_results <- beta_binomial_bench_files %>%
  combine_bench("beta_binomial")
poisson_gamma_results <- poisson_gamma_bench_files %>%
  combine_bench("poisson_gamma")
poisson_gamma_results

x <- readRDS(bench_files[3])
x
```

### 2020-04-30 hierarchical-models/large-full-beta-binomial/hierarchical-full-beta-binomial-benchmark.R

@@ -1,0 +1,45 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("bench")
+library("here")
+library("rjags"); library("runjags")
+library("R2OpenBUGS")
+library("nimble")
+library("rstan"); rstan_options(auto_write = FALSE)
+library("greta")
+options("bayes_benchmark" = TRUE)
+
+## Source code
+################################################################################
+
+source(here("hierarchical-models", "large-full-beta-binomial", "hierarchical-full-beta-binomial-jags.R"))
+source(here("hierarchical-models", "large-full-beta-binomial", "hierarchical-full-beta-binomial-bugs.R"))
+source(here("hierarchical-models", "large-full-beta-binomial", "hierarchical-full-beta-binomial-nimble.R"))
+source(here("hierarchical-models", "large-full-beta-binomial", "hierarchical-full-beta-binomial-stan.R"))
+source(here("hierarchical-models", "large-full-beta-binomial", "hierarchical-full-beta-binomial-greta.R"))
+
+options("bayes_benchmark" = FALSE)
+
+## Set Parameters
+################################################################################
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+## benchmarking
+###################################################################################
+
+num_iterations <- 50 # Number of times you waçnt to benchmark each model\
+
+rds_file_location <- here("hierarchical-models", "large-full-beta-binomial", "hierarchical-full-beta-binomial.rds")
+
+source(here("benchmark-function.R"))
+
+# choose to include stan compilation time with argument "stan_compile = TRUE"
+run_benchmark(rds_file_location)
+run_benchmark(rds_file_location, stan_compile = TRUE)
+
+

### 2020-04-30 hierarchical-models/large-full-beta-binomial/hierarchical-full-beta-binomial-greta.R

@@ -1,0 +1,70 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("greta")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+## generate/specify data
+################################################################################
+
+alpha <- 2 #beta distribution alpha
+beta <- 1 #beta distribution beta
+N <- 1000 # sample size
+
+set.seed(1)
+
+(n <- rbinom(N, 20, 0.5))
+(theta <- rbeta(N, alpha, beta))
+(y <- rbinom(N, n, theta))
+
+n <- as_data(n)
+y <- as_data(y)
+
+## specify greta model
+################################################################################
+
+alpha <- gamma(1,1)
+beta <- gamma(1,1)
+theta <- beta(alpha,beta, dim = c(N,1))
+distribution(y) <- binomial(n, theta, dim = c(N,1))
+
+greta_model <- model(theta, alpha, beta)
+
+plot(greta_model)
+
+## configure model settings
+################################################################################
+
+n_chains <- 4
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+
+## fit model
+################################################################################
+
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  greta_fit <- mcmc(
+    "model" = greta_model, "n_samples" = n_iter,
+    "warmup" = n_warmup, "chains" = n_chains
+  )
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(greta_fit)
+  
+}
+

### 2020-04-30 hierarchical-models/large-full-beta-binomial/hierarchical-full-beta-binomial-jags.R

@@ -1,0 +1,103 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rjags"); library("runjags")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+alpha <- 2 #beta distribution alpha
+beta <- 1 #beta distribution beta
+N <- 1000 # sample size
+
+set.seed(1)
+
+(n <- rbinom(N, 20, 0.5))
+(theta <- rbeta(N, alpha, beta))
+(y <- rbinom(N, n, theta))
+
+
+jags_data <- list(
+  "y" = y,
+  "N" = N,
+  "n" = n
+)
+
+
+
+
+## specify jags model
+################################################################################
+
+jags_model <- "
+  model{
+    for (i in 1:N) {
+      theta[i] ~ dbeta(alpha, beta)
+      y[i] ~ dbin(theta[i], n[i])
+    }
+    alpha ~ dgamma(1,1)
+    beta ~ dgamma(1,1)
+  }
+"
+
+jags_monitor <- c("theta", "alpha", "beta")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  jags_fit <- run.jags(
+    "model" = jags_model, "data" = jags_data, "monitor" = jags_monitor, 
+    "n.chains" = n_chains, "sample" = n_iter, "burnin" = n_warmup
+  ) 
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  jags_fit
+  
+  jags_fit_object <- jags_fit$mcmc %>% as.array()
+  dim(jags_fit_object) <- c(dim(jags_fit_object), 1)
+  dimnames(jags_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = jags_monitor
+  )
+  
+  
+  jags_fit_object %>% bayesplot::mcmc_dens()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  jags_fit_object %>% mcmc_acf_bar()
+  jags_fit_object %>% mcmc_trace()
+  
+  
+  
+  
+}
+

### 2020-04-30 hierarchical-models/large-full-beta-binomial/hierarchical-full-beta-binomial-nimble.R

@@ -1,0 +1,91 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("nimble")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+alpha <- 2 #beta distribution alpha
+beta <- 1 #beta distribution beta
+N <- 1000 # sample size
+
+set.seed(1)
+
+(n <- rbinom(N, 20, 0.5))
+(theta <- rbeta(N, alpha, beta))
+(y <- rbinom(N, n, theta))
+
+
+nimble_data <- list(
+  "y" = y,
+  "n" = n
+)
+
+nimble_constants <- list(
+  "N" = N
+)
+
+
+## specify nimble model
+################################################################################
+
+nimble_model <- nimbleCode({
+  for (i in 1:N) {
+    theta[i] ~ dbeta(alpha, beta)
+    y[i] ~ dbin(theta[i], n[i])
+  }
+  alpha ~ dgamma(1,1)
+  beta ~ dgamma(1,1)
+})
+
+nimble_monitor <- c("theta", "alpha", "beta")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+nimble_inits <- list(
+  "theta" = rbeta(10,2,1),
+  "alpha" = rgamma(1,1,1),
+  "beta" = rgamma(1,1,1)
+)
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  nimble_fit <- nimbleMCMC(
+    "code" = nimble_model, "data" = nimble_data, "constants" = nimble_constants,
+    "inits" = nimble_inits, "monitors" = nimble_monitor, "nchains" = n_chains, 
+    "niter" = n_iter, "nburnin" = n_warmup, "summary" = TRUE
+  )
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  nimble_fit$summary$all.chains
+  
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+}  
+
+
+

### 2020-05-01 Simulation Results/Simulation Results.Rmd
sim results

@@
 
+hierarchical_beta_binomial_bench_files <- c(
+  here("Simulation Results", "hierarchical-models", "beta-binomial", "hierarchical-beta-binomial-results.rds"),
+  here("Simulation Results", "hierarchical-models", "beta-binomial", "hierarchical-beta-binomial-results-greta.rds"),
+  here("Simulation Results", "hierarchical-models", "beta-binomial", "hierarchical-beta-binomial-results-stan.rds")
+)
+
+hierarchical_full_beta_binomial_bench_files <- c(
+  here("Simulation Results", "hierarchical-models", "full-beta-binomial", "hierarchical-full-beta-binomial-results.rds"),
+  here("Simulation Results", "hierarchical-models", "full-beta-binomial", "hierarchical-full-beta-binomial-results-greta.rds"),
+  here("Simulation Results", "hierarchical-models", "full-beta-binomial", "hierarchical-full-beta-binomial-results-stan.rds")
+)
+
+hierarchical_logistic_regression_bench_files <- c(
+  here("Simulation Results", "regression-models", "hierarchical-logistic-regression",  "hierarchical-logistic-regression-results.rds"),
+  here("Simulation Results", "regression-models", "hierarchical-logistic-regression",  "hierarchical-logistic-regression-results-greta.rds"),
+  here("Simulation Results", "regression-models", "hierarchical-logistic-regression",  "hierarchical-logistic-regression-results-stan.rds")
+)
+
+linear_regression_bench_files <- c(
+  here("Simulation Results", "regression-models", "linear-regression",  "linear-regression-results.rds"),
+  here("Simulation Results", "regression-models", "linear-regression",  "linear-regression-results-greta.rds"),
+  here("Simulation Results", "regression-models", "linear-regression",  "linear-regression-results-stan.rds")
+)
+
+logistic_regression_bench_files <- c(
+  here("Simulation Results", "regression-models", "logistic-regression",  "logistic-regression-results.rds"),
+  here("Simulation Results", "regression-models", "logistic-regression",  "logistic-regression-results-greta.rds"),
+  here("Simulation Results", "regression-models", "logistic-regression",  "logistic-regression-results-stan.rds")
+)
+
+poisson_regression_bench_files <- c(
+  here("Simulation Results", "regression-models", "poisson-regression",  "poisson-regression-results.rds"),
+  here("Simulation Results", "regression-models", "poisson-regression",  "poisson-regression-results-greta.rds"),
+  here("Simulation Results", "regression-models", "poisson-regression",  "poisson-regression-results-stan.rds")
+)
+
+cox_exponential_bench_files <- c(
+  here("Simulation Results", "survival-models", "cox-exponential-results.rds"),
+  here("Simulation Results", "survival-models", "cox-exponential-results-greta.rds"),
+  here("Simulation Results", "survival-models", "cox-exponential-results-stan.rds")
+)
+
+files <- logistic_regression_bench_files
+
 combine_bench <- function(files, model_name) {
@@
 
-  expression_string <- combined_results$expression[[7]]
-  expression_string
+  # expression_string <- combined_results$expression[[7]]
+  # expression_string
   get_method_name <- function(expression_string) {
@@
 
+combine_bench_indiv <- function(files, model_name) {
+  combined_results <- files %>% map_dfr(~ readRDS(.))
+  combined_results
+
+  # expression_string <- combined_results$expression[[7]]
+  # expression_string
+  get_method_name <- function(expression_string) {
+    string <- as.character(expression_string)[1]
+    string
+    if (string == "run.jags") {
+      method <- "JAGS"
+    } else if (string == "bugs") {
+      method <- "OpenBUGS"
+    } else if (string == "nimbleMCMC") {
+      method <- "Nimble"
+    } else if (string == "stan") {
+      method <- "STAN"
+    } else if (string == "mcmc") {
+      method <- "Greta"
+    } else if (string == "stan_model") {
+      method <- "STAN Compile"
+    }
+    method
+  }
+
+  method_names <- combined_results$expression %>% map_chr(get_method_name)
+
+  times <- combined_results$time[[1]]
+
+  time_results <- combined_results$time
+
+  names(time_results) <- method_names
+
+  get_iterations_for_method <- function(times, name) {
+    tibble(
+      method = name,
+      time = times,
+      model_name = model_name
+    )
+  }
+
+  res <- time_results %>% imap_dfr(get_iterations_for_method)
+
+  # 
+  # method_mean <- combined_results$time %>% map_dbl(~ mean(.))
+  # method_sd <- combined_results$time %>% map_dbl(~ sd(.))
+  # 
+  # results <- combined_results %>% 
+  #   mutate(method = method_names, mean = method_mean, sd = method_sd, 
+  #          model_name = model_name) %>% 
+  #   select(method, mean, median, min, sd, n_itr, model_name)
+  # 
+  # rows_to_keep <- results$method %>% unique() %>% 
+  #   map_dbl(~ which(results$method == .)[1])
+  # 
+  # results <- results[rows_to_keep,] %>% 
+  #   arrange(method)
+
+  res
+
+}
+
+# exp_gamma_results <- exp_gamma_bench_files %>% 
+#   combine_bench("exponential_gamma")
+# normal_gamma_tau_results <- normal_gamma_tau_bench_files %>%
+#   combine_bench("normal_gamma_tau")
+# normal_normal_results <- normal_normal_bench_files %>%
+#   combine_bench("normal_normal")
+# beta_bernoulli_results <- beta_bernoulli_bench_files %>%
+#   combine_bench("beta_bernoulli")
+# beta_binomial_results <- beta_binomial_bench_files %>%
+#   combine_bench("beta_binomial")
+# poisson_gamma_results <- poisson_gamma_bench_files %>%
+#   combine_bench("poisson_gamma")
+# hier_beta_binomial_results <- hierarchical_beta_binomial_bench_files %>%
+#   combine_bench("hierarchical_beta_binomial")
+# hier_full_bb_results <- hierarchical_full_beta_binomial_bench_files %>%
+#   combine_bench("hierarchical_full_beta_binomial")
+# hier_log_reg_results <- hierarchical_logistic_regression_bench_files %>%
+#   combine_bench("hierarchical_logistic_regression")
+# linear_reg_results <- linear_regression_bench_files %>% 
+#   combine_bench("linear regression")
+# logistic_reg_results <- logistic_regression_bench_files %>% 
+#   combine_bench("logistic regression")
+# poisson_reg_results <- poisson_regression_bench_files %>% 
+#   combine_bench("poisson regression")
+
 exp_gamma_results <- exp_gamma_bench_files %>% 
-  combine_bench("exponential_gamma")
+  combine_bench_indiv("exponential_gamma")
 normal_gamma_tau_results <- normal_gamma_tau_bench_files %>%
-  combine_bench("normal_gamma_tau")
+  combine_bench_indiv("normal_gamma_tau")
 normal_normal_results <- normal_normal_bench_files %>%
-  combine_bench("normal_normal")
+  combine_bench_indiv("normal_normal")
 beta_bernoulli_results <- beta_bernoulli_bench_files %>%
-  combine_bench("beta_bernoulli")
+  combine_bench_indiv("beta_bernoulli")
 beta_binomial_results <- beta_binomial_bench_files %>%
-  combine_bench("beta_binomial")
+  combine_bench_indiv("beta_binomial")
 poisson_gamma_results <- poisson_gamma_bench_files %>%
-  combine_bench("poisson_gamma")
-poisson_gamma_results
+  combine_bench_indiv("poisson_gamma")
+hier_beta_binomial_results <- hierarchical_beta_binomial_bench_files %>%
+  combine_bench_indiv("hierarchical_beta_binomial")
+hier_full_bb_results <- hierarchical_full_beta_binomial_bench_files %>%
+  combine_bench_indiv("hierarchical_full_beta_binomial")
+hier_log_reg_results <- hierarchical_logistic_regression_bench_files %>%
+  combine_bench_indiv("hierarchical_logistic_regression")
+linear_reg_results <- linear_regression_bench_files %>% 
+  combine_bench_indiv("linear regression")
+logistic_reg_results <- logistic_regression_bench_files %>% 
+  combine_bench_indiv("logistic regression")
+poisson_reg_results <- poisson_regression_bench_files %>% 
+  combine_bench_indiv("poisson regression")
 
-x <- readRDS(bench_files[3])
-x
 ```
+
+```{r}
+# bench_results <- bind_rows(
+#   exp_gamma_results,
+#   normal_gamma_tau_results,
+#   normal_normal_results,
+#   beta_bernoulli_results, 
+#   beta_binomial_results,
+#   poisson_gamma_results,
+#   hier_beta_binomial_results,
+#   hier_full_bb_results,
+#   hier_log_reg_results,
+#   linear_reg_results,
+#   logistic_reg_results,
+#   poisson_reg_results
+# )
+# 
+# bench_results
+# 
+# saveRDS(bench_results, file = "bench_results.RDS")
+
+bench_results_indiv <- bind_rows(
+  exp_gamma_results,
+  normal_gamma_tau_results,
+  normal_normal_results,
+  beta_bernoulli_results, 
+  beta_binomial_results,
+  poisson_gamma_results,
+  hier_beta_binomial_results,
+  hier_full_bb_results,
+  hier_log_reg_results,
+  linear_reg_results,
+  logistic_reg_results,
+  poisson_reg_results
+)
+
+bench_results_indiv
+
+bench_results_indiv <- bench_results_indiv %>% 
+  filter(!(method == "STAN Compile" & time < 5))
+
+saveRDS(bench_results_indiv, file = "bench_results_indiv.RDS")
+```
+
+```{r}
+# bench_results %>% 
+#   ggplot(aes(mean, color = method)) + 
+#   geom_boxplot() + 
+#   coord_flip() + 
+#   xlim(0,90)
+# 
+# bench_results %>% 
+#   ggplot(aes(mean, fill = method)) + 
+#   geom_histogram() + 
+#   facet_wrap(vars(method))
+#   # coord_flip() + 
+#   # xlim(0,90)
+
+```
+
+```{r}
+bench_results_indiv <- readRDS("bench_results_indiv.RDS")
+
+bench_results_indiv %>%
+  ggplot(aes(time, color = method)) +
+  geom_boxplot() +
+  coord_flip() +
+  xlim(0,90)
+
+bench_results_indiv %>%
+  ggplot(aes(time, fill = method)) +
+  geom_histogram(aes(y = stat(density)), bins = 20) + 
+  # xlim(0,100) +
+  facet_wrap(vars(method))
+
+bench_results_indiv %>%
+  ggplot(aes(time, fill = method)) +
+  geom_histogram(aes(y = stat(density)), bins = 20) + 
+  xlim(0,100) +
+  facet_wrap(vars(method))
+
+```
+
+##

### 2020-05-03 Simulation Results/Simulation Results.Rmd
tvd update

@@
 
-##
+## Accuracy benchmarking
+
+```{r}
+bugs_fit_condense <- readRDS(here("Simulation Results", "simple-models", 
+                                  "discrete", "binomial", "beta-prior", 
+                                  "bugs_fit_condense_beta_binomial.RDS"))
+jags_fit_condense <- readRDS(here("Simulation Results", "simple-models", 
+                                  "discrete", "binomial", "beta-prior", 
+                                  "jags_fit_condense_beta_binomial.RDS"))
+stan_fit_condense <- readRDS(here("Simulation Results", "simple-models", 
+                                  "discrete", "binomial", "beta-prior", 
+                                  "stan_fit_condense_beta_binomial.RDS"))
+greta_fit_condense <- readRDS(here("Simulation Results", "simple-models", 
+                                  "discrete", "binomial", "beta-prior", 
+                                  "greta_fit_condense_beta_binomial.RDS"))
+nimble_fit_condense <- readRDS(here("Simulation Results", "simple-models", 
+                                  "discrete", "binomial", "beta-prior", 
+                                  "nimble_fit_condense_beta_binomial.RDS"))
+dists <- tibble(
+  x = c(bugs_fit_condense, jags_fit_condense, stan_fit_condense, 
+        greta_fit_condense, nimble_fit_condense),
+  method = c(rep("bugs", length(bugs_fit_condense)),
+             rep("jags", length(jags_fit_condense)),
+             rep("stan", length(stan_fit_condense)),
+             rep("greta", length(greta_fit_condense)),
+             rep("nimble", length(nimble_fit_condense))
+             )
+)
+
+dists %>% 
+  ggplot(aes(x, color = method)) + 
+  geom_density(adjust = 2) + 
+  stat_function(fun = dbeta, args = list(3,9), color = "black")
+
+tibble(x = bugs_fit_condense) %>% 
+  ggplot(aes(x)) + 
+    geom_density(adjust = 2) + 
+    stat_function(fun = dbeta, args = list(3,9), color = "red")
+
+tibble(x = jags_fit_condense) %>% 
+  ggplot(aes(x)) + 
+    geom_density(adjust = 2) + 
+    stat_function(fun = dbeta, args = list(3,9), color = "red")
+
+tibble(x = stan_fit_condense) %>% 
+  ggplot(aes(x)) + 
+    geom_density(adjust = 2) + 
+    stat_function(fun = dbeta, args = list(3,9), color = "red")
+
+```
+
+```{r}
+exp_gamma_acc_files <- c(
+  here("Simulation Results", "simple-models", "continuous", "exponential", "gamma-prior", "exponential-gamma-accuracy.rds"),
+  here("Simulation Results", "simple-models", "continuous", "exponential", "gamma-prior", "exponential-gamma-accuracy-nimble.rds")
+)
+
+normal_gamma_tau_acc_files <- c(
+  here("Simulation Results", "simple-models", "continuous", "normal", "gamma-prior-tau", "normal-gamma-tau-accuracy.rds"),
+  here("Simulation Results", "simple-models", "continuous", "normal", "gamma-prior-tau", "normal-gamma-tau-accuracy-nimble.rds")
+)
+
+normal_normal_acc_files <- c(
+  here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-accuracy.rds")
+  # here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-accuracy-nimble.rds")
+)
+
+beta_bernoulli_acc_files <- c(
+  here("Simulation Results", "simple-models", "discrete", "bernoulli", "beta-prior", "beta-bernoulli-accuracy.rds"),
+  here("Simulation Results", "simple-models", "discrete", "bernoulli", "beta-prior", "beta-bernoulli-accuracy-nimble.rds")
+)
+
+beta_binomial_acc_files <- c(
+  here("Simulation Results", "simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-accuracy.rds"),
+  here("Simulation Results", "simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-accuracy-nimble.rds")
+)
+
+poisson_gamma_acc_files <- c(
+  here("Simulation Results", "simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-accuracy.rds"),
+  here("Simulation Results", "simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-accuracy-nimble.rds")
+)
+
+files <- normal_normal_acc_files
+model_name <- "poisson_gamma"
+
+combine_time_indiv <- function(files, model_name) {
+  combined_results <- files %>% map_dfc(~ readRDS(.)) %>% 
+    select(-iter, -iter1) %>% 
+    mutate(model = model_name)
+
+  combined_results
+}
+
+exp_gamma_acc_df <- exp_gamma_acc_files %>% 
+  combine_time_indiv("exponential_gamma")
+normal_gamma_tau_acc_df <- normal_gamma_tau_acc_files %>% 
+  combine_time_indiv("normal_gamma_tau")
+# normal_normal_acc_df <- normal_normal_acc_files %>% 
+#   combine_time_indiv("normal_normal")
+beta_bernoulli_acc_df <- beta_bernoulli_acc_files %>% 
+  combine_time_indiv("beta_bernoulli")
+beta_binomial_acc_df <- beta_binomial_acc_files %>% 
+  combine_time_indiv("beta_binomial")
+poisson_gamma_acc_df <- poisson_gamma_acc_files %>% 
+  combine_time_indiv("poisson_gamma")
+
+time_results <- bind_rows(
+  exp_gamma_acc_df,
+  normal_gamma_tau_acc_df,
+  beta_bernoulli_acc_df,
+  beta_binomial_acc_df,
+  poisson_gamma_acc_df
+)
+
+time_results %>% 
+  pivot_longer(
+    cols = c(jags_dist, bugs_dist, stan_dist, greta_dist, nimble_dist), 
+    names_to = "method",
+    values_to = "TVD"
+  ) %>% 
+  ggplot(aes(TVD)) + 
+    geom_histogram() + 
+    facet_wrap(facets = "method")
+
+time_results %>% 
+  pivot_longer(
+    cols = c(jags_dist, bugs_dist, stan_dist, greta_dist, nimble_dist), 
+    names_to = "method",
+    values_to = "TVD"
+  ) %>% 
+  ggplot(aes(TVD)) + 
+    geom_boxplot(aes(color = method)) + 
+    coord_flip() + 
+    xlim(0,0.25)
+
+```

### 2020-05-03 working with total variation distance.R

@@ -1,0 +1,20 @@
+jags_dist <- TotalVarDist(true_dist,jags_fit_condense)
+bugs_dist <- TotalVarDist(true_dist,bugs_fit_condense)
+stan_dist <- TotalVarDist(true_dist,stan_fit_condense)
+greta_dist <- TotalVarDist(true_dist,greta_fit_condense)
+
+jags_dist <- TotalVarDist(true_dist,jags_fit_condense, 
+                          asis.smooth.discretize = "asis")
+jags_dist
+bugs_dist <- TotalVarDist(true_dist,bugs_fit_condense, 
+                          asis.smooth.discretize = "asis")
+bugs_dist
+stan_dist <- TotalVarDist(true_dist,stan_fit_condense)
+greta_dist <- TotalVarDist(true_dist,greta_fit_condense)
+
+tibble(
+  jags_dist = jags_dist,
+  bugs_dist = bugs_dist,
+  stan_dist = stan_dist,
+  greta_dist = greta_dist
+)

### 2020-05-04 Simulation Results/Simulation Results.Rmd
ppp updates

@@
 knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
-library(tidyverse); options(tibble.width = Inf)
+library(tidyverse); options(tibble.width = Inf); theme_set(theme_bw())
 library(magrittr)
@@
     } else if (string == "stan") {
-      method <- "STAN"
+      method <- "Stan"
     } else if (string == "mcmc") {
@@
     } else if (string == "stan_model") {
-      method <- "STAN Compile"
+      method <- "Stan Compile"
     }
@@
 bench_results_indiv <- bench_results_indiv %>% 
-  filter(!(method == "STAN Compile" & time < 5))
+  filter(!(method == "Stan Compile" & time < 5))
 
-saveRDS(bench_results_indiv, file = "bench_results_indiv.RDS")
+# saveRDS(bench_results_indiv, file = "bench_results_indiv.RDS")
+# saveRDS(bench_results_indiv, file = "bench_results_indiv2.RDS")
 ```
@@
 ```{r}
-bench_results_indiv <- readRDS("bench_results_indiv.RDS")
+bench_results_indiv <- readRDS("bench_results_indiv2.RDS")
 
 bench_results_indiv %>%
-  ggplot(aes(time, color = method)) +
-  geom_boxplot() +
-  coord_flip() +
-  xlim(0,90)
+  ggplot(aes(x = method, y = time)) +
+  geom_boxplot(aes(color = method), coef = 2, outlier.size = 0.75) +
+  # geom_boxplot(aes(color = method)) +
+  # coord_flip() +
+  ylim(0,90) +
+  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan", "Nimble", "Greta", "Stan Compile")) + 
+  theme(legend.position="none",
+        plot.title = element_text(hjust = 0.5)) + 
+  xlab("Implementation") + 
+  ylab("Completion Time") + 
+  ggtitle("Implementation Completion Time Boxplots")
 
@@
   # xlim(0,100) +
-  facet_wrap(vars(method))
+  facet_wrap(vars(method)) + 
+  theme(legend.position="none",
+        plot.title = element_text(hjust = 0.5)) + 
+  ggtitle("Implementation Completion Time Histograms") + 
+  xlab("Completion Time") + 
+  ylab("Count Density")
 
@@
   xlim(0,100) +
-  facet_wrap(vars(method))
+  facet_wrap(vars(method)) + 
+  theme(legend.position="none",
+        plot.title = element_text(hjust = 0.5)) + 
+  ggtitle("Implementation Completion Time Histograms") + 
+  xlab("Completion Time") + 
+  ylab("Count Density")
+
+bench_results_indiv %>%
+  filter(method %in% c("JAGS", "Nimble", "OpenBUGS", "Stan")) %>% 
+  ggplot(aes(time, fill = method)) +
+  geom_histogram(aes(y = stat(density)), bins = 20) + 
+  # xlim(0,100) +
+  facet_wrap(vars(method)) + 
+  theme(legend.position="none",
+        plot.title = element_text(hjust = 0.5)) + 
+  ggtitle("Implementation Completion Time Histograms") + 
+  xlab("Completion Time") + 
+  ylab("Count Density")
 
@@
 
-files <- normal_normal_acc_files
+files <- poisson_gamma_acc_files
 model_name <- "poisson_gamma"
@@
     mutate(model = model_name)
+
+  stan_mean <- combined_results$stan_dist %>% mean()
+
+  combined_results <- combined_results %>% 
+    mutate(JAGS = jags_dist / stan_mean, 
+           OpenBUGS = bugs_dist / stan_mean, 
+           Stan = stan_dist / stan_mean, 
+           Nimble = nimble_dist / stan_mean, 
+           Greta = greta_dist / stan_mean
+           )
 
@@
   pivot_longer(
-    cols = c(jags_dist, bugs_dist, stan_dist, greta_dist, nimble_dist), 
+    # cols = c(jags_dist, bugs_dist, stan_dist, greta_dist, nimble_dist), 
+    cols = c(JAGS, OpenBUGS, Stan, Nimble, Greta), 
     names_to = "method",
-    values_to = "TVD"
+    values_to = "TVD_Ratio"
   ) %>% 
-  ggplot(aes(TVD)) + 
-    geom_histogram() + 
-    facet_wrap(facets = "method")
+  ggplot(aes(TVD_Ratio)) + 
+    geom_histogram(aes(fill = method)) + 
+    facet_wrap(facets = "method") + 
+  theme(legend.position="none",
+        plot.title = element_text(hjust = 0.5)) + 
+  ggtitle("Implementation TVD Accuracy Histograms") + 
+  xlab("Implementation") + 
+  ylab("TVD Ratio")
 
@@
   pivot_longer(
-    cols = c(jags_dist, bugs_dist, stan_dist, greta_dist, nimble_dist), 
+    # cols = c(jags_dist, bugs_dist, stan_dist, greta_dist, nimble_dist), 
+    cols = c(JAGS, OpenBUGS, Stan, Nimble, Greta), 
     names_to = "method",
-    values_to = "TVD"
+    values_to = "TVD_Ratio"
   ) %>% 
-  ggplot(aes(TVD)) + 
+  ggplot(aes(x = method, y = TVD_Ratio)) + 
     geom_boxplot(aes(color = method)) + 
-    coord_flip() + 
-    xlim(0,0.25)
+  ggtitle("Implementation TVD Accuracy Boxplots") + 
+  xlab("Implementation") + 
+  ylab("TVD Ratio") + 
+  scale_x_discrete(limits = c("OpenBUGS", "JAGS", "Nimble", "Stan", "Greta")) + 
+  theme(legend.position="none",
+        plot.title = element_text(hjust = 0.5))
 

### 2020-05-04 working with total variation distance.R

@@ -1,3 +1,6 @@
+library(distr)
+library(distrEx)
+
 jags_dist <- TotalVarDist(true_dist,jags_fit_condense)
 bugs_dist <- TotalVarDist(true_dist,bugs_fit_condense)
 stan_dist <- TotalVarDist(true_dist,stan_fit_condense)

### 2020-05-04 HMC Talk Notes.Rmd
ppp monday work computer

@@
 author: "Evan Miyakawa"
-date: "4/7/2020"
 output: pdf_document
@@
 ```
+
+## New Notes 5-4-2020
+
+- Might want to include Hamiltonian equations to explain kinetic and potential energy when then giving the visual illustration
+- Potential energy is specified by the problem, but kinetic energy is not
+- Good illustration might be the project up/down at 36:00 in video
+
+## Old notes
 

### 2020-05-04 PPP/Figures.Rmd
ppp monday work computer

@@
 
-log_df <- tibble(x = seq(-4,4,length.out = 100), d_norm = - dnorm(x))
+log_df <- tibble(x = seq(-4,4,length.out = 100), d_norm = dnorm(x), log_d_norm =  - log(d_norm))
 
-log_df %>% ggplot(aes(x,d_norm)) + 
+log_df %>% ggplot(aes(x,log_d_norm)) + 
   geom_line(size = 2) + 

### 2020-05-04 Simulation Results/Simulation Results.Rmd
ppp monday work computer

@@
 ```{r}
-bench_results_indiv <- readRDS("bench_results_indiv2.RDS")
+bench_results_indiv <- readRDS(here("bench_results_indiv2.RDS"))
 
@@
   ggplot(aes(x = method, y = time)) +
-  geom_boxplot(aes(color = method), coef = 2, outlier.size = 0.75) +
+  geom_boxplot(aes(color = method), coef = 2, outlier.size = 0.75, size = 0.7) +
   # geom_boxplot(aes(color = method)) +
@@
   theme(legend.position="none",
-        plot.title = element_text(hjust = 0.5)) + 
+        plot.title = element_text(hjust = 0.5),
+        text = element_text(size=20)) + 
   xlab("Implementation") + 
@@
   theme(legend.position="none",
-        plot.title = element_text(hjust = 0.5)) + 
+        plot.title = element_text(hjust = 0.5),
+        text = element_text(size=20)) + 
   ggtitle("Implementation Completion Time Histograms") + 
@@
   theme(legend.position="none",
-        plot.title = element_text(hjust = 0.5)) + 
+        plot.title = element_text(hjust = 0.5),
+        text = element_text(size=20)) + 
   ggtitle("Implementation TVD Accuracy Histograms") + 
@@
   ggplot(aes(x = method, y = TVD_Ratio)) + 
-    geom_boxplot(aes(color = method)) + 
+    geom_boxplot(aes(color = method), size = 0.7) + 
   ggtitle("Implementation TVD Accuracy Boxplots") + 
@@
   theme(legend.position="none",
-        plot.title = element_text(hjust = 0.5))
+        plot.title = element_text(hjust = 0.5),
+        text = element_text(size=20))
 

### 2020-05-04 sim-save.R

@@ -1,8 +1,10 @@
 x <- run_benchmark(rds_file_location, stan_only = TRUE, stan_compile = TRUE)
+x <- run_benchmark(rds_file_location, greta_only = TRUE)
 x <- run_benchmark(rds_file_location)
+x <- run_accuracy_benchmark(true_dist, num_iterations, nimble = TRUE)
 
 saveRDS(x, 
-        file = "ar-model-results-no-nimble.rds")
+        file = "normal-normal-accuracy-nimble.rds")
 
 x
 

### 2020-05-05 PPP/Figures.Rmd
figures udpate

@@
 library(magrittr)
-library("greta")
+
 ```
@@
 ```{r}
+library("greta")
 
@@
 ```
+
+```{r}
+set.seed(1)
+n <- 1000
+df <- tibble(x = rnorm(n), y = rnorm(n, sd = 3)) 
+big_n <- 100000
+full_df <- tibble(x = rnorm(big_n), y = rnorm(big_n, sd = 3))
+
+full_df %>% 
+  ggplot() + 
+  stat_density_2d(aes(x, y, fill = ..level..), geom = "polygon", alpha = 0.5) + 
+  # geom_point(aes(x_data, y_data)) + 
+  scale_fill_distiller(palette=1, direction=1) + 
+  theme(
+    legend.position = 'none',
+    axis.title = element_blank(),
+    axis.text = element_blank()
+  )
+
+```
+
+<!-- ```{r} -->
+<!-- full_df_with_data <- full_df %>%  -->
+<!--   mutate(x_data = rep(df$x[1], 100000)) %>%  -->
+<!--   mutate(y_data = rep(df$y[1], 100000))  -->
+
+<!-- full_df_with_data %>%  -->
+<!--   ggplot() +  -->
+<!--   stat_density_2d(aes(x, y, fill = ..level..), geom = "polygon", alpha = 0.5) +  -->
+<!--   geom_point(aes(x_data, y_data)) + -->
+<!--   scale_fill_distiller(palette=1, direction=1) +  -->
+<!--   theme( -->
+<!--     legend.position = 'none', -->
+<!--     axis.title = element_blank(), -->
+<!--     axis.text = element_blank() -->
+<!--   ) -->
+
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- full_df_with_data <- full_df %>%  -->
+<!--   mutate(x_data = c(df$x[3], rep(df$x[1], 100000 - 1))) %>%  -->
+<!--   mutate(y_data = c(df$y[3], rep(df$y[1], 100000 - 1))) -->
+
+<!-- full_df_with_data %>%  -->
+<!--   ggplot() +  -->
+<!--   stat_density_2d(aes(x, y, fill = ..level..), geom = "polygon", alpha = 0.5) +  -->
+<!--   geom_point(aes(x_data, y_data)) + -->
+<!--   scale_fill_distiller(palette=1, direction=1) +  -->
+<!--   theme( -->
+<!--     legend.position = 'none', -->
+<!--     axis.title = element_blank(), -->
+<!--     axis.text = element_blank() -->
+<!--   ) -->
+
+<!-- ``` -->
+
+```{r}
+full_df_with_data <- full_df %>% 
+  mutate(x_data = c((df$x / 1.7), rep(0, (100000 - 1000)))) %>% 
+  mutate(y_data = c((df$y / 1.7), rep(0, (100000 - 1000))))
+
+full_df_with_data %>% 
+  ggplot() + 
+  stat_density_2d(aes(x, y, fill = ..level..), geom = "polygon", alpha = 0.5) + 
+  geom_point(aes(x_data, y_data)) +
+  scale_fill_distiller(palette=1, direction=1) + 
+  theme(
+    legend.position = 'none',
+    axis.title = element_blank(),
+    axis.text = element_blank()
+  )
+
+```
+
+```{r}
+full_df_with_data <- full_df %>% 
+  mutate(x_data = c((df$x / 1.7), rep(0, (100000 - 1000)))) %>% 
+  mutate(y_data = c((df$y / 1.7), rep(0, (100000 - 1000))))
+
+full_df_with_data %>% 
+  ggplot() + 
+  geom_point(aes(x_data, y_data)) +
+  theme(
+    legend.position = 'none',
+    axis.title = element_blank(),
+    axis.text = element_blank()
+  ) + 
+  theme_void()
+
+```

### 2020-05-07 working with total variation distance.R
update

@@ -1,5 +1,24 @@
+library(tidyverse); options(tibble.width = Inf)
+library(magrittr)
+
+
 library(distr)
 library(distrEx)
+
+
+true_dist <- Beta(3,9)
+x <- rbeta(10000, 3,9)
+true_dist <- Norm(0,1)
+x <- rnorm(100000, 0, 1)
+
+x_dist <- TotalVarDist(true_dist, x, asis.smooth.discretize = "asis")
+x_dist
+x_dist <- TotalVarDist(true_dist, x, asis.smooth.discretize = "discretize")
+x_dist
+x_dist <- TotalVarDist(x, true_dist, asis.smooth.discretize = "smooth")
+x_dist
+
+
 
 jags_dist <- TotalVarDist(true_dist,jags_fit_condense)
 bugs_dist <- TotalVarDist(true_dist,bugs_fit_condense)

### 2020-05-10 Simulation Results/Simulation Results.Rmd
ppp updates

@@
 knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
-library(tidyverse); options(tibble.width = Inf); theme_set(theme_bw())
+theme_set(theme_minimal())
 library(magrittr)
@@
 
-files <- logistic_regression_bench_files
+eight_schools_bench_files <- c(
+  here("Simulation Results", "8-schools", "eight-schools-results.rds"),
+  here("Simulation Results", "8-schools", "eight-schools-results-stan.rds")
+)
+
+ma_model_bench_files <- c(
+  here("Simulation Results", "time-series", "ma-model", "ma-model-results.rds"),
+  here("Simulation Results", "time-series", "ma-model", "ma-model-results-stan.rds")
+)
+
+ar_model_bench_files <- c(
+  here("Simulation Results", "time-series", "ar-model", "ar-model-results-no-nimble.rds"),
+  here("Simulation Results", "time-series", "ar-model", "ar-model-results-greta.rds"),
+  here("Simulation Results", "time-series", "ar-model", "ar-model-stan-compile-results.rds")
+)
+
+files <- eight_schools_bench_files
+
+model_name <- "8"
 
@@
   combine_bench_indiv("poisson regression")
+cox_exp_results <- cox_exponential_bench_files %>% 
+  combine_bench_indiv("cox exponential")
+eight_schools_results <- eight_schools_bench_files %>% 
+  combine_bench_indiv("eight schools")
+ma_model_results <- ma_model_bench_files %>% 
+  combine_bench_indiv("ma model")
+ar_model_results <- ar_model_bench_files %>% 
+  combine_bench_indiv("ar model")
 
@@
   logistic_reg_results,
-  poisson_reg_results
+  poisson_reg_results,
+  cox_exp_results,
+  eight_schools_results,
+  ma_model_results,
+  ar_model_results
 )
@@
 ```{r}
-bench_results_indiv <- readRDS(here("bench_results_indiv2.RDS"))
+bench_results_indiv <- readRDS(here("bench_results_indiv.RDS"))
 
@@
   ylab("Count Density")
+
+```
+
+## Strip Charts
+
+```{r}
+bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  ggplot(aes(method, time)) + 
+  geom_jitter(width = 0.1) + 
+  facet_wrap(vars(model)) + 
+  theme_update(panel.grid.minor = element_blank()) + 
+  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile")) + 
+  ylim(0,100)
+
+bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  # filter(time < 100) %>% 
+  ggplot(aes(method, time, color = method)) + 
+  geom_jitter(width = 0.1) + 
+  facet_wrap(vars(model), scales = "free_y") + 
+  theme_update(panel.grid.minor = element_blank()) + 
+  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile"))
+
+```
+
+- First slide of strip charts
+```{r}
+first_eight <- bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  group_by(model) %>% 
+  summarize(mean_time = mean(time)) %>% 
+  arrange(mean_time) %>% 
+  pull(model) %>% 
+  .[1:8] %>% 
+  sort()
+
+first_six <- bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  group_by(model) %>% 
+  summarize(mean_time = mean(time)) %>% 
+  arrange(mean_time) %>% 
+  pull(model) %>% 
+  .[1:6] %>% 
+  sort()
+
+bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  filter(model %in% first_eight) %>% 
+  filter(time < 100) %>% 
+  ggplot(aes(method, time, color = method)) + 
+  geom_jitter(width = 0.2, size = 1) + 
+  # facet_wrap(vars(model), scales = "free_y") + 
+  facet_wrap(vars(model)) + 
+  theme_update(panel.grid.minor = element_blank()) + 
+  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile"))
+
+bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  filter(model %in% first_six) %>% 
+  filter(time < 100) %>% 
+  ggplot(aes(method, time, color = method)) + 
+  geom_jitter(width = 0.2) + 
+  # facet_wrap(vars(model), scales = "free_y") + 
+  facet_wrap(vars(model)) + 
+  theme_update(panel.grid.minor = element_blank()) + 
+  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile"))
 
@@
 ```
+
+## Strip charts for accuracy
+
+```{r}
+time_results
+```

### 2020-05-10 Simulation Results/Simulation Results.Rmd

@@
 knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+library(tidyverse); options(tibble.width = Inf)
+library(magrittr)
 theme_set(theme_minimal())
-library(magrittr)
 library(here)
@@
 ```{r}
+readRDS(here("Simulation Results", "simple-models", "continuous", "exponential", "gamma-prior", "exponential-gamma-accuracy.rds"))
+
 exp_gamma_acc_files <- c(

### 2020-05-20 Paper.Rmd
Paper updates

@@
 knitr::opts_chunk$set(eval = FALSE)
+library(knitr)
+library(png)
 ```
@@
 
-In Bayesian statistical inference, we are interested in finding the posterior distribution of a certain parameter, given a prior distribution and a likelihood for that parameter from a sample of data.
-Bayes Rule states that, given a parameter of interest $\theta$ and some known data $x$, we can find the conditional probability of $\theta$ given $y$ using $$ p(\theta | x) = \frac{p(x | \theta) p(\theta)}{p(x)},$$ where in the discrete case $p(x) = \sum_\theta p(\theta) p(x|\theta),$ and in the continuous case, $p(x) = \int p(\theta) p(y|\theta) d\theta$.
-More formally, we can derive the posterior distribution of a parameter $\theta$, $p(\theta | x)$, by $$p(\theta | x_1, x_2, \dots, x_n) = \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}},$$ where $\ell(x | \theta)$ is proportional to the likelihood of $\theta$, conditioned on the data, and $\pi(\theta)$ is the prior distribution of $\theta$.
+In Bayesian statistical inference, we are interested in finding the posterior distribution of certain parameters, given a prior distribution and a likelihood for those parameters our data.
+Much of Bayesian theory is built upon the definition of conditional probability, which states that we can find the probability of a variable $\theta$, conditioned on $x$, by $$ p(\theta | x) = \frac{p(x | \theta) p(\theta)}{p(x)}.$$ In any Bayesian inference, the first step is to find the joint probability distribution of all the observed and unobserved quantities in the problem.
+Using the definition of conditional probability, we can then find the posterior distribution of the parameters of interest.
+Bayes theorem states that the posterior distribution of a parameter $\theta$ is equal to $$p(\theta | x_1, x_2, \dots, x_n) = \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}} = \frac{\ell(\theta | \textbf{x}) \pi(\theta)}{\int{\ell(\theta | \textbf{x}) \pi(\theta) d \theta}},$$ where $\ell(\theta | \textbf{x})$ is proportional to the likelihood of $\theta$, conditioned on the data, and $\pi(\theta)$ is the prior distribution of $\theta$.
+
 For example, suppose that $X_i \sim \text{Bern}(\theta)$, and we are interested in inference on $\theta$.
@@
 \begin{align*}
-  p(\theta | x) &= \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}} \\
+  p(\theta | x) &= \frac{\ell(\theta | \textbf{x}) \pi(\theta)}{\int{\ell(\theta | \textbf{x}) \pi(\theta) d \theta}} \\
   &= \frac{\theta ^ {\sum_{i=1}^{n} x_i} (1 - \theta) ^ {n - \sum_{i=1}^{n} x_i} \frac{\Gamma (\alpha + \beta)}{\Gamma(\alpha) + \Gamma(\beta)} \theta^{\alpha - 1} (1 - \theta)^{\beta - 1}}{\int \theta ^ {\sum_{i=1}^{n} x_i} (1 - \theta) ^ {n - \sum_{i=1}^{n} x_i} \frac{\Gamma (\alpha + \beta)}{\Gamma(\alpha) + \Gamma(\beta)} \theta^{\alpha - 1} (1 - \theta)^{\beta - 1} d \theta} \\
@@
 
-## Bayesian Computation
+One challenge in bayesian inference comes in evaluating the integral in the denominator of Bayes theorem, which is used to normalize the posterior distribution.
+If we don't have distributions for the likelihood and prior that are conjugate (or in other words, play nice together), then it is impossible to analytically calculate the posterior distribution.
 
-In more complicated scenarios, it is not easy to calculate the posterior distribution of a parameter by hand, and in many cases it is impossible.
-This is where computing can help us.
-If we could somehow computationally grab samples from the posterior distribution we are interested in, then we could obtain most of the statistics that we are interested in.
+## Bayesian Computational Methods
 
-### History of Bayesian computation up to MCMC
+The need for bayesian computational methods arises in our search for an alternative approach to finding the posterior if we cannot do so analytically.
+Most of the time, the goal of any bayesian analysis is to not to find the posterior distribution, but to calculate expectations with respect to the posterior.
+However, in order for these estimates to be accurate, we often need to know about the full behavior of that posterior distribution.
+In bayesian computing, the aim is to somehow obtain individual draws of values that would come from the posterior, and then aggregate a lot of them to compute the expectations we need.
 
@@
 
-Markov Chain Monte Carlo (referred to as MCMC), is a common approach that allows for drawing samples of a parameter from a distribution that is approximately equivalent to the target posterior distribution of that parameter.
-It is an iterative algorithm, where each realization gets closer to the target density we are interested in.
-An essential property of any Markov chain is that the distribution of each draw only depends on the most recent value drawn before it.
-As we keep obtaining draws, each draw is probabilisticly more likely to have come from the target posterior distribution.
-Based on the convergence properties of MCMC, we can expect each draw to be coming from the real distribution of the parameter we are interested, once we have let the Markov chain run long enough.
+Markov Chain Monte Carlo (referred to as MCMC), is a common approach that allows for drawing samples from a target distribution, which in our case is the posterior.
+MCMC is an iterative algorithm, where we obtain one draw at a time from the parameter space from which we are trying to sample.
+In any Markov chain, the marginal distribution of each draw is conditioned on the values of our chain from the previous draw.
+As we keep obtaining draws, these marginal distributions asymptotically converge to the true target distribution.
+Therefore, if we run the markove chain for many iterations, we can expect to see draws coming from the true posterior distribution we are trying to estimate.
+MCMC is the only suitable method for obtaining the posterior distribution when we don't have conjugate priors.
 
-- MCMC is the only suitable method when we don't have a conjugate prior.
-
-### Metropolis-Hastings
+### Metropolis-Hastings Algorithm
 
@@
 The beauty of the Gibbs sampling strategy is that, no matter how complex the space of the posterior distribution is, sampling from these high-dimensional spaces can be broken down into more straightforward draws from one-dimensional distributions.
-The downside of this approach is that, in high-dimensional settings, it may take a very long time to sample from each conditional one-dimensional distribution in order to reach convergence.
-Furthermore, because the Gibbs sampler uses a random-walk behavior, if several variables are highly correlated, traversing the space can take a very high amount of iterations because the step-size needs to be small in order to not step outside the target distribution.
+Gibbs sampling is also easier to understand and implement, which makes it very versatile.
+However, full conditional distributions for each draw are needed in order to do the method.
 
-### BUGS
+Because the Gibbs sampler is a random-walk algorithm, if several variables are highly correlated, traversing the space can take a very high amount of iterations because the step-size needs to be small in order to not step outside the target distribution.
+A two-dimensional example of this is seen in Figure 1, where the joint distribution of $\theta_1$ and $\theta_2$ is highlgy correlated.
+Since each Markov transition only allows us to move in the direction of one dimension at a time, it may take many steps in order to travel up and down the parameter space, as demonstrated by the red arrows.
+In reality, the movement of the chains will be much more random since the algorithm does not know which direction will allow the chain to most efficiently travel through the posterior.
 
-The BUGS project, which stands for Bayesian inference Using Gibbs Sampling, began in 1989 in an effort to create a program that utilized MCMC to approximate draws from the posterior distribution in a Bayesian framework.
-The BUGS language has functions and distributions that allow for specifying relationship between different "nodes" in a bayesian model.
-While the language itself has a somewhat limited syntax, it is able to describe a very diverse and complex set of models.
-The BUGS language is declarative, meaning that the order of statements in a model script does not matter since the entire script is evaluated at once, instead of line-by-line in a procedural language.
+```{r, echo = FALSE, eval = TRUE, out.width = "50%"}
+include_graphics("Paper-Figures/Gibbs1.png")
+include_graphics("Paper-Figures/Gibbs2.png")
+```
 
-Many different programs have been developed that utilize the BUGS syntax to accomplish bayesian sampling, and while these programs have continually evolved, the BUGS syntax has remained very consistent.
-The Windows program WinBUGS was created by the inventors of the BUGS language, and was a more popular choice for bayesian computation until it became deprecated in 2007.
-OpenBUGS is now more commonly used in place of WinBUGS, although it's last stable release was in 2014.
-OpenBUGS communicates better with other software than WinBUGS and is more portable.
-
-In general, there are six steps needed to run bayesian models in WinBUGS/OpenBUGS:
-
-1. Specify the model through BUGS syntax and check to make sure the model code is valid.
-2. Load the appropriate data.
-3. Compiling the BUGS model.
-4. Start the simulation at appropriate initial values for the Markov chains.
-5. Simulate draws from the posterior distributions of the unknown parameters of interest.
-6. Gather the results and report summary statistics, along with other appropriate output.
+![\textit{This is a demonstration of a Gibbs sampler in a highly correlated two-dimensional parameter space.
+On the left, the ideal behavior of the markov chain is shown, moving as quickly as possible from the bottom-left to the top-right to efficiently sample from the target distribution without stepping outside of it.
+On the right, we see the more realistic behavior of the markov chain, as the random-walk behavior of the Gibbs sampler moves the chain around in directions that aren't always preferred.}](Paper-Figures/Blank.png)
 
@@
 # Bayesian Samplers
+
+### BUGS
+
+The BUGS project, which stands for Bayesian inference Using Gibbs Sampling, began in 1989 in an effort to create a program that utilized MCMC to approximate draws from the posterior distribution in a Bayesian framework.
+The BUGS language has functions and distributions that allow for specifying relationship between different "nodes" in a bayesian model.
+While the language itself has a somewhat limited syntax, it is able to describe a very diverse and complex set of models.
+The BUGS language is declarative, meaning that the order of statements in a model script does not matter since the entire script is evaluated at once, instead of line-by-line in a procedural language.
+
+Many different programs have been developed that utilize the BUGS syntax to accomplish bayesian sampling, and while these programs have continually evolved, the BUGS syntax has remained very consistent.
+The Windows program WinBUGS was created by the inventors of the BUGS language, and was a more popular choice for bayesian computation until it became deprecated in 2007.
+OpenBUGS is now more commonly used in place of WinBUGS, although it's last stable release was in 2014.
+OpenBUGS communicates better with other software than WinBUGS and is more portable.
+
+In general, there are six steps needed to run bayesian models in WinBUGS/OpenBUGS:
+
+1. Specify the model through BUGS syntax and check to make sure the model code is valid.
+2. Load the appropriate data.
+3. Compiling the BUGS model.
+4. Start the simulation at appropriate initial values for the Markov chains.
+5. Simulate draws from the posterior distributions of the unknown parameters of interest.
+6. Gather the results and report summary statistics, along with other appropriate output.
 
@@
 
-\textit{\textbf{Gamma}}: The $Gam(\alpha,\lambda)$ parameterization, where $\alpha$ is a shape parameter and $\lambda$ is a rate parameter, is used by \texttt{OpenBUGS}, \texttt{JAGS}, \texttt{Nimble}, and \texttt{greta}.
-The $Gam(\alpha,\beta)$ parameterization is used by \texttt{STAN}, using the scale parameter $\beta = 1 / \lambda$.
+\textit{\textbf{Gamma}}: The $Gam(\alpha,\lambda)$ parameterization, where $\alpha$ is a shape parameter and $\lambda$ is a rate parameter, is used by all five implementations.
 

### 2020-08-13 Paper.Rmd
added paragraph on Metropolis Hastings

@@
 
+When using Markov Chain methods to sample from the Bayesian posterior distribution, the Metropolis-Hastings Algorithm gives a general approach to construct transitions from one draw to the next.
+The algorithm has two steps: a proposal step, where we suggest a new draw, and a correction step, where we try to reject any proposals that move too far away from the typical set from which we are trying to sample (Betancourt paper).
+
+A probability density for the proposal is needed, which we will call $Q(q'|q)$, where $q$ is our current point, and $q'$ is the proposed point.
+After sampling from the proposal distribution, the probability of accepting this proposal is calculated as
+
+$$a(q'|q) = \min \left( 1, \frac{Q(q'|q) \pi(q')}{Q(q|q') \pi(q)} \right).$$
+
+Anytime that a proposal moves closer to the typical set, it will automatically be accepted.
+If a proposal moves further away, it will be accepted with probability $\frac{Q(q'|q) \pi(q')}{Q(q|q') \pi(q)}$.
+Choosing a proposal distribution is important.
+The distribution should be easy to sample from, for any $q$, and should also allow for easy computation of the acceptance probability "ratio".
+Furthermore, the distribution should strike a balance between each proposed jump traveling a reasonable distance from the current point, while not having proposals rejected too frequently (BDA3).
+
+A gaussian distribution, such as $Q(q'|q) = N(q'|q, \Sigma)$ is often used for the proposal, including in Gibbs Sampling methods.
+In this case, because the distribution is symmetric and $Q(q'|q) = Q(q|q')$, then the acceptance probability simplifies to
+
+$$a(q'|q) = \min \left( 1, \frac{\pi(q')}{\pi(q)} \right).$$
+
+There are some issues that arise with the Metropolis-Hastings algorithm.
+Each proposal is biased towards larger volumes (such as the tails of the distribution).
+Although the algorithm will reject these proposals because the density is smaller, this can lead to many rejected proposals.
+This is especially true in higher dimensions, where in some instances, virtually every proposal step will be rejected, and even finishing one "journey" across the typical set might be unattainable.
+In better scenarios, there is still often autocorrelation in the draws that has to be accounted for.
+
 ### Gibbs Sampling
@@
 
-A main of objective of Hamiltonian Monte Carlo is to use the geometry of the typical set to help us know where to sample from, instead of blindly using a guess-and-check method (Betancourt HMC Youtube talk).
-This method utilizes Hamiltonian Dynamics, which is a physics system that has three elements: position, momentum, and potential energy variables.
-To use Hamiltonian Dynamics to traverse the typical set we are interested in, each parameters of interest is transformed into a position variable, which can be visualized as a frictionless puck sliding over a surface of varying height.
-Each position variable $p$ has an associated momentum variable $q$ that defines the direction and magnitude of the movement of the "puck".
-At any given time, the value of the parameter can be found as the position of $p$ in some dimension.
+A main of objective of Hamiltonian Monte Carlo methods is to use our knowledge of the geometry of the target distribution to guide our sampling efforts, instead of utilizing a random-walk algorithm (Betancourt HMC Youtube talk).
+The density function of the target distribution is translated into a potential energy function, which is used in Hamiltonian Dynamics, a physics system that we employ in order to yield monte carlo draws.
+To use this dynamic system to obtain draws, each parameter is transformed into a position variable, where the position of the variable can be interpreted as the value of the parameter at that point in time.
+Each position variable $q$ also has one associated momentum variable $p$ which will guide the movement of the position variable inside the simulation.
 
-We define the joint probability of the position $p$ and momentum $q$ as $\pi(q,p) = \pi(q|p) \pi(p)$.
-This structure allows us at anytime to throw out $q$ and recover the value of $p$, which holds the position values of the parameter(s).
-The Hamiltonian is defined as
+The joint probability distribution of the position $p$ and momentum $q$ is $\pi(q,p) = \pi(p|q) \pi(q)$.
+Although both $q$ and $p$ are being used jointly to determine movement, we can find the marginal distribution $\pi(q)$ at any time to recover the values of the position variables.
+
+The Hamiltonian is a set of equations defined as
 
@@
 \end{align*}
-where $K$ is potential energy and $V$ is kinetic energy. Taking partial derivatives of the Hamiltonian yields Hamilton's equations, which is where we obtain our vector field.
+where $K$ is potential energy and $V$ is kinetic energy.
+Taking partial derivatives of the Hamiltonian yields Hamilton's equations, which is where we obtain our vector field that guides the movement of the position variables:
 
@@
 
-Hamiltonian Dynamics can be used to sample from the target distribution by translating the density function of the distribution to a potential energy function, and obtaining position variables (and associated momentum variables) for parameters of interest.
-We can simulate a markov chain where each iteration consists of the following steps:
+We can simulate a Hamiltonian monte carlo markov chain where each iteration consists of the following steps:
 
@@
 
-Hamiltonian Monte Carlo methods are very robust, in that the markov chains seldom get caught in complex regions of the posterior compared to other common MCMC methods.
-When the chains do get caught, the HMC chains break so obviously that errors can be easily spotted and reported to the user.
+```{r, echo = FALSE, eval = TRUE, out.width = "50%", fig.align='center'}
+include_graphics("Paper-Figures/HMC projecting down.png")
+```
 
-### STAN
-STAN 2015
-- Based on a probabilistic programming language that is more flexible and expressive than BUGS/JAGS.
-- Based on HMC
-- Works in command line, Python, and R
-- "STAN is a platform for Bayesian inference that tries to abstract as much computation away as possible, so users can focus on building complex models"
+![\textit{A visual representation from Betancourt of lifting up off the parameter space (illustrated in red) to a higher-dimension phase space to sample when using HMC.}](Paper-Figures/Blank.png)
+
+While the probability distribution for the position variables is defined by the target distribution, the choice of the conditional distribution the momentum, which is equivalent the kinetic energy function, is up to the user.
+This choice is very important as it can drastically affect the simulation.
+There are several common gaussian distributions used for these distributions, which are used by default in the implementations that utilize HMC methods.
+
+Hamiltonian Monte Carlo methods are very robust, in that the markov chains get "stuck" in complex regions of the posterior less frequently than other common MCMC methods.
+When the markov chains do get stuck, it is easy for the program running the sampler to spot and report to the user.
 
@@
 
-OpenBugs is the oldest of the Bayesian samplers that we use.
-OpenBugs is designed for Windows machines, but there is a workaround to run it on MacOS and Linux, using WINE, a program designed to run Windows software on other operating systems.
-OpenBugs can be used through the actual click-and-point application, or through \texttt{R}, using the \texttt{R2OpenBugs} package, which allows for the user to send model scripts and data to OpenBugs without needing to leave the \texttt{R} interface.
-The \texttt{R} command to run OpenBugs is \texttt{bugs()}.
-Using this method causes OpenBugs to run invisibly in the background and return the results, unless an argument is specified to debug code in the actual application itself, which can be very useful.
+OpenBUGS is the oldest of the Bayesian samplers that we use, maintained only as recently as 2014.
+OpenBugs is a program for Windows machines, but there are workarounds to run it on MacOS and Linux, such as WINE or a virtual machine.
+OpenBUGS can be used through the actual click-and-point application, or through \texttt{R}, using the \texttt{R2OpenBugs} package, which allows for the user to send model scripts and data to OpenBugs without needing to leave the \texttt{R} interface.
+OpenBUGS is based on the BUGS model specification language, which is a small-to-moderately sized set of functions and commands that allow for extensive specification of bayesian models.
+Several other common bayesian sampling implementations are made to closely mimic the BUGS language as well.
+
+<!-- Using this method causes OpenBugs to run invisibly in the background and return the results, unless an argument is specified to debug code in the actual application itself, which can be very useful.
 When using R to run OpenBugs, there is little way to know what went wrong if the code errors somewhere without interacting with the OpenBugs program itself.
@@
 It is recommend to use \texttt{debug = T} only when debugging, as it does slow down the sampling process slightly.
+-->
+
 OpenBugs has many tools to visualize the performance the sampler which are easy to use.
@@
 One unique feature in OpenBugs is the Doodle tool, which allows for model specification through the creation of a diagram of nodes representing the parameters.
-Though using Doodle can be helpful for illustration, it is not used often in practice.
+Though using Doodle can be helpful for illustration, it is not used often in daily practice.
+
+```{r, echo = FALSE, eval = TRUE, out.width = "50%", fig.align='center'}
+include_graphics("Paper-Figures/Doodle.png")
+```
+
+![\textit{An example of the OpenBUGS visual tool DoodleBUGS.}](Paper-Figures/Blank.png)
 
@@
 
-In terms of model specification, JAGS is very similar to OpenBugs, but works more seemlessly in R using the package \texttt{rjags}.
-JAGS was developed independently of OpenBUGS, but uses the BUGS language for most of its model specification syntax.
+In terms of model specification, JAGS is very similar to OpenBugs, as it is based on the BUGS model specification language.
+However, JAGS is always interacted with entirely in R, making the bayesian inference process more seemless, using the R package \texttt{rjags}.
 The \texttt{runjags} is an optional package that allows for the JAGS model script to be inputted as a string, and run using the \texttt{run.jags()} function.
-Since JAGS runs inside \texttt{R} itself, it is very easy to work with.
-Helpful errors are given when there is an issue with the model code or sampling process.
-JAGS is often a very fast sampler, especially for less complicated models.
+Built on C++, JAGS is able to run MCMC sampling process very quickly.
+There is plenty of documentation and support available for JAGS, which is still maintained.
 
@@
 
-Nimble is an extension of Bugs that allows for bayesian computation within \texttt{R}, but uses C++ under the hood to compile models and algorithms for speed.
-Using the \texttt{nimble} package, Nimble can perform MCMC bayesian computations, but also allows for other types of algorithms to be loaded or custom-built.
+Nimble allows for bayesian computation within \texttt{R}, like JAGS and OpenBUGS, but is also made for more flexible statistical modeling.
+It uses C++ under the hood to compile models and algorithms.
+Using the \texttt{nimble} package, Nimble can perform MCMC bayesian sampling using common algorithms, but also allows for other types of algorithms to be loaded or custom-built.
 It can also compile other code without BUGS models into C++ for other uses.
-For the purpose of bayesian sampling using MCMC, Nimble is used very similarly to BUGS in practicality.
-The flexibility of Nimble is not needed in most cases when doing MCMC.
-The model specification language is almost identical to OpenBugs.
-A model can be both compiled and run using the function \texttt{nimbleMCMC()}.
-Because the model is compiled using C++ each time, this can drastically increase the total computation time as compared to other sampling methods.
-Occasionally, Nimble can also freeze up or fail, which can take a bit of time to fix.
+For the purpose of bayesian sampling using MCMC, Nimble is used very similarly to JAGS and OpenBUGS in practicality, as its model specification language is similar.
+The flexibility of Nimble is not needed in most cases when doing bayesian MCMC methods.
+The default sampling method in Nimble is Gibbs sampling, but other methods can be used as well.
+Because the model has to be compiled first for C++ before sampling each time, this can drastically increase the total computation time as compared to other sampling methods.
 
@@
 STAN can be used in several programming languages, including \texttt{R} and Python.
-While many bayesian samplers use Gibbs sampling, STAN uses Hamiltonian Monte Carlo under the hood, a more computationally efficient sampler that is based on an advanced physics simulation.
-This can greatly reduce the computational time for many types of statistical models, especially for more complicated ones.
-Like Nimble, STAN compiles each model, but one big advantage is that it can store each compiled model in a .Rds file, which allows for repeated use without the need to re-compile each time.
+STAN uses Hamiltonian Monte Carlo sampling methods, which ideally can reduce the computational time for many types of statistical models.
+Like Nimble, STAN compiles each model for C++, but one convenient feature is that the user can store each compiled model in a file, which allows for repeated use without the need to re-compile each time.
 In \texttt{R}, STAN is used through the \texttt{rstan} package, which allows for data to be loaded through normal \texttt{R} data structures.
 The biggest practical difference in using STAN compared to other bayesian computational methods is in the model specification process.
-STAN requires a model object in the form of a .stan file, which is created in a very unique language that is more formal than in BUGS.
+STAN requires a model object in the form of a .stan file, which is created in a very unique language that is more formal than BUGS syntax.
 A .stan file, either created through a character string in R or in a seperate text file, can be used by the \texttt{stan()} function, which takes the file along with the data to perform the sampling.
-There is a very robust set of manuals for STAN users, compared to other methods.
+There is a very robust set of manuals for STAN users, including many papers written on the subject.
 
@@
 
-Greta, a very recent addition to the list of bayes computational methods, is made specifically for use in \texttt{R}, making the model and data specification process as easy as possible.
-Not only does it utilize Hamiltonian Monte Carlo methods for sampling, but it also uses Google TensorFlow to perform its calculations, and allows for more flexible use of CPUs and GPUs.
-Unfortunately, as of the writing of this, it is still in need of lots of development, and lacks an appropriate amount of documentation and support for users.
-Computationally, it is very slow compared to the other bayes methods, especially in simple cases.
-There are several types of semi-common models that greta seems unable to properly handle in model specfication at the moment.
-Greta also attempts to make model specification very user-friendly, but uses some very common function names in the process that might override other often-used \texttt{R} functions.
-Greta does offer a model visualization tool that does not exist in any of the other Bayesian samplers, in that it can provide a graphical representation of the specified model through the use of nodes and arrows.
+Greta, a more recent addition to the list of bayesian computational implementations, is made specifically for easy use in \texttt{R}.
+The model and data specification process is intended to be as straightforward as possible users who are familiar with \texttt{R} but not as familiar with other bayes methods.
+Not only does it utilize Hamiltonian Monte Carlo methods for sampling, but it also uses Google TensorFlow,a machine learning platform from Google, to perform its calculations.
+This allows for more flexible use of CPUs and GPUs.
+Greta offers a model visualization tool that does not exist in any of the other Bayesian samplers, in that it can provide a graphical representation of the specified model through the use of nodes and arrows.
 This tool can be very useful in trying to explain how all the components of a model work together.
+
+Unfortunately, as of the writing of this, Greta is still in need of lots of development, and lacks an appropriate amount of documentation and support for users, compared to the other implementations mentioned previously.
+The sampling process can slow compared to the other bayes methods when using identical sampling configurations, especially for simple models.
+There are also several types of models that Greta seems unable to properly specify at the moment.
 
@@
 
-To illustrate how each bayesian computational method is used in practice, we will demonstrate how to construct a sampler for a simple Beta-Binomial model for each of the five methods we are studying.
+To illustrate how each bayesian computational method is used in practice, we will demonstrate how to construct a model and provide data for a simple Beta-Binomial model for each of the five implementations.
 
@@
 
+An extra step is required when using Nimble, which requires for any constant variables to be seperated from other data variables.
+
+```{r}
+model_data <- list(
+  "y" = y
+)
+
+nimble_constants <- list(
+  "n" = n 
+)
+```
+
 ## Model Configuration
 
-Next, we will pick some basic model parameters. Here, we will use four chains, a warmup of 1,000 iterations, and a total iteration count of 10,000.
+Next, we will pick some basic bayesian sampling parameters.
+Here, we will use four markov chains, an initial warmup (or burnin, depending on the sampling method) of 1,000 iterations, and a total iteration count of 10,000.
 
@@
 n_iter <- 1e4L
-n_warmup <- 1e3L
+n_initial <- 1e3L
 ```
 
-## OpenBUGS / R2OpenBUGS
+## Model Specification
+
+For several implementations, the model script code can be specified using a seperate text file that is read by the function that compiles and runs the sampler.
+There are also ways to input the model script as an R object itself.
+For each implementation, if possible, we will demonstrate how to use both methods of model specification.
+
+## OpenBUGS
 
 When using OpenBUGS through R, we will use the `R2OpenBUGS` library in order to pipe the data and model code into the program rather than using manual entry.
-This requires the creation of a temporary file that holds the model text that will be used.
+One way to specify the model code is by creating a text file that will be read into R when compiling and running the sampler.
+We will create the text file `model.bugs` with the BUGS model script code inside.
+
+```{txt}
+## model.bugs
+
+model {
+  y ~ dbin(p,n)
+  p ~ dbeta(1,1)
+}
+
+##
+```
+
+In the R script, we will load the `R2OpenBUGS` library, specify a vector of parameters to monitor (which in this case has just one parameter), and run the sampler using the `bugs()` function.
+As is the case with all bayesian sampling implementations, the function call needs the model, data, parameters, and model configuration settings specified.
+The `"inits" = NULL` argument tells OpenBUGS to randomly generate initial values for the parameters based on the prior distributions for those parameters, instead of the user having to provide them.
 
@@
 
-## model script code
+## parameters to monitor
+bugs_monitor <- c("p")
+
+## run the model
+bugs_fit <- bugs(
+  "model.file" = "model.bugs", 
+  "data" = model_data, 
+  "parameters.to.save" = bugs_monitor, 
+  "inits" = NULL, 
+  "n.chains" = n_chains, 
+  "n.iter" = n_iter, 
+  "n.burnin" = n_initial
+)
+```
+
+The alternative approach to model specification when using `R2OpenBUGS` is to provide the model code entirely through R.
+The user can wrap the model script code inside a function, and then create a temporary text file that will be read into the `bugs()` function.
+
+```{r}
+## model script code in a function
 bugs_model <- function() {
@@
 
-## parameters to monitor
-bugs_monitor <- "p"
-
 ## temporary model script text file creation
-bugs.file <- file.path(tempdir(), "model.txt")
+model.bugs <- file.path(tempdir(), "model.txt")
 write.model(bugs_model, bugs.file)
-
-## run the model
-bugs_fit <- bugs(
-    "model.file" = bugs.file, "data" = model_data, "parameters.to.save" = bugs_monitor, 
-    "inits" = NULL, "n.chains" = n_chains, "n.iter" = n_iter, "n.burnin" = n_warmup
-)
 ```
@@
 
-In JAGS, the model script simply needs to be inputted as a string to the `run.jags()` function.
+Model specification in JAGS is very similar to OpenBUGS in that we can either create a text file, or specify the model through R.
+Since the specification language for JAGS is based on the BUGS language, the code will often look very similar.
+
+```{txt}
+## model.jags
+
+model {
+  y ~ dbin(p,n)
+  p ~ dbeta(1,1)
+}
+
+##
+```
+
+In R, the model script file location is inputted in the `"model"` argument.
+
 ```{r}
@@
 library("rjags"); library("runjags")
-
-## model script code
-jags_model <- "
-  model{
-    y ~ dbin(p,n)
-    p ~ dbeta(1,1)
-  }
-"
 
@@
 jags_fit <- run.jags(
-    "model" = jags_model, "data" = model_data, "monitor" = jags_monitor, 
-    "n.chains" = n_chains, "sample" = n_iter, "burnin" = n_warmup
+  "model" = "model.jags", 
+  "data" = model_data, 
+  "monitor" = jags_monitor, 
+  "n.chains" = n_chains, 
+  "sample" = n_iter, 
+  "burnin" = n_warmup
 ) 
+```
+
+To create the model script inside R, we can wrap the script in quotes to make it a character object.
+
+```{r}
+## model script code
+model.jags <- "
+  model{
+    y ~ dbin(p,n)
+    p ~ dbeta(1,1)
+  }
+"
 ```
@@
 
-An additional step needed when using Nimble, which is to specify initial values for all parameters of interest, since the sampler cannot generate initial values on its own.
+The only way to specify a model when using Nimble is to use the `NimbleCode()` function and wrap the code inside using curly braces.
+Initial values can't be generated automatically and must be specified for all parameters of interest.
+
 ```{r}
@@
 nimble_fit <- nimbleMCMC(
-    "code" = nimble_model, "data" = model_data, 
-    "inits" = nimble_inits, "monitors" = nimble_monitor, "nchains" = n_chains, 
-    "niter" = n_iter, "nburnin" = n_warmup, "summary" = TRUE
+  "code" = nimble_model,
+  "data" = model_data,
+  "inits" = nimble_inits,
+  "monitors" = nimble_monitor,
+  "nchains" = n_chains,
+  "niter" = n_iter,
+  "nburnin" = n_warmup,
+  "summary" = TRUE
 )
@@
 
-Here is the .stan file for this example, which is saved as "stan_file.stan". Each STAN file contains a data block, parameters block, and model block at a minimum.
-```{r}
+Here is the .stan file for this example, which is saved as "stan_file.stan".
+Each STAN file contains a data block, parameters block, and model block at a minimum.
+The data block initializes each data variable, the parameters block initializes any parameters to monitor, and the model block contains more familiar looking model script code.
+
+```{txt}
+## model.stan
+
 data {
@@
 }
+
+##
 ```
@@
 stan_fit <- stan(
-    "file" = "stan_file.stan", "data" = model_data, 
-    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
+  "file" = "model.stan",
+  "data" = model_data,
+  "chains" = n_chains,
+  "iter" = n_iter,
+  "warmup" = n_warmup
 )
@@
 
-Each data variable needs to be converted to a greta array before it can be used in the model.
+Model specification in Greta is more intuitive for R users, but requires the extra step of converting existing data objects into Greta arrays, using the `as_data()` function from the `greta` package. 
 ```{r}
@@
 Next, the priors for each parameter of interested are defined using functions that describe the prior distribution paramter values.
-The distribution of the response variable given in the data is then described using the `distribution()` function.
-The model is compiled using the `model()` function, before being sampled by `mcmc()`.
+The relationship between the data and parameters is defined using the  `distribution()` function.
+The model is compiled using the `model()` function, which requires a vector of all the parameters to model.
+Finally, sampling is done by calling `mcmc()`.
 
@@
 ## compile model and run
-greta_model <- model(p)
+greta_model <- model(c(p))
 greta_fit <- mcmc(
-    "model" = greta_model, "n_samples" = n_iter,
-    "warmup" = n_warmup,"chains" = n_chains
+  "model" = greta_model,
+  "n_samples" = n_iter,
+  "warmup" = n_warmup,
+  "chains" = n_chains
 )

### 2021-04-28 Paper.Rmd
add updated paper and simulations

@@
 ---
-title: "Bayes Computational Methods"
+title: "Bayesian Computational Strategies and Implementations"
 author: "Evan Miyakawa"
@@
 
+\textcolor{red}{Introductory paragraph.}
+
 ## The Bayes Paradigm
 
-In Bayesian statistical inference, we are interested in finding the posterior distribution of certain parameters, given a prior distribution and a likelihood for those parameters our data.
+In Bayesian statistical inference, we are interested in finding the posterior probability distribution of certain parameters, given a prior distribution and a likelihood for those parameters in our data.
 Much of Bayesian theory is built upon the definition of conditional probability, which states that we can find the probability of a variable $\theta$, conditioned on $x$, by $$ p(\theta | x) = \frac{p(x | \theta) p(\theta)}{p(x)}.$$ In any Bayesian inference, the first step is to find the joint probability distribution of all the observed and unobserved quantities in the problem.
 Using the definition of conditional probability, we can then find the posterior distribution of the parameters of interest.
-Bayes theorem states that the posterior distribution of a parameter $\theta$ is equal to $$p(\theta | x_1, x_2, \dots, x_n) = \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}} = \frac{\ell(\theta | \textbf{x}) \pi(\theta)}{\int{\ell(\theta | \textbf{x}) \pi(\theta) d \theta}},$$ where $\ell(\theta | \textbf{x})$ is proportional to the likelihood of $\theta$, conditioned on the data, and $\pi(\theta)$ is the prior distribution of $\theta$.
+Bayes theorem states that the posterior distribution of a parameter $\theta$ is equal to $$p(\theta | \textbf{x}) = \frac{\ell(\theta | \textbf{x}) \pi(\theta)}{\int{\ell(\theta | \textbf{x}) \pi(\theta) d \theta}},$$ where $\ell(\theta | \textbf{x})$ is proportional to the likelihood of $\theta$, conditioned on the data $\textbf{x}$, and $\pi(\theta)$ is the prior distribution of $\theta$.
+
+<!-- Bayes theorem states that the posterior distribution of a parameter $\theta$ is equal to $$p(\theta | x_1, x_2, \dots, x_n) = \frac{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta)}{\int{\ell(\theta | x_1, x_2, \dots, x_n) \pi(\theta) d \theta}} = \frac{\ell(\theta | \textbf{x}) \pi(\theta)}{\int{\ell(\theta | \textbf{x}) \pi(\theta) d \theta}},$$ where $\ell(\theta | \textbf{x})$ is proportional to the likelihood of $\theta$, conditioned on the data, and $\pi(\theta)$ is the prior distribution of $\theta$.
+-->
 
 For example, suppose that $X_i \sim \text{Bern}(\theta)$, and we are interested in inference on $\theta$.
-A reasonable prior to assign to $\theta$ would be a $\text{Beta}(\alpha, \beta)$, distribution, given that $\theta$ is constrained to be between 0 and 1, which is a property of the Beta distribution.
+A common prior distribution for $\theta$ would be a $\text{Beta}(\alpha, \beta)$, given that $\theta$ is constrained to be between 0 and 1, which is a property of the Beta distribution.
 Therefore, given that $\theta \sim \text{Beta}(\alpha, \beta)$ for some known values of $\alpha$ and $\beta$, then
@@
 \end{align*}
-which is the kernel of a $\text{Beta}\left(\sum (x_i) + \alpha, n - \sum (x_i) + \beta \right)$ distribution.
+which is the kernel of a $\text{Beta} \left( \sum (x_i) + \alpha, n - \sum (x_i) + \beta \right)$ distribution.
 Therefore, we can analytically calculate the posterior distribution of $\theta$ as $$p(\theta | x) \sim \text{Beta}\left(\sum (x_i) + \alpha, n - \sum (x_i) + \beta\right).$$
@@
 However, in order for these estimates to be accurate, we often need to know about the full behavior of that posterior distribution.
-In bayesian computing, the aim is to somehow obtain individual draws of values that would come from the posterior, and then aggregate a lot of them to compute the expectations we need.
+In bayesian computing, the aim is to obtain individual draws of values that would come from the posterior, and then aggregate a large number of them to compute the expectations we need.
 
@@
 
-Markov Chain Monte Carlo (referred to as MCMC), is a common approach that allows for drawing samples from a target distribution, which in our case is the posterior.
+Approaches using Markov Chain Monte Carlo methods (referred to as MCMC), are commonly used to drawing samples from a target distribution, which in our case is the posterior distribution of our parameter(s).
 MCMC is an iterative algorithm, where we obtain one draw at a time from the parameter space from which we are trying to sample.
@@
 As we keep obtaining draws, these marginal distributions asymptotically converge to the true target distribution.
-Therefore, if we run the markove chain for many iterations, we can expect to see draws coming from the true posterior distribution we are trying to estimate.
+Therefore, if we run the markov chain for many iterations, we can expect to see draws coming from the true posterior distribution we are trying to estimate.
 MCMC is the only suitable method for obtaining the posterior distribution when we don't have conjugate priors.
@@
 
-When using Markov Chain methods to sample from the Bayesian posterior distribution, the Metropolis-Hastings Algorithm gives a general approach to construct transitions from one draw to the next.
-The algorithm has two steps: a proposal step, where we suggest a new draw, and a correction step, where we try to reject any proposals that move too far away from the typical set from which we are trying to sample (Betancourt paper).
+When using Markov Chain methods to sample from a posterior distribution, the Metropolis-Hastings Algorithm gives a general approach to construct transitions from one draw to the next.
+The algorithm has two steps: a proposal step, where we suggest a new draw, and a correction step, where we potentially reject a proposal that move too far away from the typical set from which we are trying to sample (Betancourt paper).
 
@@
 Although the algorithm will reject these proposals because the density is smaller, this can lead to many rejected proposals.
-This is especially true in higher dimensions, where in some instances, virtually every proposal step will be rejected, and even finishing one "journey" across the typical set might be unattainable.
-In better scenarios, there is still often autocorrelation in the draws that has to be accounted for.
+This is especially true in higher dimensions, where in some instances, virtually every proposal step will be rejected, and finishing just one "journey" across the typical set might be unattainable.
+In simpler scenarios, there is still often autocorrelation in the draws for which we must account (Betancourt).
 
@@
 
-Gibbs Sampling is one of the most common Markov Chain Monte Carlo Algorithms used for Bayesian computation.
+Gibbs Sampling is one of the most common Markov Chain Monte Carlo algorithms used for Bayesian computation.
 A special case of the Metropolis-Hastings algorithm, a Gibbs sampler works by generating a multi-dimensional markov chain, with one dimension for each unknown variable of interest.
@@
     - Continue this process for the remainder of the $k$ elements of $\theta$.
-3. Repeat step 2 for the total decided number of iterations that the sampler will run.
+3. Repeat step 2 for the total decided number of iterations that the sampler will run (BDA3).
 
@@
 
-A main of objective of Hamiltonian Monte Carlo methods is to use our knowledge of the geometry of the target distribution to guide our sampling efforts, instead of utilizing a random-walk algorithm (Betancourt HMC Youtube talk).
-The density function of the target distribution is translated into a potential energy function, which is used in Hamiltonian Dynamics, a physics system that we employ in order to yield monte carlo draws.
+The main of objective of Hamiltonian Monte Carlo (HMC) algorithms is to use the knowledge of the geometry of the target distribution to guide sampling efforts, instead of using a random-walk algorithm (Betancourt HMC Youtube talk).
+Under HMC, the density function of the target distribution is translated into a potential energy function, which is used in Hamiltonian Dynamics, a physics system that is employed to yield Monte Carlo draws.
 To use this dynamic system to obtain draws, each parameter is transformed into a position variable, where the position of the variable can be interpreted as the value of the parameter at that point in time.
@@
 
-We can simulate a Hamiltonian monte carlo markov chain where each iteration consists of the following steps:
+We can simulate a Hamiltonian Monte Carlo Markov chain where each iteration consists of the following steps:
 
@@
 
-![\textit{A visual representation from Betancourt of lifting up off the parameter space (illustrated in red) to a higher-dimension phase space to sample when using HMC.}](Paper-Figures/Blank.png)
+![\textit{A visual representation from Betancourt of lifting up off the parameter space (illustrated in red) to a higher-dimension phase space to sample when using HMC (Betancourt).}](Paper-Figures/Blank.png)
 
-While the probability distribution for the position variables is defined by the target distribution, the choice of the conditional distribution the momentum, which is equivalent the kinetic energy function, is up to the user.
+While the probability distribution for the position variables is defined by the target distribution, the choice of the conditional distribution of the momentum, which is equivalent the kinetic energy function, is up to the user.
 This choice is very important as it can drastically affect the simulation.
-There are several common gaussian distributions used for these distributions, which are used by default in the implementations that utilize HMC methods.
+There are several common gaussian distributions used for these distributions, which are used by default in the implementations that employ HMC methods (Betancourt).
 
-Hamiltonian Monte Carlo methods are very robust, in that the markov chains get "stuck" in complex regions of the posterior less frequently than other common MCMC methods.
-When the markov chains do get stuck, it is easy for the program running the sampler to spot and report to the user.
+Hamiltonian Monte Carlo algorithms are very robust, in that the markov chains get "stuck" in complex regions of the posterior less frequently than other common MCMC methods.
+When the Markov chains do get stuck, it is easy for the program running the sampler to spot the issues and report them to the user.
 
-# Bayesian Samplers
+There are some inherent challenges that come with HMC methods that are worth mentioning.
+In every transition step between MCMC draws, a system of differential equations needs to be solved in order to direct the movement of the position variables.
+This can be difficult in some situations.
+Additionally, getting the "physics simulation" set up properly for a given model and data set requires several tuning parameters to be set correctly.
+While some of these parameters can be picked automatically through an adaptation phase before sampling, the user needs to have some knowledge of how to set these initially.
+This is less of an issue for the no-U-turn sampler, an HMC algorithm that can automatically tune the parameters in a warmup and adaptation phase that occurs before sampling (Gelman, Intro to Stan).
 
-### BUGS
+# Platforms for Bayesian Sampling and Inference
 
-The BUGS project, which stands for Bayesian inference Using Gibbs Sampling, began in 1989 in an effort to create a program that utilized MCMC to approximate draws from the posterior distribution in a Bayesian framework.
+The term "probabilistic programming language" refers to programming languages and tools that natively work with probabilistic models and do inference on them more generally (Hardesty).
+Within bayesian computing, there are several platforms that are used to do MCMC sampling, to great effect.
+A bayesian computing platform usually consists of a specification language used to write and program the models, and the engine that performs the computations.
+We have chosen several of the most common of these platforms to look at more in detail in this paper, with the objective of portraying the strengths, weaknesses, and unique features of each.
+
+<!-- ### BUGS -->
+
+<!-- The BUGS project, which stands for Bayesian inference Using Gibbs Sampling, began in 1989 in an effort to create a program that utilized MCMC to approximate draws from the posterior distribution in a Bayesian framework.
 The BUGS language has functions and distributions that allow for specifying relationship between different "nodes" in a bayesian model.
@@
 The BUGS language is declarative, meaning that the order of statements in a model script does not matter since the entire script is evaluated at once, instead of line-by-line in a procedural language.
+-->
 
-Many different programs have been developed that utilize the BUGS syntax to accomplish bayesian sampling, and while these programs have continually evolved, the BUGS syntax has remained very consistent.
+<!-- Many different programs have been developed that utilize the BUGS syntax to accomplish bayesian sampling, and while these programs have continually evolved, the BUGS syntax has remained very consistent.
 The Windows program WinBUGS was created by the inventors of the BUGS language, and was a more popular choice for bayesian computation until it became deprecated in 2007.
@@
 OpenBUGS communicates better with other software than WinBUGS and is more portable.
+-->
 
-In general, there are six steps needed to run bayesian models in WinBUGS/OpenBUGS:
+<!-- In general, there are six steps needed to run bayesian models in WinBUGS/OpenBUGS: -->
 
-1. Specify the model through BUGS syntax and check to make sure the model code is valid.
-2. Load the appropriate data.
-3. Compiling the BUGS model.
-4. Start the simulation at appropriate initial values for the Markov chains.
-5. Simulate draws from the posterior distributions of the unknown parameters of interest.
-6. Gather the results and report summary statistics, along with other appropriate output.
+<!-- 1. Specify the model through BUGS syntax and check to make sure the model code is valid. -->
+<!-- 2. Load the appropriate data. -->
+<!-- 3. Compiling the BUGS model. -->
+<!-- 4. Start the simulation at appropriate initial values for the Markov chains. -->
+<!-- 5. Simulate draws from the posterior distributions of the unknown parameters of interest. -->
+<!-- 6. Gather the results and report summary statistics, along with other appropriate output. -->
 
@@
 
-OpenBUGS is the oldest of the Bayesian samplers that we use, maintained only as recently as 2014.
-OpenBugs is a program for Windows machines, but there are workarounds to run it on MacOS and Linux, such as WINE or a virtual machine.
-OpenBUGS can be used through the actual click-and-point application, or through \texttt{R}, using the \texttt{R2OpenBugs} package, which allows for the user to send model scripts and data to OpenBugs without needing to leave the \texttt{R} interface.
+OpenBUGS is the oldest of the Bayesian sampling platforms that we use, maintained only as recently as 2014.
+OpenBUGS is a program for Windows machines, but there are workarounds to run it on MacOS and Linux, such as WINE or a virtual machine.
+OpenBUGS can be used through the actual click-and-point application, or through R, using the `R2OpenBUGS` package, which allows for the user to send model scripts and data to OpenBUGS without needing to leave the R interface.
 OpenBUGS is based on the BUGS model specification language, which is a small-to-moderately sized set of functions and commands that allow for extensive specification of bayesian models.
-Several other common bayesian sampling implementations are made to closely mimic the BUGS language as well.
+The BUGS language is declarative, meaning that order in which expressions are listed in the model code does not matter (Lunn, BUGS Project).
 
-<!-- Using this method causes OpenBugs to run invisibly in the background and return the results, unless an argument is specified to debug code in the actual application itself, which can be very useful.
-When using R to run OpenBugs, there is little way to know what went wrong if the code errors somewhere without interacting with the OpenBugs program itself.
+The program uses Gibbs sampling methods by default, but also allows for Metropolis-Hastings and Slice sampling.
+Depending on the model type, OpenBUGS will attempt to group nodes that are correlated together and update them at the same time in each Gibbs sampling transition.
+In most cases this is not feasible, and parameter nodes are updated one at a time, which can be very slow if there are many parameters or there is a high level of correlation between them (OpenBUGS Manual).
+
+<!-- Using this method causes OpenBUGS to run invisibly in the background and return the results, unless an argument is specified to debug code in the actual application itself, which can be very useful.
+When using R to run OpenBUGS, there is little way to know what went wrong if the code errors somewhere without interacting with the OpenBUGS program itself.
 Using the \texttt{debug = T} argument within \texttt{bugs()} will cause the application to run in the forefront as if it were actually being used directly by the user.
@@
 
-OpenBugs has many tools to visualize the performance the sampler which are easy to use.
-The biggest drawback of using the OpenBugs application directly is that specifiying the data can be tedious if the data is not very small in size or design.
+OpenBUGS has tools to visualize the performance diagnostics of the sampling, which are easy to use.
+The biggest drawback of using the OpenBUGS application directly is that specifying the data by hand can be tedious if the data is not very small in size or design structure.
 Using R to specify the data is much easier.
-One unique feature in OpenBugs is the Doodle tool, which allows for model specification through the creation of a diagram of nodes representing the parameters.
-Though using Doodle can be helpful for illustration, it is not used often in daily practice.
+
+One main purpose of the BUGS language and software is to easily work with directed acyclical graphs (DAGs).
+The syntax makes specifying child-parent relationships easy, and the program can interpret the model code and automatically determine the nodes relevant to each full conditional distribution calculation used in Gibbs sampling.
+Because of how flexible the BUGS language is, any DAG can be represented through a BUGS model (Lunn).
+One of the unique feature in OpenBUGS is the Doodle tool, which allows for model specification through the creation of a DAG itself by the user, where each represents a variable.
+A Doodle can be converted into a model, though this is not used often in daily practice.
 
@@
 
+There are a few well-known weaknesses to OpenBUGS.
+The software only works on Windows platforms, although there have been ways to use the program on MacOS through The program itself can be prone to failure on occasion, and there are no built-in diagnostic checks to alert the user if the sampling has gone poorly.
+Diagnosing errors in the model code can be somewhat tedious if the errors given aren't very descriptive.
+OpenBUGS does not support logical expressions such as the if-else statement, although there are workarounds, such as the `step()` function that is built-in (Lunn).
+
 ## JAGS
 
-In terms of model specification, JAGS is very similar to OpenBugs, as it is based on the BUGS model specification language.
-However, JAGS is always interacted with entirely in R, making the bayesian inference process more seemless, using the R package \texttt{rjags}.
-The \texttt{runjags} is an optional package that allows for the JAGS model script to be inputted as a string, and run using the \texttt{run.jags()} function.
-Built on C++, JAGS is able to run MCMC sampling process very quickly.
-There is plenty of documentation and support available for JAGS, which is still maintained.
+JAGS (Just Another Gibbs Sampler) is similar in purpose to OpenBUGS, in that it does analysis of Bayesian models using MCMC methods, and is based on the BUGS model specification language.
+However, JAGS is designed to be interacted with entirely in R, making the bayesian inference process more seamless, using the R package `rjags`.
+The `runjags` is an optional package that allows for the JAGS model script to be inputted as a string, and run using the `run.jags()` function.
+JAGS is also built on C++, which makes running the MCMC sampling process very fast.
+Sampling diagnostics can be visualized through the `coda` package.
+There is plenty of documentation and support available for JAGS, which is still maintained (JAGS manual).
 
@@
 
-Nimble allows for bayesian computation within \texttt{R}, like JAGS and OpenBUGS, but is also made for more flexible statistical modeling.
+Nimble allows for bayesian computation within R, like JAGS and OpenBUGS, but is also made for more flexible statistical modeling.
 It uses C++ under the hood to compile models and algorithms.
-Using the \texttt{nimble} package, Nimble can perform MCMC bayesian sampling using common algorithms, but also allows for other types of algorithms to be loaded or custom-built.
-It can also compile other code without BUGS models into C++ for other uses.
-For the purpose of bayesian sampling using MCMC, Nimble is used very similarly to JAGS and OpenBUGS in practicality, as its model specification language is similar.
-The flexibility of Nimble is not needed in most cases when doing bayesian MCMC methods.
-The default sampling method in Nimble is Gibbs sampling, but other methods can be used as well.
-Because the model has to be compiled first for C++ before sampling each time, this can drastically increase the total computation time as compared to other sampling methods.
+Using the `nimble` package, Nimble can perform MCMC bayesian sampling using common algorithms, but also allows for other types of algorithms to be loaded or custom-built (Michaud).
+It can also compile other types of algorithms without the BUGS syntax, written in a familiar fashion for R users.
 
-## STAN
+For the purpose of bayesian sampling using MCMC, Nimble is used very similarly to JAGS and OpenBUGS in practicality, as its model specification language is close to the same.
+The flexibility of Nimble is not needed in most cases when doing standard bayesian MCMC methods.
+Nimble can load and use BUGS model and data files if desired.
+The default sampling method in Nimble is Gibbs sampling, but other algorithms are built-in as well, like Slice sampling, reversible jump MCMC, and Sequential Monte Carlo (Nimble Manual).
+Because the model has to be compiled first into C++ before sampling each time, this can increase the total computation time as compared to other sampling methods.
+There is a very accessible user manual for Nimble available online.
 
-STAN is a language in-and-of itself used for statistcal modeling and data analysis.
-STAN can be used in several programming languages, including \texttt{R} and Python.
-STAN uses Hamiltonian Monte Carlo sampling methods, which ideally can reduce the computational time for many types of statistical models.
-Like Nimble, STAN compiles each model for C++, but one convenient feature is that the user can store each compiled model in a file, which allows for repeated use without the need to re-compile each time.
-In \texttt{R}, STAN is used through the \texttt{rstan} package, which allows for data to be loaded through normal \texttt{R} data structures.
-The biggest practical difference in using STAN compared to other bayesian computational methods is in the model specification process.
-STAN requires a model object in the form of a .stan file, which is created in a very unique language that is more formal than BUGS syntax.
-A .stan file, either created through a character string in R or in a seperate text file, can be used by the \texttt{stan()} function, which takes the file along with the data to perform the sampling.
-There is a very robust set of manuals for STAN users, including many papers written on the subject.
+## Stan
 
-## Greta
+Stan is a probabilistic programming language in-and-of itself, used for bayesian statistical modeling.
+Stan has two key differences from the BUGS-based implementations: it has its own language and syntax, and it uses a different MCMC algorithms.
+Stan has interfaces for R, Python, and the command line.
 
-Greta, a more recent addition to the list of bayesian computational implementations, is made specifically for easy use in \texttt{R}.
-The model and data specification process is intended to be as straightforward as possible users who are familiar with \texttt{R} but not as familiar with other bayes methods.
+The Stan language is designed to be more flexible and expressive than the BUGS-type languages.
+Explicit variable declarations are required in Stan, and, unlike the BUGS models, and statements are executed in the order that they appear in the model file.
+It also supports more programming techniques, such as conditional statements using if-else syntax (Carpenter).
+
+Stan uses Hamiltonian Monte Carlo sampling methods, which ideally is more efficient and robust than Gibbs or Metropolis-Hastings strategies.
+In some cases, this can reduce the computational time for many types of statistical models.
+Like Nimble, Stan model code is translated and compiled to C++ for faster use.
+In addition, Stan conveniently gives the user the option to store a compiled model in a file, which allows for repeated use without the need to re-compile each time.
+Stan supports both the Euclidean HMC sampler, and the no-U-turn sampler (NUTS), which is the default method (Carpenter).
+NUTS automatically adapts the number of leapfrog steps in the warmup phase and does not require user-specified tuning parameters (Stan User Manual).
+
+In R, Stan is used through the `rstan` package, which allows for data to be loaded through normal R data structures.
+The biggest practical difference in using Stan compared to other packages is in the model specification process.
+Stan requires a model object in the form of a .stan file.
+This file, either created through a character string in R or in a separate text file, can be used by the `stan()` function, which takes the file along with the data to perform the sampling.
+There is a very robust set of manuals for Stan users, including many papers written on the subject.
+
+## greta
+
+Greta, a more recent addition to the list of bayesian computational implementations, is made specifically for easy use in R.
+The model and data specification process is intended to be as straightforward as possible for R users, who need not be familiar with other bayesian inference tools.
 Not only does it utilize Hamiltonian Monte Carlo methods for sampling, but it also uses Google TensorFlow,a machine learning platform from Google, to perform its calculations.
 This allows for more flexible use of CPUs and GPUs.
-Greta offers a model visualization tool that does not exist in any of the other Bayesian samplers, in that it can provide a graphical representation of the specified model through the use of nodes and arrows.
-This tool can be very useful in trying to explain how all the components of a model work together.
+Greta offers a model visualization tool that does not exist in any of the other Bayesian samplers, in that it can provide a graphical representation of the specified model, similar to a directed acyclical graph (greta Manual).
 
-Unfortunately, as of the writing of this, Greta is still in need of lots of development, and lacks an appropriate amount of documentation and support for users, compared to the other implementations mentioned previously.
-The sampling process can slow compared to the other bayes methods when using identical sampling configurations, especially for simple models.
-There are also several types of models that Greta seems unable to properly specify at the moment.
+As of the writing of this, greta is still in need of lots of development, and lacks an appropriate amount of documentation and support for users, compared to the other implementations mentioned previously.
+The sampling process can slow compared to the other bayesian methods when using identical sampling configurations, especially for simple models.
+There are also several types of models for which greta seems unable to properly specify at the moment.
 
@@
 
-## OpenBUGS
+### OpenBUGS
 
@@
 }
-
-##
 ```
@@
 
-## JAGS
+### JAGS
 
@@
 
-## Nimble
+### Nimble
 
@@
 
-## STAN
+### Stan
 
-Running models in STAN requires a .stan file which contains the model script code, written in the unique STAN syntax.
+Running models in Stan requires a .stan file which contains the model script code, written in the unique Stan syntax.
 This can either be a seperate file, or can be inputted as a character string.
-It is generally recommended to use a seperate .stan file, espcially when using RStudio, which can do syntax highlighting and point out errors in STAN code.
+It is generally recommended to use a seperate .stan file, espcially when using RStudio, which can do syntax highlighting and point out errors in Stan code.
 
 Here is the .stan file for this example, which is saved as "stan_file.stan".
-Each STAN file contains a data block, parameters block, and model block at a minimum.
+Each Stan file contains a data block, parameters block, and model block at a minimum.
 The data block initializes each data variable, the parameters block initializes any parameters to monitor, and the model block contains more familiar looking model script code.
@@
 
-## Greta
+### greta
 
-Model specification in Greta is more intuitive for R users, but requires the extra step of converting existing data objects into Greta arrays, using the `as_data()` function from the `greta` package. 
+Model specification in greta is more intuitive for R users, but requires the extra step of converting existing data objects into greta arrays, using the `as_data()` function from the `greta` package. 
 ```{r}
@@
 
-When using Markov Chain Monte Carlo methods to estimate Bayesian posteriors of parameters, the number of independent Markov chains to be run needs to be specified.
+When using Markov Chain Monte Carlo methods, the number of independent Markov chains needs to be specified.
 It is recommended to use several chains to explore the space of the target distribution, as a single chain may not fully traverse the entire region on its own.
@@
 
-The number of steps that each Markov chain will take is a very important setting to get right.
+The number of total draws that each Markov chain will collect is a very important setting to get right.
 If each chain is not set up to run long enough, the parameter space may not be fully explored and the estimates will be inaccurate.
@@
 
-Since our inferences are based on the assmumption that our draws are from the target distribution, it is wise to discard the samples at the beginning of each chain, since it takes time for the sampler to "settle down".
+Since our inferences are based on the assmumption that our draws are from the target distribution, it is wise to discard the samples at the beginning of each chain, since it takes time for the Markov chains to converge to the target distribution.
 The terms "warmup" and "burn-in" are often used to indicate the number of iterations that we will discard at the beginning of each chain when calculating statistics about our posterior distribution.
-We can choose this number in the function call for each sampler.
-In \texttt{OpenBUGS}, \texttt{Nimble}, and \texttt{Jags}, the argument is referred to as burnin, but in \texttt{STAN} and \texttt{greta} it is called warmup.
-Some samplers (\texttt{greta}, \texttt{Jags, Nimble}) treat this warmup argument as a number of iterations to run before starting to track the sampling results, but others (\texttt{STAN}, \texttt{BUGS}) have this be the actual number of iterations to discard out of the total number of specified iterations, so it is important to know which designation a sampler uses.
+While "warmup" is sometimes used interchangeably with "burnin", the term is usually associated with Hamiltonian Monte Carlo samplers, indicating the number of iterations that the sampler will spend adapting and fine-tuning parameters before collecing official draws.
+Gibbs Samplers usually use "burnin" to indicate the number of initial draws that will be discarded at the beginning of each chain before keeping track of the draws.
+
+We can choose these values for warmup/burnin in the function call for each sampler.
+Some samplers (greta, JAGS, Nimble) treat this argument as a number of iterations to run before starting to track the sampling results, but others (Stan, OpenBUGS) have this be the actual number of iterations to discard out of the total number of specified iterations, so it is important to know which designation a sampler uses.
+For example, if one were to have 5000 iterations with 1000 burnin iterations, in JAGS this would lead to 1000 discared iterations, followed by 5000 draws, but in OpenBUGS, there would 5000 total draws, with the first 1000 discared, leaving 4000 posterior draws for each chain.
 
@@
 
-In some cases, there is high correlation between sequential samples in a given chain, which can bias the resulting posterior inference, which can be diagnosed using some graphical procedures.
-In this situation, it is wise to "thin" the results by only looking at every ith observation from a chain, such as every 10th observation.
+In some cases, there is high autocorrelation between sequential samples in a given chain, which can bias the resulting posterior inference.
+These autocorrelation issues can be diagnosed using some graphical procedures.
+If there is autocorrelation, it is wise to "thin" the results by only looking at every ith observation from a chain, such as every 10th observation.
 While each sampler defaults to including every simulated posterior draw, a user can specify a "thinning rate", which is the rate at which a sampler will keep memory of the observations.
@@
 
-In order to start sampling, each Markov chain needs to have an initial value at which to start.
+In order to start sampling, each Markov chain needs to have an initial value at which to start for each parameter.
 These initial values can be either be specified, or can often be randomly generated based on the prior distributions for each parameter of interest.
@@
 In these cases, a more strategic choice of initial values is advised.
-\texttt{Nimble} unfortunately requires for these values to be specified instead of randomly generated.
-Using \texttt{R} to randomly generate the values from the prior distributions can be used in this situation:
+Nimble unfortunately requires for these values to be specified instead of randomly generated.
+Using R to randomly generate the values from the prior distributions can be used in this situation:
 
@@
 
-When describing a bayesian model through code, there are instances were slightly different implementations are needed based on the sampler being used.
+When describing a bayesian model through code, there are instances where slightly different specifications are needed based on the sampler being used.
 Both in model and data specification, the user needs to understand the intricacies of the specific sampling tool being used in order to get the analysis to run properly.
@@
 
-##### OpenBUGS
-In \texttt{OpenBUGS}, the earliest method created of the lot, the model specification code must be wrapped inside the expression \texttt{model\{\}}:
+**OpenBUGS**
+
+In OpenBUGS, the model specification code must be wrapped inside the expression `model\{\}`:
 
@@
 
-Since we are using the \texttt{R2OpenBUGS} library to run \texttt{OpenBUGS} through an \texttt{R} session, a text file with the model code is passed to the \texttt{bugs()} function.
-We can specify the model in \texttt{R} and create the appropriate text file:
+We can either create this text file ourselves, or we can have R generate the text file automatically. The code below demonstrates how to create the model text file entirely in R.
 
@@
 
-##### JAGS
-The \texttt{JAGS} sampler was intended to allow for model specification code to be almost identical to \texttt{BUGS}, so the implementation is very similar.
-There are two common approaches, one using the \texttt{rjags} package, and one using the \texttt{runjags} package.
+### JAGS
+
+The JAGS sampler was intended to allow for model specification code to be almost identical to OpenBUGS, so the implementation is very similar.
+Like in OpenBUGS, the user can create a text file with the model code, or specify the model in R directly.
+When using the R approach, there are two common approaches, one using the `rjags` package, and one using the `runjags` package.
 We use the latter, which allows for the model code to be specified in a multi-line character string:
@@
 
-##### Nimble
-Model specification in \texttt{Nimble} is extremely similar to \texttt{BUGS} as well.
-Model code is wrapped inside the \texttt{nimble} package function \texttt{nimbleCode()}, and then fed into \texttt{nimbleMCMC()} sampler function:
+### Nimble
+
+Model specification in Nimble is extremely similar to OpenBUGS as well. Model code is wrapped inside the `nimble` package function `nimbleCode()`, and then fed into `nimbleMCMC()` sampler function:
 
@@
 
-##### STAN
-The STAN model specification language is significantly different than other samplers, in that a .stan file is fed to the \texttt{C++} compiler.
+### Stan
+
+The Stan model specification language is significantly different than other samplers, in that a .stan file is fed to the C++ compiler.
 A .stan file has a very unique model specification format, which divides code into "data", "parameter", and "model" blocks at a minimum.
@@
 
-A character string specifying the path to the .stan file is then passed to \texttt{stan()}.
+A character string specifying the path to the .stan file is then passed to `stan()`.
 
@@
 
-Specifying data for the bayesian samplers in \texttt{R} is very straightforward.
-In \texttt{OpenBUGS}, \texttt{JAGS}, and \texttt{STAN}, a list containing each data object or constant can be passed to the function that calls the sampler.
-For example, in \texttt{JAGS}, the data might constructed like this:
+Specifying data for the bayesian samplers in R is very straightforward.
+In OpenBUGS, JAGS, and Stan, a list containing each data object or constant can be passed to the function that calls the sampler.
+For example, in JAGS, the data might constructed like this:
 
@@
 
-In \texttt{Nimble}, the constants need to be specified in a separate list from the data.
+In Nimble, the constants need to be specified in a separate list from the data.
 
@@
 
-In \texttt{OpenBUGS}, \texttt{JAGS}, \texttt{Nimble}, and \texttt{STAN}, specifying parameters for the sampler to monitor is as easy as providing a character vector with the names of the parameters to monitor:
+In OpenBUGS, JAGS, Nimble, and Stan, specifying parameters for the sampler to monitor is as easy as providing a character vector with the names of the parameters to monitor:
 
@@
 
-## STAN Probabilty and Sampling Statements
+## Stan Probabilty and Sampling Statements
 
@@
 
-This kind of probability statement does not actually perform any sampling, but is instead tranformed into a statement about the incremental log probability density of a target distribution.
+This kind of probability statement does not actually perform any sampling, but is instead transformed into a statement about the incremental log probability density of a target distribution.
 The log probability density function can be specified directly.
@@
 
-## Greta's Unique Specification Language
+## greta's Unique Specification Language
 
-Implementing models in \texttt{greta} is very unique in that each data variable or parameter is initialized seperately with its own line of \texttt{R} code.
-Each data object in \texttt{R} is redefined using the \texttt{as\_data()} function in the \texttt{greta} package:
+Implementing models in greta is very unique in that each data variable or parameter is initialized seperately with its own line of R code.
+Each data object in R is redefined using the `as\_data()` function in the greta package:
 ```{r}
@@
 
-Once all the variables have been defined, the relationship between the data and the unknown parameters is established using the \texttt{distribution()} function:
+Once all the variables have been defined, the relationship between the data and the unknown parameters is established using the `distribution()` function:
 
@@
 
-Finally, the greta model is constructed using \texttt{model()} function, which takes each of the monitored parameters as arguments:
+Finally, the greta model is constructed using `model()` function, which takes each of the monitored parameters as arguments:
 
@@
 
-# How to Specify Tricky Models
+# Benchmarking
+
+There are several important performance aspects to consider when assessing bayesian computational strategies.
+The main two that we will focus on are computation time and computational accuracy.
+A reliable sampler should both be efficient in computation and give trustworthy posterior estimations.
+We have conducted simulations to benchmark OpenBUGS, JAGS, Nimble, Stan, and greta to compare their performance in each area.
+
+## Computation Time Benchmarking
+
+In order to assess each sampler's performance across many different scenarios, we chose 16 different model types on which to measure the computation time performance, including simple conjugate models, regression models, hierarchical models, time series models, survival models, and others.
+For each platform, we ran computations on each model, both for a small dataset, and a large dataset.
+Each model computation was done a total of 200 times for each implementation.
+In all computations, the MCMC sampling was performed with four chains, using 10,000 posterior draws, with a warmup / burnin of 1000 iterations prior to sampling.
+Each sampling was computed in parallel, on the same workstation iMac (2017, 8 GB 2400 MHz DDR4 Memory, 3.4 GHz Quad-Core Intel Core i5 Processor, running MacOS Catalina).
+
+## Computation Time Results
+
+\textcolor{red}{Computation time results here.}
+
+## Posterior Estimation Accuracy Benchmarking
+
+When using MCMC methods to approximate a posterior distribution, the samples that we collect should, in theory, be coming from the posterior itself.
+While we cannot determine if a single draw is from that distribution, there are ways to aggregate all of the draws and see how closely the empirical distribution matches the "true" distribution.
+For certain conjugate probability models, we can analytically calculate the posterior distribution for the parameters of interest.
+In these cases, we will use the Total Variation Distance to measure the distance between the true posterior distribution and the estimated posterior distribution, in the form of an empirical distribution obtained from MCMC sampling.
+
+### Total Variation Distance
+
+Total Variation Distance (TVD) is a metric that defines the distance between two probability measures on the same space. For two probabilty measures $P$ and $Q$, the TVD is defined as
+
+$$\delta(P,Q) = \textrm{sup}_{A \in B} \left| P(A) - Q(A)\right|.$$
+
+When we are dealing with probability distributions on $\mathbb{R}$, the TVD becomes
+
+$$\delta(P,Q) = \frac{1}{2} \int_\mathbb{R} \left|p(x) - q(x)\right| dx.$$
+
+\textcolor{red}{Proof if needed.}
+
+When our probability models allow us to calculate the analytical posterior distribution, we can use the TVD to calculate the distance between the "true" distribution and the estimated one.
+Using this method, we can compare the posterior estimation accuracy of different bayesian computation implementations (Casement) (Michel).
+
+To compare the posterior estimation accuracy of our five implementations, we picked six different data models with conjugate prior distributions (such as binomial data with a beta prior for $\theta$), and randomly simulated data to analytically calculate the true posterior distribution in each case.
+Then, for each of OpenBUGS, JAGS, Nimble, Stan, and greta, we ran a set of 50 simulations using the same data, each time running the MCMC sampler with four chains, 10,000 iterations, and a warmup or burnin of 1000, and obtained the samples from the posterior for each to calculate the Total Variation Distance between the true posterior and empirical posterior.
+For each model, we have aggregated the TVD values for each implementation.
+
+### Total Variation Distance Results
+
+\textcolor{red}{TVD results here.}
+
+### Posterior Estimation for Non-Conjugate Models
+
+In most cases where we do not have a conjugate data distribution and prior distribution, an analytic posterior cannot be computed.
+To obtain an approximation of the true posterior to the best of our ability, we will use Hamiltonian Monte Carlo (through Stan) to get a very large number of draws (way more than needed) from the posterior distribution.
+We will then treat this empirical posterior distribution as our "true" distribution, and compare it to empirical posteriors from OpenBUGS, JAGS, Nimble, Stan, greta, when they are used with much smaller iterations.
+
+### Results
+
+\textcolor{red}{Results here.}
+
+# Installation
+
+\textcolor{red}{Insert section on installation process for each implementation here.}
+
+# Conclusion
+
+\textcolor{red}{Insert conclusion and discussion here.}
+
+\newpage
+
+# Appendix
+
+## How to Specify Tricky Models
 
@@
 
-##### Truncated Distribution
+**Truncated Distribution**
 
@@
 
-## Model Text for OpenBUGS using R2OpenBUGS
+## Model Text for OpenBUGS using R2OpenBUGS to create the text file
 tau ~ dnorm(0,1) %_% I(0, )
@@
 
-## Greta
+## greta
 tau <- normal(0,1, truncation = c(0,Inf))
 
-## STAN
+## Stan
 # Parameter block initialization
@@
 
-##### Flat Priors
+**Flat Priors**
 
 Sometimes using a flat, diffuse prior is appropriate.
-Most of the samplers allow for improper flat priors, although \texttt{JAGS} does not.
-In \texttt{STAN}, a parameter has is given a flat, improper prior by default if it doesn't have a distribution assigned.
+Most of the samplers allow for improper flat priors, although JAGS does not.
+In Stan, a parameter has is given a flat, improper prior by default if it doesn't have a distribution assigned.
 Simply declaring the parameter will do the job.
@@
 
-## Greta
+## greta
 theta <- variable()
 
-## STAN
+## Stan
 # Parameter block initialization
@@
 
-##### Distribution Parameter Order
+**Distribution Parameter Order**
 
@@
 Some use the $\sigma$ parameterization of the normal distribution, where $\sigma$ is the standard deviation, and others use $\tau$ instead, where $\tau = 1 / \sigma^2$.
-Among bayesian samplers, \texttt{JAGS}, \texttt{OpenBUGS}, and \texttt{Nimble} use the $\tau$ parameteriztion, while \texttt{STAN} and \texttt{greta} use $\sigma$.
+Among bayesian samplers, JAGS, OpenBUGS, and Nimble use the $\tau$ parameteriztion, while Stan and greta use $\sigma$.
 
-\textit{\textbf{Normal}}: The $N(\mu,\sigma)$ parameterization is used by \texttt{STAN} and \texttt{greta}, where $\sigma$ is the standard deviation.
-The $N(\mu,\tau)$ parameterization is used by \texttt{JAGS}, \texttt{OpenBUGS}, and \texttt{Nimble}, which uses the precision parameter $\tau = 1 / \sigma^2$.
+\textit{\textbf{Normal}}: The $N(\mu,\sigma)$ parameterization is used by Stan and greta, where $\sigma$ is the standard deviation.
+The $N(\mu,\tau)$ parameterization is used by JAGS, OpenBUGS, and Nimble, which uses the precision parameter $\tau = 1 / \sigma^2$.
 The same parameterization guidelines are the same for the student-t distribution as well.
 
-\textit{\textbf{Binomial}}: The $Bin(n,p)$ parameterization is used by \texttt{OpenBUGS}, \texttt{JAGS}, and \texttt{Nimble}, while the $Bin(p,n)$ parameterization is used by \texttt{STAN} and \texttt{greta}.
+\textit{\textbf{Binomial}}: The $Bin(n,p)$ parameterization is used by OpenBUGS, JAGS, and Nimble, while the $Bin(p,n)$ parameterization is used by Stan and greta.
 
@@
 
-\textit{\textbf{Weibull}}: The $Weibull(v,\lambda)$ parameterization is used by \texttt{OpenBUGS}, \texttt{JAGS}, \texttt{Nimble}, and \texttt{greta}, where $\lambda$ is a shape parameter.
-The $Weibull(v,\sigma)$ parameterization is used by \texttt{STAN}, where $\sigma = 1 / \lambda$.
+\textit{\textbf{Weibull}}: The $Weibull(v,\lambda)$ parameterization is used by OpenBUGS, JAGS, Nimble, and greta, where $\lambda$ is a shape parameter.
+The $Weibull(v,\sigma)$ parameterization is used by Stan, where $\sigma = 1 / \lambda$.
 
-##### Equations inside distribution parameters
+**Equations inside distribution parameters**
 
-In order to utilized calculated parameters inside \texttt{OpenBUGS}, a variable must be defined using the assignment operator that is the result of the calculation:
+In order to utilized calculated parameters inside OpenBUGS, a variable must be defined using the assignment operator that is the result of the calculation:
 
@@
 
-Fortunately, the other samplers allow for variable calculations inside the actual parameter definition of a distribution function. Here is an example in \texttt{JAGS}:
+Fortunately, the other samplers allow for variable calculations inside the actual parameter definition of a distribution function. Here is an example in JAGS:
 

### 2021-04-28 8-Schools/8-Schools-bugs.R

@@ -87,9 +87,25 @@ if (!currently_benchmarking()) {
   
   bugs_fit$summary
   
+  bugs_fit_object <- bugs_fit$sims.array[,,1]
+  dim(bugs_fit_object) <- c(dim(bugs_fit_object), 1)
+  dimnames(bugs_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = bugs_monitor
+  )
+  
+  
+  bugs_fit_object %>% mcmc_areas()
+  bugs_fit_object %>% mcmc_intervals()
+  
   
   ## assess convergence issues 
   ###################################################################################
+  
+  bugs_fit_object %>% mcmc_acf_bar()
+  bugs_fit_object %>% mcmc_trace()
+  bugs_fit_object %>% mcmc_hist_by_chain()
   
 }
 

### 2021-04-28 hierarchical-models/beta-binomial/hierarchical-beta-binomial-jags.R

@@ -84,7 +84,8 @@ if (!currently_benchmarking()) {
   )
   
   
-  jags_fit_object %>% bayesplot::mcmc_dens()
+  jags_fit_object %>% mcmc_areas()
+  jags_fit_object %>% mcmc_intervals()
   
   
   ## assess convergence issues 
@@ -92,6 +93,7 @@ if (!currently_benchmarking()) {
   
   jags_fit_object %>% mcmc_acf_bar()
   jags_fit_object %>% mcmc_trace()
+  jags_fit_object %>% mcmc_hist_by_chain()
   
   
   

### 2021-04-28 hierarchical-models/large-full-beta-binomial/hierarchical-full-beta-binomial-bugs.R

@@ -91,9 +91,25 @@ if (!currently_benchmarking()) {
   
   bugs_fit$summary
   
+  bugs_fit_object <- bugs_fit$sims.array[,,1]
+  dim(bugs_fit_object) <- c(dim(bugs_fit_object), 1)
+  dimnames(bugs_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = bugs_monitor
+  )
+  
+  
+  bugs_fit_object %>% mcmc_areas()
+  bugs_fit_object %>% mcmc_intervals()
+  
   
   ## assess convergence issues 
   ###################################################################################
+  
+  bugs_fit_object %>% mcmc_acf_bar()
+  bugs_fit_object %>% mcmc_trace()
+  bugs_fit_object %>% mcmc_hist_by_chain()
   
 }
 

### 2021-04-28 hierarchical-models/large-full-beta-binomial/hierarchical-full-beta-binomial-jags.R

@@ -87,7 +87,8 @@ if (!currently_benchmarking()) {
   )
   
   
-  jags_fit_object %>% bayesplot::mcmc_dens()
+  jags_fit_object %>% mcmc_areas()
+  jags_fit_object %>% mcmc_intervals()
   
   
   ## assess convergence issues 
@@ -95,6 +96,7 @@ if (!currently_benchmarking()) {
   
   jags_fit_object %>% mcmc_acf_bar()
   jags_fit_object %>% mcmc_trace()
+  jags_fit_object %>% mcmc_hist_by_chain()
   
   
   

### 2021-04-28 hierarchical-models/large-full-beta-binomial/hierarchical-full-beta-binomial-stan.R

@@ -53,8 +53,7 @@ source(here("currently-benchmarking.R"))
 if (!currently_benchmarking()) {
   stan_fit <- stan(
     "file" = stan_file, "data" = stan_data, 
-    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup, 
-    "control" = list("adapt_delta" = 0.99)
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
   )
   
   
@@ -66,7 +65,7 @@ if (!currently_benchmarking()) {
   summary(stan_fit)$summary
   get_posterior_mean(stan_fit)
   stan_dens(stan_fit) + theme_bw()
-  stan_fit %>% as.array() %>% bayesplot::mcmc_dens()
+  stan_fit %>% as.array() %>% mcmc_areas()
   
   
   
@@ -74,8 +73,9 @@ if (!currently_benchmarking()) {
   ###################################################################################
   
   stan_fit %>% as.array() %>% mcmc_acf_bar()
-  stan_fit %>% as.array() %>% mcmc_pairs()
   stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
   
   # see each chain
   stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)

### 2021-04-28 hierarchical-models/poisson-gamma/hierarchical-poisson-gamma-bugs.R

@@ -1,0 +1,107 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("R2OpenBUGS")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+N <- 8 # sample size
+y <- c(10, 9, 21, 4, 5, 21, 1, 1) # data
+t <- c(47.6, 53.4, 44.4, 45, 51.1, 58, 55.3, 43.7) # known constants
+
+bugs_data <- list(
+  "y" = y,
+  "N" = N,
+  "t" = t
+)
+
+## specify bugs model
+################################################################################
+
+bugs_model <- function() {
+  for (i in 1:N) {
+    theta[i] ~ dgamma(alpha, beta) # prior on the thetas
+    lambda[i] <- theta[i] * t[i] # define Poisson rate
+    y[i] ~ dpois(lambda[i]) # likelihood
+  }
+  alpha ~ dexp(1) # hyperprior parameters
+  beta ~ dgamma(0.1, 1.0)
+}
+
+
+bugs.file <- file.path(tempdir(), "model.txt")
+write.model(bugs_model, bugs.file)
+
+bugs_monitor <- "theta"
+
+
+## Specify path to WINE if using WINE 
+###################################################################################
+
+if (getwd() == "/Users/evanmiyakawa/hubiC/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
+  WINE="/usr/local/Cellar/wine/4.0.1/bin/wine"
+  WINEPATH="/usr/local/Cellar/wine/4.0.1/bin/winepath"
+  OpenBUGS.pgm="/Users/evanmiyakawa/OpenBugs/OpenBUGS323/OpenBUGS.exe" #~
+} else {
+  WINE="/Users/evan_miyakawa1/Cellar/wine/4.0.1/bin/wine"
+  WINEPATH="/Users/evan_miyakawa1/Cellar/wine/4.0.1/bin/winepath"
+  OpenBUGS.pgm="/Users/evan_miyakawa1/OpenBugs/OpenBUGS323/OpenBUGS.exe"
+}
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
[…]

### 2021-04-28 large-data-parameters-models/hierarchical-models/full-beta-binomial/large-hierarchical-full-beta-binomial-bugs.R

@@ -1,0 +1,115 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("R2OpenBUGS")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+alpha <- 2 #beta distribution alpha
+beta <- 1 #beta distribution beta
+N <- 1000 # sample size
+
+set.seed(1)
+
+(n <- rbinom(N, 20, 0.5))
+(theta <- rbeta(N, alpha, beta))
+(y <- rbinom(N, n, theta))
+
+
+bugs_data <- list(
+  "y" = y,
+  "N" = N,
+  "n" = n
+)
+
+
+
+## specify bugs model
+################################################################################
+
+bugs_model <- function() {
+  for (i in 1:N) {
+    theta[i] ~ dbeta(alpha, beta)
+    y[i] ~ dbin(theta[i], n[i])
+  }
+  alpha ~ dgamma(1,1)
+  beta ~ dgamma(1,1)
+}
+
+
+bugs.file <- file.path(tempdir(), "model.txt")
+write.model(bugs_model, bugs.file)
+
+bugs_monitor <- c("theta", "alpha", "beta")
+
+
+## Specify path to WINE if using WINE 
+###################################################################################
+
+if (getwd() == "/Users/evanmiyakawa/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
+  WINE="/usr/local/Cellar/wine/4.0.1/bin/wine"
+  WINEPATH="/usr/local/Cellar/wine/4.0.1/bin/winepath"
+  OpenBUGS.pgm="/Users/evanmiyakawa/OpenBugs/OpenBUGS323/OpenBUGS.exe" #~
+} else {
+  WINE="/Users/evan_miyakawa1/Cellar/wine/4.0.1/bin/wine"
+  WINEPATH="/Users/evan_miyakawa1/Cellar/wine/4.0.1/bin/winepath"
+  OpenBUGS.pgm="/Users/evan_miyakawa1/OpenBugs/OpenBUGS323/OpenBUGS.exe"
+}
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
[…]

### 2021-04-28 large-data-parameters-models/old/ar-model/ar-model-greta.R

@@ -1,0 +1,94 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("greta")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+## generate/specify data
+################################################################################
+
+N <- 200L # sample size
+alpha <- 0 # ar(1) mean
+beta <- c(-0.6) # ar(1) coefficient
+sigma <- 1 # ar(1) standard deviation
+
+
+set.seed(1)
+(y <- arima.sim(model = list(ar = beta), sd = sigma, n = N) %>% as.numeric())
+
+t <- 1:length(y)
+ggplot(data.frame(t = t, y = y), aes(t,y)) + geom_line()
+
+y_1 <- y[1:(length(y) - 1)] # data excluding the last point
+z_1 <- y[2:length(y)] # data excluding the first point
+
+y <- as_data(y)
+y_1 <- as_data(y_1) # 
+z_1 <- as_data(z_1)
+
+
+## specify greta model
+################################################################################
+
+alpha <- normal(0,1000)
+beta <- normal(0,1000)
+sigma <- normal(0,1000, truncation = c(0,Inf))
+distribution(y) <- normal(alpha + beta * y, sigma)
+distribution(z_1) <- normal(alpha + beta * y_1, sigma)
+
+greta_model <- model(alpha, beta, sigma)
+
+plot(greta_model)
+
+## configure model settings
+################################################################################
+
+n_chains <- 4
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
+################################################################################
+
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  greta_fit <- mcmc(
+    "model" = greta_model, "n_samples" = n_iter,
+    "warmup" = n_warmup, "chains" = n_chains
+  )
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(greta_fit)
+  
+  greta_fit_object <- greta_fit %>% as.array()
+  dim(greta_fit_object) <- c(dim(greta_fit_object), 1)
+  dimnames(greta_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = c("theta")
+  )
+  
+  
+  greta_fit_object %>% mcmc_areas()
[…]

### 2021-04-28 large-data-parameters-models/old/ar-model/ar-model-jags.R

@@ -1,0 +1,105 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rjags"); library("runjags")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+N <- 200L # sample size
+alpha <- 0 # ar(1) mean
+beta <- c(-0.6) # ar(1) coefficient
+sigma <- 1 # ar(1) standard deviation
+
+
+set.seed(1)
+(y <- arima.sim(model = list(ar = beta), sd = sigma, n = N)  %>% as.numeric())
+
+t <- 1:length(y)
+ggplot(data.frame(t = t, y = y), aes(t,y)) + geom_line()
+
+
+jags_data <- list(
+  "N" = N,
+  "y" = y
+)
+
+
+## specify jags model
+################################################################################
+
+jags_model <- "
+  model{
+    for (i in 2:N) {
+      y[i] ~ dnorm(alpha + beta * y[i-1], tau)
+    }
+    alpha ~ dnorm(0,0.0001)
+    beta ~ dnorm(0,0.0001)
+    tau ~ dnorm(0,0.0001)  I(0, )
+    sigma <- sqrt(1 / tau)
+  }
+"
+
+jags_monitor <- c("alpha", "beta", "sigma")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  jags_fit <- run.jags(
+    "model" = jags_model, "data" = jags_data, "monitor" = jags_monitor, 
+    "n.chains" = n_chains, "sample" = n_iter, "burnin" = n_warmup
+  ) 
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  jags_fit
+  
+  jags_fit_object <- jags_fit$mcmc %>% as.array()
+  dim(jags_fit_object) <- c(dim(jags_fit_object), 1)
+  dimnames(jags_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = jags_monitor
+  )
+  
+  
[…]

### 2021-04-28 large-data-parameters-models/old/ar-model/ar-model-nimble.R

@@ -1,0 +1,112 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("nimble")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+N <- 200L # sample size
+alpha <- 0 # ar(1) mean
+beta <- c(-0.6) # ar(1) coefficient
+sigma <- 1 # ar(1) standard deviation
+
+
+set.seed(1)
+(y <- arima.sim(model = list(ar = beta), sd = sigma, n = N)  %>% as.numeric())
+
+t <- 1:length(y)
+ggplot(data.frame(t = t, y = y), aes(t,y)) + geom_line()
+
+
+nimble_data <- list(
+  "y" = y
+)
+
+nimble_constants <- list(
+  "N" = N
+)
+
+
+## specify nimble model
+################################################################################
+
+nimble_model <- nimbleCode({
+  for (i in 2:N) {
+    y_hat[i] <- alpha + beta * y[i-1]
+    y[i] ~ dnorm(y_hat[i], tau)
+  }
+  alpha ~ dnorm(0,0.0001)
+  beta ~ dnorm(0,0.0001)
+  tau ~ T(dnorm(0,0.0001),0,)
+  sigma <- sqrt(1 / tau)
+})
+
+nimble_monitor <- c("alpha", "beta", "sigma")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+nimble_inits <- list(
+  "alpha" = rnorm(1,0,1000),
+  "beta" = rnorm(1,0,1000),
+  "tau" = rnorm(1,0,0.001) %>% abs()
+)
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  nimble_fit <- nimbleMCMC(
+    "code" = nimble_model, "data" = nimble_data, "constants" = nimble_constants,
+    "inits" = nimble_inits, "monitors" = nimble_monitor, "nchains" = n_chains, 
+    "niter" = n_iter, "nburnin" = n_warmup, "summary" = TRUE
+  )
+  
+  
+  ## assess fit
+  ################################################################################
+  
[…]

### 2021-04-28 large-data-parameters-models/old/ar-model/ar-model-stan.R

@@ -1,0 +1,87 @@
+## load required packages and set basic options
+################################################################################
+
+library("here")
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rstan"); rstan_options(auto_write = TRUE)
+library("bayesplot")
+library("bench")
+
+
+
+## generate/specify data
+################################################################################
+
+N <- 200L # sample size
+alpha <- 0 # ar(1) mean
+beta <- c(-0.6) # ar(1) coefficient
+sigma <- 1 # ar(1) standard deviation
+
+
+set.seed(1)
+(y <- arima.sim(model = list(ar = beta), sd = sigma, n = N)) %>% as.numeric()
+
+t <- 1:length(y)
+ggplot(data.frame(t = t, y = y), aes(t,y)) + geom_line()
+
+
+stan_data <- list(
+  "N" = N,
+  "y" = y
+)
+
+
+
+## specify stan model
+################################################################################
+
+# read it in from file
+stan_file <- here("time-series", "ar-model", "ar-model.stan")
+# file.show(stan_file)
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {  
+  stan_fit <- stan(
+    "file" = stan_file, "data" = stan_data, 
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
+  )
+  
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(stan_fit)$summary
+  get_posterior_mean(stan_fit)
+  stan_dens(stan_fit) + theme_bw()
+  stan_fit %>% as.array() %>% mcmc_areas()
+  
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  stan_fit %>% as.array() %>% mcmc_acf_bar()
+  stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
+  
+  # see each chain
+  stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)
+  
+}
+
+

### 2021-04-28 large-data-parameters-models/old/gamma-prior/poisson-gamma-benchmark.R

@@ -1,0 +1,59 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("bench")
+library("here")
+library("rjags"); library("runjags")
+library("R2OpenBUGS")
+library("nimble")
+library("rstan"); rstan_options(auto_write = FALSE)
+library("greta")
+options("bayes_benchmark" = TRUE)
+
+## Source code
+################################################################################
+
+source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-jags.R"))
+source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-bugs.R"))
+source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-nimble.R"))
+source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-stan.R"))
+source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-greta.R"))
+
+options("bayes_benchmark" = FALSE)
+
+## Set Parameters
+################################################################################
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+## benchmarking
+###################################################################################
+
+num_iterations <- 2 # Number of times you want to benchmark each model\
+
+rds_file_location <- here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma.rds")
+
+source(here("benchmark-function.R"))
+
+# choose to include stan compilation time with argument "stan_compile = TRUE"
+run_benchmark(rds_file_location)
+run_benchmark(rds_file_location, stan_compile = TRUE)
+run_benchmark(rds_file_location, stan_compile = TRUE, stan_only = TRUE)
+
+## benchmarking accuracy
+###################################################################################
+
+num_iterations <- 20 # Number of times you want to benchmark each model\
+
+source(here("benchmark-accuracy-function.R"))
+
+true_dist <- distr::Gammad(7,0.5)
+true_mean <- 3.5
+
+run_accuracy_benchmark(true_dist,num_iterations, n_iter, n_warmup)
+
+
+

### 2021-04-28 large-data-parameters-models/old/gamma-prior/poisson-gamma-greta.R

@@ -1,0 +1,83 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("greta")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+## generate/specify data
+################################################################################
+
+theta <- 5 # poisson theta
+
+set.seed(1)
+
+(y <- as_data(rpois(1, theta)))
+
+
+
+## specify greta model
+################################################################################
+
+theta <- gamma(3,1)
+distribution(y) <- poisson(theta)
+
+greta_model <- model(theta)
+
+plot(greta_model)
+
+## configure model settings
+################################################################################
+
+n_chains <- 4
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+
+## fit model
+################################################################################
+
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  greta_fit <- mcmc(
+    "model" = greta_model, "n_samples" = n_iter,
+    "warmup" = n_warmup, "chains" = n_chains
+  )
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(greta_fit)
+  
+  greta_fit_object <- greta_fit %>% as.array()
+  dim(greta_fit_object) <- c(dim(greta_fit_object), 1)
+  dimnames(greta_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = c("theta")
+  )
+  
+  
+  greta_fit_object %>% mcmc_areas()
+  greta_fit_object %>% mcmc_intervals()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  greta_fit_object %>% mcmc_acf_bar()
+  greta_fit_object %>% mcmc_trace()
+  greta_fit_object %>% mcmc_hist_by_chain()
+  
+}
+

### 2021-04-28 large-data-parameters-models/old/gamma-prior/poisson-gamma-nimble.R

@@ -1,0 +1,96 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("nimble")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+theta <- 5 # poisson theta
+
+set.seed(1)
+
+(y <- rpois(1, theta))
+
+nimble_data <- list(
+  "y" = y
+)
+
+nimble_constants <- list()
+
+
+## specify nimble model
+################################################################################
+
+nimble_model <- nimbleCode({
+  y ~ dpois(theta)
+  theta ~ dgamma(3,1)
+})
+
+nimble_monitor <- c("theta")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+nimble_inits <- list(
+  "theta" = rgamma(1,3,1)
+)
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  nimble_fit <- nimbleMCMC(
+    "code" = nimble_model, "data" = nimble_data, "constants" = nimble_constants,
+    "inits" = nimble_inits, "monitors" = nimble_monitor, "nchains" = n_chains, 
+    "niter" = n_iter, "nburnin" = n_warmup, "summary" = TRUE
+  )
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  nimble_fit$summary$all.chains
+  
+  nimble_fit_vector <- nimble_fit$samples %>% as_vector()
+  nimble_fit_object <- array(nimble_fit_vector, 
+                             dim = c(length(nimble_fit_vector) / n_chains, n_chains))
+  dim(nimble_fit_object) <- c(dim(nimble_fit_object), 1)
+  dimnames(nimble_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = nimble_monitor
+  )
+  
+  
+  nimble_fit_object %>% mcmc_areas()
+  nimble_fit_object %>% mcmc_intervals()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  nimble_fit_object %>% mcmc_acf_bar()
[…]

### 2021-04-28 large-data-parameters-models/old/hierarchical-logistic-regression/hierarchical-logistic-regression-bugs.R

@@ -1,0 +1,129 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("R2OpenBUGS")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+n <- 30L # sample size
+d <- 2L
+l <- 2L
+ll <- sapply(1:l, function(x) rep(x,n / l)) %>% as.numeric()
+
+set.seed(1)
+(x <- matrix(rnorm(n * d), nrow = n, byrow = TRUE))
+(y <- rbinom(n, 1, 0.5))
+
+bugs_data <- list(
+  "N" = n,
+  "D" = d,
+  "L" = l,
+  "ll" = ll,
+  "y" = y, 
+  "x" = x
+)
+
+
+
+## specify bugs model
+################################################################################
+
+bugs_model <- function() {
+  for (d in 1:D) {
+    mu[d] ~ dnorm(0, 0.0001)
+    tau[d] ~ dnorm(0,0.0001) %_% I(0,)
+    for (l in 1:L) {
+      beta[l,d] ~ dnorm(mu[d], tau[d])
+    }
+  }
+  for (n in 1:N) {
+    for (d in 1:D) {
+      temp[n,d] <- x[n,d] * beta[ll[n],d]
+    }
+    temp_sum[n] <- temp[n,1] + temp[n,2] 
+    logit(theta[n]) <- temp_sum[n]
+    y[n] ~ dbern(theta[n])
+  }
+}
+
+
+bugs.file <- file.path(tempdir(), "model.txt")
+write.model(bugs_model, bugs.file)
+
+bugs_monitor <- c("mu", "beta", "theta", "tau")
+
+
+## Specify path to WINE if using WINE 
+###################################################################################
+
+if (getwd() == "/Users/evanmiyakawa/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
+  WINE="/usr/local/Cellar/wine/4.0.1/bin/wine"
+  WINEPATH="/usr/local/Cellar/wine/4.0.1/bin/winepath"
+  OpenBUGS.pgm="/Users/evanmiyakawa/OpenBugs/OpenBUGS323/OpenBUGS.exe" #~
+} else {
+  WINE="/Users/evan_miyakawa1/Cellar/wine/4.0.1/bin/wine"
[…]

### 2021-04-28 large-data-parameters-models/old/hierarchical-logistic-regression/hierarchical-logistic-regression-greta.R

@@ -1,0 +1,116 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("greta")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+## generate/specify data
+################################################################################
+n <- 30L # sample size
+d <- 2L
+l <- 3L
+ll <- sapply(1:l, function(x) rep(x,n / l)) %>% as.numeric()
+
+set.seed(1)
+(x <- matrix(rnorm(n * d), nrow = n, byrow = TRUE))
+(y <- rbinom(n, 1, 0.5))
+
+bugs_data <- list(
+  "N" = n,
+  "D" = d,
+  "L" = l,
+  "ll" = ll,
+  "y" = y, 
+  "x" = x
+)
+
+x <- as_data(x)
+y <- as_data(y)
+
+
+
+## specify greta model
+################################################################################
+
+
+#### NOT COMPLETE ############
+mu <- normal(0, 1000, dim = c(1,d))
+sigma <- normal(0,1000, dim = c(d,1), truncation = c(0,Inf))
+mu_matrix <- zeros(3,2)
+mu_matrix[1,] <- mu
+mu_matrix[2,] <- mu
+mu_matrix[3,] <- mu
+sigma_matrix <- zeros(3,2)
+sigma_matrix[1,] <- sigma
+sigma_matrix[2,] <- sigma
+sigma_matrix[3,] <- sigma
+beta <- normal(mu_matrix, sigma_matrix)
+
+theta <- zeros(n,1)
+i <- 1
+for (i in 1:n) {
+  theta[i,] <- x[i,] %*% t(beta[ll[i],])
+}
+
+distribution(y) <- bernoulli(ilogit(theta))
+
+greta_model <- model(mu, beta, sigma, theta)
+
+plot(greta_model)
+
+## configure model settings
+################################################################################
+
+n_chains <- 4
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+
+## fit model
+################################################################################
+
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  greta_fit <- mcmc(
+    "model" = greta_model, "n_samples" = n_iter,
+    "warmup" = n_warmup, "chains" = n_chains
+  )
+  
+  
+  
+  ## assess fit
[…]

### 2021-04-28 large-data-parameters-models/old/hierarchical-logistic-regression/hierarchical-logistic-regression-jags.R

@@ -1,0 +1,111 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rjags"); library("runjags")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+n <- 30L # sample size
+d <- 2L
+l <- 3L
+ll <- sapply(1:l, function(x) rep(x,n / l)) %>% as.numeric()
+
+set.seed(1)
+(x <- matrix(rnorm(n * d), nrow = n, byrow = TRUE))
+(y <- rbinom(n, 1, 0.5))
+
+jags_data <- list(
+  "N" = n,
+  "D" = d,
+  "L" = l,
+  "ll" = ll,
+  "y" = y, 
+  "x" = x
+)
+
+
+
+
+## specify jags model
+################################################################################
+
+jags_model <- "
+  model{
+    for (d in 1:D) {
+      mu[d] ~ dnorm(0, 1 / (100 ^ 2))
+      tau[d] ~ dnorm(0,0.0001)  I(0, )
+      for (l in 1:L) {
+        beta[l,d] ~ dnorm(mu[d], tau[d])
+      }
+    }
+    for (n in 1:N) {
+      logit(theta[n]) <- x[n,] %*% beta[ll[n],]
+      y[n] ~ dbern(theta[n])
+    }
+  }
+"
+
+jags_monitor <- c("mu", "beta", "theta", "tau")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  jags_fit <- run.jags(
+    "model" = jags_model, "data" = jags_data, "monitor" = jags_monitor, 
+    "n.chains" = n_chains, "sample" = n_iter, "burnin" = n_warmup
+  ) 
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  jags_fit
+  
+  jags_fit_object <- jags_fit$mcmc %>% as.array()
+  dim(jags_fit_object) <- c(dim(jags_fit_object), 1)
+  dimnames(jags_fit_object) <- list(
+    "iterations" = NULL, 
[…]

### 2021-04-28 large-data-parameters-models/old/hierarchical-logistic-regression/hierarchical-logistic-regression-stan.R

@@ -1,0 +1,90 @@
+## load required packages and set basic options
+################################################################################
+
+library("here")
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rstan"); rstan_options(auto_write = TRUE)
+library("bayesplot")
+library("bench")
+
+
+
+## generate/specify data
+################################################################################
+
+n <- 30L # sample size
+d <- 2L
+l <- 3L
+ll <- sapply(1:l, function(x) rep(x,n / l)) %>% as.numeric()
+
+set.seed(1)
+(x <- matrix(rnorm(n * d), nrow = n, byrow = TRUE))
+(y <- rbinom(n, 1, 0.5))
+
+stan_data <- list(
+  "N" = n,
+  "D" = d,
+  "L" = l,
+  "ll" = ll,
+  "y" = y, 
+  "x" = x
+)
+
+
+
+## specify stan model
+################################################################################
+
+# read it in from file
+stan_file <- here("regression-models", "hierarchical-logistic-regression", "hierarchical-logistic-regression.stan")
+# file.show(stan_file)
+
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {  
+  stan_fit <- stan(
+    "file" = stan_file, "data" = stan_data, 
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
+  )
+  
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(stan_fit)$summary
+  get_posterior_mean(stan_fit)
+  stan_dens(stan_fit) + theme_bw()
+  stan_fit %>% as.array() %>% mcmc_areas()
+  
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  stan_fit %>% as.array() %>% mcmc_acf_bar()
+  stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
+  
+  # see each chain
+  stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)
+  
+}
+
+
+

### 2021-04-28 large-data-parameters-models/old/large-full-beta-binomial/hierarchical-full-beta-binomial-nimble.R

@@ -1,0 +1,108 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("nimble")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+alpha <- 2 #beta distribution alpha
+beta <- 1 #beta distribution beta
+N <- 500 # sample size
+
+set.seed(1)
+
+(n <- rbinom(N, 20, 0.5))
+(theta <- rbeta(N, alpha, beta))
+(y <- rbinom(N, n, theta))
+
+
+nimble_data <- list(
+  "y" = y,
+  "n" = n
+)
+
+nimble_constants <- list(
+  "N" = N
+)
+
+
+## specify nimble model
+################################################################################
+
+nimble_model <- nimbleCode({
+  for (i in 1:N) {
+    theta[i] ~ dbeta(alpha, beta)
+    y[i] ~ dbin(theta[i], n[i])
+  }
+  alpha ~ dgamma(1,1)
+  beta ~ dgamma(1,1)
+})
+
+nimble_monitor <- c("theta", "alpha", "beta")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+nimble_inits <- list(
+  "theta" = rbeta(N,2,1),
+  "alpha" = rgamma(1,1,1),
+  "beta" = rgamma(1,1,1)
+)
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  nimble_fit <- nimbleMCMC(
+    "code" = nimble_model, "data" = nimble_data, "constants" = nimble_constants,
+    "inits" = nimble_inits, "monitors" = nimble_monitor, "nchains" = n_chains, 
+    "niter" = n_iter, "nburnin" = n_warmup, "summary" = TRUE
+  )
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  nimble_fit$summary$all.chains
+  
+  nimble_fit_vector <- nimble_fit$samples %>% as_vector()
+  nimble_fit_object <- array(nimble_fit_vector, 
+                             dim = c(length(nimble_fit_vector) / n_chains, n_chains))
[…]

### 2021-04-28 large-data-parameters-models/old/linear-regression/linear-regression-benchmark.R

@@ -1,0 +1,46 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("bench")
+library("here")
+library("rjags"); library("runjags")
+library("R2OpenBUGS")
+library("nimble")
+library("rstan"); rstan_options(auto_write = FALSE)
+library("greta")
+options("bayes_benchmark" = TRUE)
+
+## Source code
+################################################################################
+
+source(here("large-data-parameters", "linear-regression", "linear-regression-jags.R"))
+# source(here("large-data-parameters", "linear-regression", "linear-regression-bugs.R"))
+# source(here("large-data-parameters", "linear-regression", "linear-regression-nimble.R"))
+source(here("large-data-parameters", "linear-regression", "linear-regression-stan.R"))
+source(here("large-data-parameters", "linear-regression", "linear-regression-greta.R"))
+
+
+options("bayes_benchmark" = FALSE)
+
+## Set Parameters
+################################################################################
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+## benchmarking
+###################################################################################
+
+num_iterations <- 1 # Number of times you want to benchmark each model\
+
+rds_file_location <- here("large-data-parameters", "linear-regression", "linear-regression.rds")
+
+source(here("benchmark-function.R"))
+
+# choose to include stan compilation time with argument "stan_compile = TRUE"
+run_benchmark(rds_file_location)
+run_benchmark(rds_file_location, stan_compile = TRUE)
+
+

### 2021-04-28 large-data-parameters-models/old/linear-regression/linear-regression-greta.R

@@ -1,0 +1,74 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("greta")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+## generate/specify data
+################################################################################
+
+n <- 10000L # sample size
+alpha <- -5 # intercept
+beta <- 1 # single coefficient
+sigma <- 3 # standard deviation
+
+set.seed(1)
+
+(x <- rnorm(n, 5, 4)) # observed x values
+y_hat <- alpha + beta * x
+(y <- rnorm(n,y_hat,sigma))
+
+x <- as_data(x)
+y <- as_data(y)
+
+
+
+## specify greta model
+################################################################################
+
+alpha <- normal(0, 1000)
+beta <- normal(0, 1000)
+sigma <- normal(0,1000, truncation = c(0,Inf))
+
+distribution(y) <- normal(alpha + beta * x, sigma)
+
+greta_model <- model(alpha, beta, sigma)
+
+plot(greta_model)
+
+## configure model settings
+################################################################################
+
+n_chains <- 4
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+
+## fit model
+################################################################################
+
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  greta_fit <- mcmc(
+    "model" = greta_model, "n_samples" = n_iter,
+    "warmup" = n_warmup, "chains" = n_chains
+  )
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(greta_fit)
+  
+}
+

### 2021-04-28 large-data-parameters-models/old/linear-regression/linear-regression-nimble.R

@@ -1,0 +1,108 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("nimble")
+library("bench")
+
+
+
+## generate/specify data
+################################################################################
+
+n <- 10000L # sample size
+alpha <- -5 # intercept
+beta <- 1 # single coefficient
+sigma <- 3 # standard deviation
+
+set.seed(1)
+
+(x <- rnorm(n, 5, 4)) # observed x values
+y_hat <- alpha + beta * x
+(y <- rnorm(n,y_hat,sigma))
+
+nimble_data <- list(
+  "y" = y,
+  "x" = x
+)
+
+nimble_constants <- list(
+  "N" = n
+)
+
+
+## specify nimble model
+################################################################################
+
+nimble_model <- nimbleCode({
+  for (i in 1:N) {
+    y_hat[i] <- alpha + beta * x[i]
+    y[i] ~ dnorm(y_hat[i], tau)
+  }
+  alpha ~ dnorm(0,0.0001)
+  beta ~ dnorm(0,0.0001)
+  tau ~ T(dnorm(0,0.0001), 0,)
+  sigma <- sqrt(1 / tau)
+})
+nimble_monitor = c("alpha", "beta", "sigma")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+nimble_inits <- list(
+  "alpha" = rnorm(1,0,1000),
+  "beta" = rnorm(1,0,1000),
+  "tau" = 1 / (abs(rnorm(1,0,1000)) ^ 2)
+)
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  nimble_fit <- nimbleMCMC(
+    "code" = nimble_model, "constants" = nimble_constants, "data" = nimble_data,
+    "inits" = nimble_inits, "monitors" = nimble_monitor, "nchains" = n_chains, 
+    "niter" = n_iter, "nburnin" = n_warmup, "summary" = TRUE
+  )
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  nimble_fit$summary$all.chains
+  
+  nimble_fit_vector <- nimble_fit$samples %>% as_vector()
[…]

### 2021-04-28 large-data-parameters-models/regression-models/logistic-regression/large-logistic-regression-bugs.R

@@ -1,0 +1,136 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("R2OpenBUGS")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+n <- 10L # sample size
+alpha <- -5 # intercept
+beta <- 1 # single coefficient
+
+set.seed(1)
+
+(x <- rnorm(n, 5, 1)) # observed x values
+theta_0 <- rnorm(n,alpha,0.5) + rnorm(n,beta,0.5) * x 
+(theta <- exp(theta_0) / (1 + exp(theta_0))) # generated values of bernoulli theta
+(y <- rbinom(n,1,theta))
+
+
+bugs_data <- list(
+  "n" = n,
+  "y" = y, 
+  "x" = x, 
+  "tau" = (1 / 1000)^2
+)
+bugs_data <- list(
+  "n" = n,
+  "y" = y, 
+  "x" = x
+)
+
+
+
+
+## specify bugs model
+################################################################################
+
+bugs_model <- function() {
+  for (i in 1:n) {
+    logit(theta[i]) <- alpha + beta * x[i]
+    y[i] ~ dbern(theta[i])
+  }
+  alpha ~ dnorm(0,0.0001)
+  beta ~ dnorm(0,0.0001)
+}
+
+bugs_model <- function() {
+  for (i in 1:n) {
+    log(theta[i]) <- alpha + beta * x[i]
+    y[i] ~ dpois(theta[i])
+  }
+  alpha ~ dnorm(0,0.0001)
+  beta ~ dnorm(0,0.0001)
+}
+
+
+bugs.file <- file.path(tempdir(), "model.txt")
+write.model(bugs_model, bugs.file)
+
+bugs_monitor <- c("alpha", "beta")
+
+
+## Specify path to WINE if using WINE 
+###################################################################################
+
+if (getwd() == "/Users/evanmiyakawa/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
+  WINE="/usr/local/Cellar/wine/4.0.1/bin/wine"
+  WINEPATH="/usr/local/Cellar/wine/4.0.1/bin/winepath"
+  OpenBUGS.pgm="/Users/evanmiyakawa/OpenBugs/OpenBUGS323/OpenBUGS.exe" #~
+} else {
[…]

### 2021-04-28 large-data-parameters-models/regression-models/logistic-regression/large-logistic-regression-nimble.R

@@ -1,0 +1,109 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("nimble")
+library("bench")
+library("MASS")
+
+
+
+## generate/specify data
+################################################################################
+
+n <- 500L # sample size
+alpha <- 0 # intercept
+m <- 10 # number of coefficients
+
+set.seed(1)
+
+beta <- rnorm(m, 0, 0.1) # single coefficient
+
+(x <- mvrnorm(n, rep(0, m), diag(0.5, m))) # observed x values
+
+i <- 1
+gen_theta_0 <- function(i) {
+  rnorm(1, alpha, 0.01) + sum(x %*% mvrnorm(1, beta, diag(0.01, m)))
+}
+
+(theta_0 <- 1:n %>% map_dbl(gen_theta_0))
+
+# theta_0 <- rnorm(n,alpha,0.5) + mvrnorm(1,beta, diag(m, 0.5)) %*% x 
+
+(theta <- exp(theta_0) / (1 + exp(theta_0))) # generated values of bernoulli theta
+(y <- rbinom(n,1,theta))
+
+
+nimble_data <- list(
+  "y" = y, 
+  "x" = x
+)
+
+nimble_constants <- list(
+  "n" = n,
+  "m" = m
+)
+
+
+
+
+## specify nimble model
+################################################################################
+
+nimble_model <- nimbleCode({
+  for (i in 1:n) {
+    logit(theta[i]) <- alpha + (x[i,1:m] %*% beta[1:m,1])[1,1]
+    y[i] ~ dbern(theta[i])
+  }
+  alpha ~ dnorm(0,0.0001)
+  for (i in 1:m) {
+    beta[i,1] ~ dnorm(0,0.0001)
+  }
+  
+})
+
+nimble_monitor <- c("alpha", "beta")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+nimble_inits <- list(
+  "alpha" = rnorm(1,0,1),
+  "beta" = matrix(rnorm(m,0,1), ncol = 1)
+)
+
+
+## fit model
+################################################################################
+if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+  
+  nimble_fit <- nimbleMCMC(
+
[…]

### 2021-04-28 large-data-parameters-models/simple-models/continuous/normal/normal-prior/large-normal-normal-benchmark.R

@@ -1,0 +1,60 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("bench")
+library("here")
+library("rjags"); library("runjags")
+library("R2OpenBUGS")
+library("nimble")
+library("rstan"); rstan_options(auto_write = FALSE)
+options("bayes_benchmark" = TRUE)
+
+## Source code
+################################################################################
+
+source(here("simple-models", "continuous", "normal", "normal-prior", "normal-normal-jags.R"))
+source(here("simple-models", "continuous", "normal", "normal-prior", "normal-normal-bugs.R"))
+source(here("simple-models", "continuous", "normal", "normal-prior", "normal-normal-nimble.R"))
+source(here("simple-models", "continuous", "normal", "normal-prior", "normal-normal-stan.R"))
+source(here("simple-models", "continuous", "normal", "normal-prior", "normal-normal-greta.R"))
+
+options("bayes_benchmark" = FALSE)
+
+## Set Parameters
+################################################################################
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+## benchmarking
+###################################################################################
+
+num_iterations <- 50 # Number of times you want to benchmark each model\
+
+rds_file_location <- here("simple-models", "continuous", "normal", "normal-prior", "normal-normal.rds")
+
+source(here("benchmark-function.R"))
+
+# choose to include stan compilation time with argument "stan_compile = TRUE"
+run_benchmark(rds_file_location)
+run_benchmark(rds_file_location, stan_compile = TRUE)
+
+## benchmarking accuracy
+###################################################################################
+
+num_iterations <- 20 # Number of times you want to benchmark each model\
+
+source(here("benchmark-accuracy-function.R"))
+
+true_dist <- distr::Norm(0.9031468, sqrt(0.2))
+
+run_accuracy_benchmark(true_dist,num_iterations, n_iter / 2, n_warmup / 2)
+
+
+
+
+
+
+

### 2021-04-28 large-data-parameters-models/simple-models/continuous/normal/normal-prior/large-normal-normal-greta.R

@@ -1,0 +1,87 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("greta")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+## generate/specify data
+################################################################################
+
+mu <- 50    # normal mu
+sigma <- 20 # normal sigma
+n <- 1e5   # sample size
+
+set.seed(1)
+
+(y <- rnorm(n, mu, sigma))
+
+y <- as_data(y)
+n <- as_data(n)
+
+## specify greta model
+################################################################################
+
+mu <- normal(0,1000)
+
+distribution(y) <- normal(mu,sigma)
+
+greta_model <- model(mu)
+
+plot(greta_model)
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  greta_fit <- mcmc(
+    "model" = greta_model, 
+    "n_samples" = n_iter,
+    "warmup" = n_warmup,
+    "chains" = n_chains
+  )
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(greta_fit)
+  
+  greta_fit_object <- greta_fit %>% as.array()
+  dim(greta_fit_object) <- c(dim(greta_fit_object), 1)
+  dimnames(greta_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = c("theta")
+  )
+  
+  
+  greta_fit_object %>% mcmc_areas()
+  greta_fit_object %>% mcmc_intervals()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  greta_fit_object %>% mcmc_acf_bar()
+  greta_fit_object %>% mcmc_trace()
+  greta_fit_object %>% mcmc_hist_by_chain()
+  
+}
+

### 2021-04-28 large-data-parameters-models/simple-models/continuous/normal/normal-prior/large-normal-normal-jags.R

@@ -1,0 +1,97 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rjags"); library("runjags")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+mu <- 50    # normal mu
+sigma <- 20 # normal sigma
+n <- 1e5   # sample size
+
+set.seed(1)
+
+(y <- rnorm(n, mu, sigma))
+
+jags_data <- list(
+  "y" = y,
+  "tau" = 1 / sigma ^2,
+  "N" = n
+)
+
+
+## specify jags model
+################################################################################
+
+jags_model <- "
+  model{
+    for (i in 1:N) {
+      y[i] ~ dnorm(mu, tau)
+    }
+    mu ~ dnorm(0, 0.0001)
+  }
+"
+
+jags_monitor <- c("mu")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  jags_fit <- run.jags(
+    "model" = jags_model, "data" = jags_data, "monitor" = jags_monitor, 
+    "n.chains" = n_chains, "sample" = n_iter, "burnin" = n_warmup
+  ) 
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  jags_fit
+  
+  jags_fit_object <- jags_fit$mcmc %>% as.array()
+  dim(jags_fit_object) <- c(dim(jags_fit_object), 1)
+  dimnames(jags_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = monitor
+  )
+  
+  
+  jags_fit_object %>% mcmc_areas()
+  jags_fit_object %>% mcmc_intervals()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  jags_fit_object %>% mcmc_acf_bar()
+  jags_fit_object %>% mcmc_trace()
+  jags_fit_object %>% mcmc_hist_by_chain()
+  
+  
+  
+  
+}
+

### 2021-04-28 large-data-parameters-models/simple-models/continuous/normal/normal-prior/large-normal-normal-nimble.R

@@ -1,0 +1,79 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("nimble")
+library("bench")
+
+
+
+## generate/specify data
+################################################################################
+
+mu <- 50    # normal mu
+sigma <- 20 # normal sigma
+n <- 1e5   # sample size
+
+set.seed(1)
+
+(y <- rnorm(n, mu, sigma))
+
+nimble_data <- list(
+  "y" = y
+)
+
+nimble_constants <- list(
+  "tau" = 1 / sigma ^2,
+  "N" = n
+)
+
+
+## specify nimble model
+################################################################################
+
+nimble_model <- nimbleCode({
+  for (i in 1:N) {
+    y[i] ~ dnorm(mu, tau)
+  }
+  mu ~ dnorm(0,0.0001)
+})
+nimble_monitor = c("mu")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+nimble_inits <- list(
+  "mu" = rnorm(1)
+)
+
+
+## fit model
+################################################################################
+if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
+  
+  nimble_fit <- nimbleMCMC(
+    "code" = nimble_model, "constants" = nimble_constants, "data" = nimble_data,
+    "inits" = nimble_inits, "monitors" = nimble_monitor, "nchains" = n_chains, 
+    "niter" = n_iter, "nburnin" = n_warmup, "summary" = TRUE
+  )
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  nimble_fit$summary$all.chains
+  
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+}  
+
+

### 2021-04-28 large-data-parameters-models/simple-models/continuous/normal/normal-prior/large-normal-normal-stan.R

@@ -1,0 +1,86 @@
+## load required packages and set basic options
+################################################################################
+
+library("here")
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rstan"); rstan_options(auto_write = TRUE)
+library("bayesplot")
+library("bench")
+
+
+
+## generate/specify data
+################################################################################
+
+mu <- 50    # normal mu
+sigma <- 20 # normal sigma
+n <- 1e5   # sample size
+
+set.seed(1)
+
+(y <- rnorm(n, mu, sigma))
+
+stan_data <- list(
+  "y" = y,
+  "sigma" = sigma,
+  "N" = n
+)
+
+
+
+## specify stan model
+################################################################################
+
+# read it in from file
+stan_file <- here("large-data-parameters-models", "simple-models", "continuous", "normal", "normal-prior", "large-normal-normal.stan")
+
+# file.show(stan_file)
+
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {  
+  stan_fit <- stan(
+    "file" = stan_file, "data" = stan_data, 
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
+  )
+  
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(stan_fit)$summary
+  get_posterior_mean(stan_fit)
+  stan_dens(stan_fit) + theme_bw()
+  stan_fit %>% as.array() %>% mcmc_areas()
+  
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  stan_fit %>% as.array() %>% mcmc_acf_bar()
+  stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
+  
+  # see each chain
+  stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)
+  
+}
+
+

### 2021-04-28 large-data-parameters-models/simple-models/discrete/binomial/beta-prior/large-beta-binomial-bugs.R

@@ -1,0 +1,106 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("R2OpenBUGS")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+p <- .25 # binomial p
+n <-  10 # binomial n
+
+set.seed(1)
+
+(y <- rbinom(1, n, p))
+
+bugs_data <- list(
+  "y" = y,
+  "N" = n
+)
+
+
+
+## specify bugs model
+################################################################################
+
+bugs_model <- function() {
+  y ~ dbin(p,N)
+  p ~ dbeta(1,1)
+}
+
+
+bugs.file <- file.path(tempdir(), "model.txt")
+write.model(bugs_model, bugs.file)
+
+bugs_monitor <- "p"
+
+
+## Specify path to WINE if using WINE 
+###################################################################################
+
+if (getwd() == "/Users/evanmiyakawa/hubiC/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
+  WINE="/usr/local/Cellar/wine/4.0.1/bin/wine"
+  WINEPATH="/usr/local/Cellar/wine/4.0.1/bin/winepath"
+  OpenBUGS.pgm="/Users/evanmiyakawa/OpenBugs/OpenBUGS323/OpenBUGS.exe" #~
+} else {
+  WINE="/Users/evan_miyakawa1/Cellar/wine/4.0.1/bin/wine"
+  WINEPATH="/Users/evan_miyakawa1/Cellar/wine/4.0.1/bin/winepath"
+  OpenBUGS.pgm="/Users/evan_miyakawa1/OpenBugs/OpenBUGS323/OpenBUGS.exe"
+}
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  bugs_fit <- bugs(
+    "model.file" = bugs.file, "data" = bugs_data, "parameters.to.save" = bugs_monitor, 
[…]

### 2021-04-28 large-data-parameters-models/simple-models/discrete/binomial/beta-prior/large-beta-binomial-greta.R

@@ -1,0 +1,88 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("greta")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+## generate/specify data
+################################################################################
+
+p <- .25 # binomial p
+n <-  10 # binomial n
+N <-  1e3 # sample size
+
+set.seed(1)
+
+(y <- rbinom(N, n, p))
+
+y <- as_data(y)
+n <- as_data(n)
+
+
+## specify greta model
+################################################################################
+
+p <- beta(1,1)
+
+distribution(y) <- binomial(n,p)
+
+greta_model <- model(p)
+
+plot(greta_model)
+
+## configure model settings
+################################################################################
+
+n_chains <- 50L
+n_iter <- (1e4L * 4) / n_chains
+n_warmup <- 1e3L
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  greta_fit <- mcmc(
+    "model" = greta_model, 
+    "n_samples" = n_iter,
+    "warmup" = n_warmup,
+    "chains" = n_chains
+  )
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(greta_fit)
+  
+  greta_fit_object <- greta_fit %>% as.array()
+  dim(greta_fit_object) <- c(dim(greta_fit_object), 1)
+  dimnames(greta_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = c("theta")
+  )
+  
+  
+  greta_fit_object %>% mcmc_areas()
+  greta_fit_object %>% mcmc_intervals()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  greta_fit_object %>% mcmc_acf_bar()
+  greta_fit_object %>% mcmc_trace()
+  greta_fit_object %>% mcmc_hist_by_chain()
+  
+}
+

### 2021-04-28 large-data-parameters-models/simple-models/discrete/binomial/beta-prior/large-beta-binomial-jags.R

@@ -1,0 +1,98 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rjags"); library("runjags")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+p <- .25 # binomial p
+n <-  10 # binomial n
+N <-  1e5 # sample size
+
+set.seed(1)
+
+(y <- rbinom(N, n, p))
+
+jags_data <- list(
+  "y" = y,
+  "n" = n,
+  "N" = N
+)
+
+
+
+## specify jags model
+################################################################################
+
+jags_model <- "
+  model{
+    for (i in 1:N) {
+      y[i] ~ dbin(p,n)
+    }
+    p ~ dbeta(1,1)
+  }
+"
+
+jags_monitor <- c("p")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  jags_fit <- run.jags(
+    "model" = jags_model, "data" = jags_data, "monitor" = jags_monitor, 
+    "n.chains" = n_chains, "sample" = n_iter, "burnin" = n_warmup
+  ) 
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  jags_fit
+  
+  jags_fit_object <- jags_fit$mcmc %>% as.array()
+  dim(jags_fit_object) <- c(dim(jags_fit_object), 1)
+  dimnames(jags_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = monitor
+  )
+  
+  
+  jags_fit_object %>% mcmc_areas()
+  jags_fit_object %>% mcmc_intervals()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  jags_fit_object %>% mcmc_acf_bar()
+  jags_fit_object %>% mcmc_trace()
+  jags_fit_object %>% mcmc_hist_by_chain()
+  
+  
+  
+  
+}
+

### 2021-04-28 large-data-parameters-models/simple-models/discrete/binomial/beta-prior/large-beta-binomial-stan.R

@@ -1,0 +1,86 @@
+## load required packages and set basic options
+################################################################################
+
+library("here")
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rstan"); rstan_options(auto_write = TRUE)
+library("bayesplot")
+
+
+
+## generate/specify data
+################################################################################
+
+p <- .25 # binomial p
+n <-  10 # binomial n
+N <-  1e5 # sample size
+
+set.seed(1)
+
+(y <- rbinom(N, n, p))
+
+stan_data <- list(
+  "n" = n,
+  "y" = y,
+  "N" = N
+)
+
+
+
+## specify stan model
+################################################################################
+
+# read it in from file
+stan_file <- here("large-data-parameters-models", "simple-models", "discrete", "binomial", "beta-prior", "large-beta-binomial.stan")
+
+# file.show(stan_file)
+
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {  
+  stan_fit <- stan(
+    "file" = stan_file, "data" = stan_data, 
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
+  )
+  
+  
+  
+  
+  ## assess fit
+  ################################################################################
+  
+  summary(stan_fit)$summary
+  get_posterior_mean(stan_fit)
+  stan_dens(stan_fit) + theme_bw()
+  stan_fit %>% as.array() %>% mcmc_areas()
+  
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  stan_fit %>% as.array() %>% mcmc_acf_bar()
+  stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
+  
+  # see each chain
+  stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)
+  
+}
+
+
+

### 2021-04-28 large-data-parameters-models/time-series/ar-model/large-ar-model-benchmark.R

@@ -1,0 +1,45 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("bench")
+library("here")
+library("rjags"); library("runjags")
+library("R2OpenBUGS")
+library("nimble")
+library("rstan"); rstan_options(auto_write = FALSE)
+library("greta")
+options("bayes_benchmark" = TRUE)
+
+## Source code
+################################################################################
+
+source(here("time-series", "ar-model", "ar-model-jags.R"))
+source(here("time-series", "ar-model", "ar-model-bugs.R"))
+# source(here("time-series", "ar-model", "ar-model-nimble.R"))
+source(here("time-series", "ar-model", "ar-model-stan.R"))
+# source(here("time-series", "ar-model", "ar-model-greta.R"))
+
+options("bayes_benchmark" = FALSE)
+
+## Set Parameters
+################################################################################
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+## benchmarking
+###################################################################################
+
+num_iterations <- 5 # Number of times you want to benchmark each model\
+
+rds_file_location <- here("time-series", "ar-model.rds")
+
+source(here("benchmark-function.R"))
+
+# choose to include stan compilation time with argument "stan_compile = TRUE"
+run_benchmark(rds_file_location)
+run_benchmark(rds_file_location, stan_compile = TRUE)
+
+

### 2021-04-28 large-data-parameters-models/time-series/ar-model/large-ar-model-bugs.R

@@ -1,0 +1,123 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("R2OpenBUGS")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+N <- 200L # sample size
+alpha <- 0 # ar(1) mean
+beta <- c(-0.6) # ar(1) coefficient
+sigma <- 1 # ar(1) standard deviation
+
+
+set.seed(1)
+(y <- arima.sim(model = list(ar = beta), sd = sigma, n = N) %>% as.numeric())
+
+t <- 1:length(y)
+ggplot(data.frame(t = t, y = y), aes(t,y)) + geom_line()
+
+
+bugs_data <- list(
+  "N" = N,
+  "y" = y
+)
+
+
+
+
+## specify bugs model
+################################################################################
+
+bugs_model <- function() {
+  for (i in 2:N) {
+    y_hat[i] <- alpha + beta * y[i-1]
+    y[i] ~ dnorm(y_hat[i], tau)
+  }
+  alpha ~ dnorm(0,0.0001)
+  beta ~ dnorm(0,0.0001)
+  tau ~ dnorm(0,0.0001) %_% I(0, )
+  sigma <- sqrt(1 / tau)
+}
+
+
+bugs.file <- file.path(tempdir(), "model.txt")
+write.model(bugs_model, bugs.file)
+
+bugs_monitor <- c("alpha", "beta", "sigma")
+
+
+## Specify path to WINE if using WINE 
+###################################################################################
+
+if (getwd() == "/Users/evanmiyakawa/Git Projects/Bayes-Computational-Methods/Bayes-Computational-Methods") {
+  WINE="/usr/local/Cellar/wine/4.0.1/bin/wine"
+  WINEPATH="/usr/local/Cellar/wine/4.0.1/bin/winepath"
+  OpenBUGS.pgm="/Users/evanmiyakawa/OpenBugs/OpenBUGS323/OpenBUGS.exe" #~
+} else {
+  WINE="/Users/evan_miyakawa1/Cellar/wine/4.0.1/bin/wine"
+  WINEPATH="/Users/evan_miyakawa1/Cellar/wine/4.0.1/bin/winepath"
+  OpenBUGS.pgm="/Users/evan_miyakawa1/OpenBugs/OpenBUGS323/OpenBUGS.exe"
+}
+
+
+
+
+## configure model settings
[…]

### 2021-04-28 large-data-parameters-models/time-series/ar-model/large-ar-model-jags.R

@@ -1,0 +1,116 @@
+## load required packages and set basic options
+################################################################################
+
+library("tidyverse"); theme_set(theme_minimal())
+library("parallel"); options(mc.cores = detectCores())
+library("rjags"); library("runjags")
+library("bayesplot")
+library("bench")
+library("here")
+
+
+
+## generate/specify data
+################################################################################
+
+N <- 1e4L # sample size
+alpha <- 0 # ar(1) mean
+M <- 6 #Number of coefficients
+beta <- c(-0.8, -0.4, 0.2, 0.15, 0.05, -0.03) # ar(6) coefficients
+sigma <- rnorm(length(beta), 1, sd = 0.2) # ar(1) standard deviation
+
+
+set.seed(1)
+(y <- arima.sim(model = list(ar = beta), sd = sigma, n = N)  %>% as.numeric())
+y <- c(
+  rep(0, length(beta)),
+  y
+)
+
+t <- 1:length(y)
+ggplot(data.frame(t = t, y = y), aes(t,y)) + geom_line()
+
+
+jags_data <- list(
+  "N" = length(y),
+  "M" = M,
+  "y" = y
+)
+
+
+## specify jags model
+################################################################################
+
+### try step function instead of ifelse
+
+jags_model <- "
+  model{
+    for (i in (M+1):N) {
+      y_hat[i] <- alpha + sum(beta[1:M] * y[(i-M):(i-1)])
+      y[i] ~ dnorm(y_hat[i], tau)
+    }
+    alpha ~ dnorm(0,0.0001)
+    for (j in 1:M) {
+      beta[j] ~ dnorm(0,0.0001)
+    }
+    tau ~ dnorm(0,0.0001)  I(0, )
+    sigma <- sqrt(1 / tau)
+  }
+"
+
+jags_monitor <- c("alpha", "beta", "sigma")
+
+
+## configure model settings
+################################################################################
+
+n_chains <- 4L
+n_iter <- 1e4L
+n_warmup <- 1e3L
+
+
+
+## fit model
+################################################################################
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {
+  
+  
+  jags_fit <- run.jags(
+    "model" = jags_model, "data" = jags_data, "monitor" = jags_monitor, 
+    "n.chains" = n_chains, "sample" = n_iter, "burnin" = n_warmup
+  ) 
+  
+  
+  
+  ## assess fit
[…]

### 2021-04-28 regression-models/hierarchical-logistic-regression/hierarchical-logistic-regression-bugs.R

@@ -103,9 +103,25 @@ if (!currently_benchmarking()) {
   
   bugs_fit$summary
   
+  bugs_fit_object <- bugs_fit$sims.array[,,1]
+  dim(bugs_fit_object) <- c(dim(bugs_fit_object), 1)
+  dimnames(bugs_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = bugs_monitor
+  )
+  
+  
+  bugs_fit_object %>% mcmc_areas()
+  bugs_fit_object %>% mcmc_intervals()
+  
   
   ## assess convergence issues 
   ###################################################################################
+  
+  bugs_fit_object %>% mcmc_acf_bar()
+  bugs_fit_object %>% mcmc_trace()
+  bugs_fit_object %>% mcmc_hist_by_chain()
   
 }
 

### 2021-04-28 regression-models/hierarchical-logistic-regression/hierarchical-logistic-regression-stan.R

@@ -52,8 +52,9 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
-  
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {  
   stan_fit <- stan(
     "file" = stan_file, "data" = stan_data, 
     "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
@@ -68,7 +69,7 @@ if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])
   summary(stan_fit)$summary
   get_posterior_mean(stan_fit)
   stan_dens(stan_fit) + theme_bw()
-  stan_fit %>% as.array() %>% bayesplot::mcmc_dens()
+  stan_fit %>% as.array() %>% mcmc_areas()
   
   
   
@@ -76,8 +77,9 @@ if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])
   ###################################################################################
   
   stan_fit %>% as.array() %>% mcmc_acf_bar()
-  stan_fit %>% as.array() %>% mcmc_pairs()
   stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
   
   # see each chain
   stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)

### 2021-04-28 regression-models/linear-regression/linear-regression-bugs.R

@@ -93,9 +93,25 @@ if (!currently_benchmarking()) {
   
   bugs_fit$summary
   
+  bugs_fit_object <- bugs_fit$sims.array[,,1]
+  dim(bugs_fit_object) <- c(dim(bugs_fit_object), 1)
+  dimnames(bugs_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = bugs_monitor
+  )
+  
+  
+  bugs_fit_object %>% mcmc_areas()
+  bugs_fit_object %>% mcmc_intervals()
+  
   
   ## assess convergence issues 
   ###################################################################################
+  
+  bugs_fit_object %>% mcmc_acf_bar()
+  bugs_fit_object %>% mcmc_trace()
+  bugs_fit_object %>% mcmc_hist_by_chain()
   
 }
 

### 2021-04-28 regression-models/logistic-regression/logistic-regression-bugs.R

@@ -110,9 +110,25 @@ if (!currently_benchmarking()) {
   
   bugs_fit$summary
   
+  bugs_fit_object <- bugs_fit$sims.array[,,1]
+  dim(bugs_fit_object) <- c(dim(bugs_fit_object), 1)
+  dimnames(bugs_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = bugs_monitor
+  )
+  
+  
+  bugs_fit_object %>% mcmc_areas()
+  bugs_fit_object %>% mcmc_intervals()
+  
   
   ## assess convergence issues 
   ###################################################################################
+  
+  bugs_fit_object %>% mcmc_acf_bar()
+  bugs_fit_object %>% mcmc_trace()
+  bugs_fit_object %>% mcmc_hist_by_chain()
   
 }
 

### 2021-04-28 regression-models/logistic-regression/logistic-regression-stan.R

@@ -69,7 +69,7 @@ if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])
   summary(stan_fit)$summary
   get_posterior_mean(stan_fit)
   stan_dens(stan_fit) + theme_bw()
-  stan_fit %>% as.array() %>% bayesplot::mcmc_dens()
+  stan_fit %>% as.array() %>% mcmc_areas()
   
   
   
@@ -77,8 +77,9 @@ if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])
   ###################################################################################
   
   stan_fit %>% as.array() %>% mcmc_acf_bar()
-  stan_fit %>% as.array() %>% mcmc_pairs()
   stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
   
   # see each chain
   stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)

### 2021-04-28 regression-models/poisson-regression/poisson-regression-greta.R

@@ -70,5 +70,25 @@ if (!currently_benchmarking()) {
   
   summary(greta_fit)
   
+  greta_fit_object <- greta_fit %>% as.array()
+  dim(greta_fit_object) <- c(dim(greta_fit_object), 1)
+  dimnames(greta_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = c("theta")
+  )
+  
+  
+  greta_fit_object %>% mcmc_areas()
+  greta_fit_object %>% mcmc_intervals()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  greta_fit_object %>% mcmc_acf_bar()
+  greta_fit_object %>% mcmc_trace()
+  greta_fit_object %>% mcmc_hist_by_chain()
+  
 }
 

### 2021-04-28 simple-models/continuous/exponential/gamma-prior/exponential-gamma-bugs.R

@@ -81,9 +81,25 @@ if (!currently_benchmarking()) {
   
   bugs_fit$summary
   
+  bugs_fit_object <- bugs_fit$sims.array[,,1]
+  dim(bugs_fit_object) <- c(dim(bugs_fit_object), 1)
+  dimnames(bugs_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = bugs_monitor
+  )
+  
+  
+  bugs_fit_object %>% mcmc_areas()
+  bugs_fit_object %>% mcmc_intervals()
+  
   
   ## assess convergence issues 
   ###################################################################################
+  
+  bugs_fit_object %>% mcmc_acf_bar()
+  bugs_fit_object %>% mcmc_trace()
+  bugs_fit_object %>% mcmc_hist_by_chain()
   
 }
 

### 2021-04-28 simple-models/continuous/exponential/gamma-prior/exponential-gamma-jags.R

@@ -77,7 +77,8 @@ if (!currently_benchmarking()) {
   )
   
   
-  jags_fit_object %>% bayesplot::mcmc_dens()
+  jags_fit_object %>% mcmc_areas()
+  jags_fit_object %>% mcmc_intervals()
   
   
   ## assess convergence issues 
@@ -85,6 +86,7 @@ if (!currently_benchmarking()) {
   
   jags_fit_object %>% mcmc_acf_bar()
   jags_fit_object %>% mcmc_trace()
+  jags_fit_object %>% mcmc_hist_by_chain()
   
   
   

### 2021-04-28 simple-models/continuous/normal/normal-prior/normal-normal-benchmark.R

@@ -48,7 +48,8 @@ num_iterations <- 20 # Number of times you want to benchmark each model\
 
 source(here("benchmark-accuracy-function.R"))
 
-true_dist <- distr::Norm(0.9031468, sqrt(0.2))
+true_mean <- 0.9031468
+true_dist <- distr::Norm(true_mean, sqrt(0.2))
 
 run_accuracy_benchmark(true_dist,num_iterations, n_iter / 2, n_warmup / 2)
 

### 2021-04-28 simple-models/continuous/normal/normal-prior/normal-normal-jags.R

@@ -88,7 +88,8 @@ if (!currently_benchmarking()) {
   )
   
   
-  jags_fit_object %>% bayesplot::mcmc_dens()
+  jags_fit_object %>% mcmc_areas()
+  jags_fit_object %>% mcmc_intervals()
   
   
   ## assess convergence issues 
@@ -96,6 +97,7 @@ if (!currently_benchmarking()) {
   
   jags_fit_object %>% mcmc_acf_bar()
   jags_fit_object %>% mcmc_trace()
+  jags_fit_object %>% mcmc_hist_by_chain()
   
   
   

### 2021-04-28 simple-models/discrete/bernoulli/beta-prior/beta-bernoulli-nimble.R

@@ -63,10 +63,27 @@ if (!currently_benchmarking()) {
   
   nimble_fit$summary$all.chains
   
+  nimble_fit_vector <- nimble_fit$samples %>% as_vector()
+  nimble_fit_object <- array(nimble_fit_vector, 
+                             dim = c(length(nimble_fit_vector) / n_chains, n_chains))
+  dim(nimble_fit_object) <- c(dim(nimble_fit_object), 1)
+  dimnames(nimble_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = nimble_monitor
+  )
+  
+  
+  nimble_fit_object %>% mcmc_areas()
+  nimble_fit_object %>% mcmc_intervals()
   
   
   ## assess convergence issues 
   ###################################################################################
+  
+  nimble_fit_object %>% mcmc_acf_bar()
+  nimble_fit_object %>% mcmc_trace()
+  nimble_fit_object %>% mcmc_hist_by_chain()
   
 }  
 

### 2021-04-28 simple-models/discrete/bernoulli/beta-prior/beta-bernoulli-stan.R

@@ -51,8 +51,7 @@ if (!currently_benchmarking()) {
   
   stan_fit <- stan(
     "file" = stan_file, "data" = stan_data, 
-    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup, 
-    "control" = list("adapt_delta" = 0.99)
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
   )
   
   
@@ -64,7 +63,7 @@ if (!currently_benchmarking()) {
   summary(stan_fit)$summary
   get_posterior_mean(stan_fit)
   stan_dens(stan_fit) + theme_bw()
-  stan_fit %>% as.array() %>% bayesplot::mcmc_dens()
+  stan_fit %>% as.array() %>% mcmc_areas()
   
   
   
@@ -72,8 +71,9 @@ if (!currently_benchmarking()) {
   ###################################################################################
   
   stan_fit %>% as.array() %>% mcmc_acf_bar()
-  stan_fit %>% as.array() %>% mcmc_pairs()
   stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
   
   # see each chain
   stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)

### 2021-04-28 simple-models/discrete/binomial/beta-prior/beta-binomial-bugs.R

@@ -82,9 +82,25 @@ if (!currently_benchmarking()) {
   
   bugs_fit$summary
   
+  bugs_fit_object <- bugs_fit$sims.array[,,1]
+  dim(bugs_fit_object) <- c(dim(bugs_fit_object), 1)
+  dimnames(bugs_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = bugs_monitor
+  )
+  
+  
+  bugs_fit_object %>% mcmc_areas()
+  bugs_fit_object %>% mcmc_intervals()
+  
   
   ## assess convergence issues 
   ###################################################################################
+  
+  bugs_fit_object %>% mcmc_acf_bar()
+  bugs_fit_object %>% mcmc_trace()
+  bugs_fit_object %>% mcmc_hist_by_chain()
   
 }
 

### 2021-04-28 simple-models/discrete/binomial/beta-prior/beta-binomial-nimble.R

@@ -63,10 +63,27 @@ if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])
   
   nimble_fit$summary$all.chains
   
+  nimble_fit_vector <- nimble_fit$samples %>% as_vector()
+  nimble_fit_object <- array(nimble_fit_vector, 
+                             dim = c(length(nimble_fit_vector) / n_chains, n_chains))
+  dim(nimble_fit_object) <- c(dim(nimble_fit_object), 1)
+  dimnames(nimble_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = nimble_monitor
+  )
+  
+  
+  nimble_fit_object %>% mcmc_areas()
+  nimble_fit_object %>% mcmc_intervals()
   
   
   ## assess convergence issues 
   ###################################################################################
+  
+  nimble_fit_object %>% mcmc_acf_bar()
+  nimble_fit_object %>% mcmc_trace()
+  nimble_fit_object %>% mcmc_hist_by_chain()
   
 }  
 

### 2021-04-28 simple-models/discrete/binomial/beta-prior/beta-binomial-stan.R

@@ -46,12 +46,12 @@ n_warmup <- 1e3L
 
 ## fit model
 ################################################################################
-if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])) {
-  
+source(here("currently-benchmarking.R"))
+
+if (!currently_benchmarking()) {  
   stan_fit <- stan(
     "file" = stan_file, "data" = stan_data, 
-    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup, 
-    "control" = list("adapt_delta" = 0.99)
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
   )
   
   
@@ -63,7 +63,7 @@ if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])
   summary(stan_fit)$summary
   get_posterior_mean(stan_fit)
   stan_dens(stan_fit) + theme_bw()
-  stan_fit %>% as.array() %>% bayesplot::mcmc_dens()
+  stan_fit %>% as.array() %>% mcmc_areas()
   
   
   
@@ -71,8 +71,9 @@ if (is.null(options()[["bayes_benchmark"]]) || !(options()[["bayes_benchmark"]])
   ###################################################################################
   
   stan_fit %>% as.array() %>% mcmc_acf_bar()
-  stan_fit %>% as.array() %>% mcmc_pairs()
   stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
   
   # see each chain
   stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)

### 2021-04-28 simple-models/discrete/poisson/gamma-prior/poisson-gamma-benchmark.R

@@ -6,20 +6,21 @@ library("parallel"); options(mc.cores = detectCores())
 library("bench")
 library("here")
 library("rjags"); library("runjags")
-library("R2OpenBUGS")
-library("nimble")
+# library("R2OpenBUGS")
 library("rstan"); rstan_options(auto_write = FALSE)
 library("greta")
+library("nimble")
 options("bayes_benchmark" = TRUE)
 
 ## Source code
 ################################################################################
 
 source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-jags.R"))
-source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-bugs.R"))
-source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-nimble.R"))
+# source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-bugs.R"))
 source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-stan.R"))
 source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-greta.R"))
+source(here("simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-nimble.R"))
+
 
 options("bayes_benchmark" = FALSE)
 

### 2021-04-28 simple-models/discrete/poisson/gamma-prior/poisson-gamma-jags.R

@@ -74,7 +74,8 @@ if (!currently_benchmarking()) {
   )
   
   
-  jags_fit_object %>% bayesplot::mcmc_dens()
+  jags_fit_object %>% mcmc_areas()
+  jags_fit_object %>% mcmc_intervals()
   
   
   ## assess convergence issues 
@@ -82,7 +83,7 @@ if (!currently_benchmarking()) {
   
   jags_fit_object %>% mcmc_acf_bar()
   jags_fit_object %>% mcmc_trace()
-  
+  jags_fit_object %>% mcmc_hist_by_chain()
   
   
   

### 2021-04-28 simple-models/discrete/poisson/gamma-prior/poisson-gamma-stan.R

@@ -49,8 +49,7 @@ source(here("currently-benchmarking.R"))
 if (!currently_benchmarking()) {
   stan_fit <- stan(
     "file" = stan_file, "data" = stan_data, 
-    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup, 
-    "control" = list("adapt_delta" = 0.99)
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
   )
   
   
@@ -62,7 +61,7 @@ if (!currently_benchmarking()) {
   summary(stan_fit)$summary
   get_posterior_mean(stan_fit)
   stan_dens(stan_fit) + theme_bw()
-  stan_fit %>% as.array() %>% bayesplot::mcmc_dens()
+  stan_fit %>% as.array() %>% mcmc_areas()
   
   
   
@@ -70,8 +69,9 @@ if (!currently_benchmarking()) {
   ###################################################################################
   
   stan_fit %>% as.array() %>% mcmc_acf_bar()
-  stan_fit %>% as.array() %>% mcmc_pairs()
   stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
   
   # see each chain
   stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)

### 2021-04-28 survival-models/cox-exponential-greta.R

@@ -69,5 +69,25 @@ if (!currently_benchmarking()) {
   
   summary(greta_fit)
   
+  greta_fit_object <- greta_fit %>% as.array()
+  dim(greta_fit_object) <- c(dim(greta_fit_object), 1)
+  dimnames(greta_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = c("theta")
+  )
+  
+  
+  greta_fit_object %>% mcmc_areas()
+  greta_fit_object %>% mcmc_intervals()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  greta_fit_object %>% mcmc_acf_bar()
+  greta_fit_object %>% mcmc_trace()
+  greta_fit_object %>% mcmc_hist_by_chain()
+  
 }
 

### 2021-04-28 survival-models/cox-exponential-stan.R

@@ -79,8 +79,7 @@ source(here("currently-benchmarking.R"))
 if (!currently_benchmarking()) {
   stan_fit <- stan(
     "file" = stan_file, "data" = stan_data,# "init" = stan_inits,
-    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup, 
-    "control" = list("adapt_delta" = 0.99)
+    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
   )
   
   
@@ -92,7 +91,7 @@ if (!currently_benchmarking()) {
   summary(stan_fit)$summary
   get_posterior_mean(stan_fit)
   stan_dens(stan_fit) + theme_bw()
-  stan_fit %>% as.array() %>% bayesplot::mcmc_dens()
+  stan_fit %>% as.array() %>% mcmc_areas()
   
   
   
@@ -100,8 +99,9 @@ if (!currently_benchmarking()) {
   ###################################################################################
   
   stan_fit %>% as.array() %>% mcmc_acf_bar()
-  stan_fit %>% as.array() %>% mcmc_pairs()
   stan_fit %>% as.array() %>% mcmc_trace()
+  stan_fit %>% as.array() %>% mcmc_hist_by_chain()
+  
   
   # see each chain
   stan_fit %>% rstan::extract(permuted = FALSE, inc_warmup = TRUE)

### 2021-04-28 time-series/ar-model/ar-model-greta.R

@@ -22,7 +22,7 @@ set.seed(1)
 (y <- arima.sim(model = list(ar = beta), sd = sigma, n = N) %>% as.numeric())
 
 t <- 1:length(y)
-ggplot(data.frame(t = t, y = y), aes(t,y)) + geom_line()
+# ggplot(data.frame(t = t, y = y), aes(t,y)) + geom_line()
 
 y_1 <- y[1:(length(y) - 1)] # data excluding the last point
 z_1 <- y[2:length(y)] # data excluding the first point
@@ -69,6 +69,26 @@ if (!currently_benchmarking()) {
   ################################################################################
   
   summary(greta_fit)
+  
+  greta_fit_object <- greta_fit %>% as.array()
+  dim(greta_fit_object) <- c(dim(greta_fit_object), 1)
+  dimnames(greta_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = c("theta")
+  )
+  
+  
+  greta_fit_object %>% mcmc_areas()
+  greta_fit_object %>% mcmc_intervals()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  greta_fit_object %>% mcmc_acf_bar()
+  greta_fit_object %>% mcmc_trace()
+  greta_fit_object %>% mcmc_hist_by_chain()
   
 }
 

### 2021-04-28 time-series/ma-model/ma-model-greta.R

@@ -12,17 +12,18 @@ library("here")
 ## generate/specify data
 ################################################################################
 
-N <- 200L # sample size
-theta <- c(0.7)
+N <- 50L # sample size
+theta <- c(0.6)
 sigma <- 1
 
 set.seed(1)
 (y <- arima.sim(model = list(ma = theta), sd = sigma, n = N) %>% as.numeric())
 
+# epsilon <- y
 epsilon <- y * 1.01
 
 t <- 1:length(y)
-ggplot(data.frame(t = t, y = y), aes(t,y)) + geom_line()
+# ggplot(data.frame(t = t, y = y), aes(t,y)) + geom_line()
 
 y_1 <- y[1]
 y_2 <- y[2:length(y)]
@@ -50,14 +51,14 @@ distribution(y) <- normal(mu + theta * epsilon, sigma)
 
 greta_model <- model(mu, theta, sigma)
 
-plot(greta_model)
+# plot(greta_model)
 
 ## configure model settings
 ################################################################################
 
-n_chains <- 1
-n_iter <- 5000
-n_warmup <- 2000
+n_chains <- 4
+n_iter <- 1e4L
+n_warmup <- 1e3L
 
 
 
@@ -80,6 +81,26 @@ if (!currently_benchmarking()) {
   ################################################################################
   
   summary(greta_fit)
+  
+  greta_fit_object <- greta_fit %>% as.array()
+  dim(greta_fit_object) <- c(dim(greta_fit_object), 1)
+  dimnames(greta_fit_object) <- list(
+    "iterations" = NULL, 
+    "chains" = 1:n_chains, 
+    "parameters" = c("theta")
+  )
+  
+  
+  greta_fit_object %>% mcmc_areas()
+  greta_fit_object %>% mcmc_intervals()
+  
+  
+  ## assess convergence issues 
+  ###################################################################################
+  
+  greta_fit_object %>% mcmc_acf_bar()
+  greta_fit_object %>% mcmc_trace()
+  greta_fit_object %>% mcmc_hist_by_chain()
   
 }
 

### 2021-05-07 simulation-results/Simulation Results.Rmd

--- 2020-05-10 Simulation Results/Simulation Results.Rmd
@@
 author: "Evan Miyakawa"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-```{r}
+```{r, eval = FALSE}
 exp_gamma_bench_files <- c(
@@
 
-```{r}
+```{r, eval = FALSE}
 # bench_results <- bind_rows(
@@
 
-# saveRDS(bench_results_indiv, file = "bench_results_indiv.RDS")
+saveRDS(bench_results_indiv, file = "bench_results_indiv.RDS")
 # saveRDS(bench_results_indiv, file = "bench_results_indiv2.RDS")
@@
 
-## Strip Charts
+<!-- ## Strip Charts -->
 
-```{r}
-bench_results_indiv %>% 
-  mutate(model = model_name %>% 
-           str_replace_all("_", " ")) %>% 
-  ggplot(aes(method, time)) + 
-  geom_jitter(width = 0.1) + 
-  facet_wrap(vars(model)) + 
-  theme_update(panel.grid.minor = element_blank()) + 
-  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile")) + 
-  ylim(0,100)
+<!-- ```{r} -->
+<!-- bench_results_indiv %>%  -->
+<!--   mutate(model = model_name %>%  -->
+<!--            str_replace_all("_", " ")) %>%  -->
+<!--   ggplot(aes(method, time)) +  -->
+<!--   geom_jitter(width = 0.1) +  -->
+<!--   facet_wrap(vars(model)) +  -->
+<!--   theme_update(panel.grid.minor = element_blank()) +  -->
+<!--   scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile")) +  -->
+<!--   ylim(0,100) -->
 
-bench_results_indiv %>% 
-  mutate(model = model_name %>% 
-           str_replace_all("_", " ")) %>% 
-  # filter(time < 100) %>% 
-  ggplot(aes(method, time, color = method)) + 
-  geom_jitter(width = 0.1) + 
-  facet_wrap(vars(model), scales = "free_y") + 
-  theme_update(panel.grid.minor = element_blank()) + 
-  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile"))
+<!-- bench_results_indiv %>%  -->
+<!--   mutate(model = model_name %>%  -->
+<!--            str_replace_all("_", " ")) %>%  -->
+<!--   # filter(time < 100) %>%  -->
+<!--   ggplot(aes(method, time, color = method)) +  -->
+<!--   geom_jitter(width = 0.1) +  -->
+<!--   facet_wrap(vars(model), scales = "free_y") +  -->
+<!--   theme_update(panel.grid.minor = element_blank()) +  -->
+<!--   scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile")) -->
 
-```
+<!-- ``` -->
 
-- First slide of strip charts
-```{r}
-first_eight <- bench_results_indiv %>% 
-  mutate(model = model_name %>% 
-           str_replace_all("_", " ")) %>% 
-  group_by(model) %>% 
-  summarize(mean_time = mean(time)) %>% 
-  arrange(mean_time) %>% 
-  pull(model) %>% 
-  .[1:8] %>% 
-  sort()
+<!-- - First slide of strip charts -->
+<!-- ```{r} -->
+<!-- first_nine <- bench_results_indiv %>%  -->
+<!--   mutate(model = model_name %>%  -->
+<!--            str_replace_all("_", " ")) %>%  -->
+<!--   group_by(model) %>%  -->
+<!--   summarize(mean_time = mean(time)) %>%  -->
+<!--   arrange(mean_time) %>%  -->
+<!--   pull(model) %>%  -->
+<!--   .[1:9] %>%  -->
+<!--   sort() -->
 
-first_six <- bench_results_indiv %>% 
-  mutate(model = model_name %>% 
-           str_replace_all("_", " ")) %>% 
-  group_by(model) %>% 
-  summarize(mean_time = mean(time)) %>% 
-  arrange(mean_time) %>% 
-  pull(model) %>% 
-  .[1:6] %>% 
-  sort()
+<!-- # first_six <- bench_results_indiv %>%  -->
+<!-- #   mutate(model = model_name %>%  -->
+<!-- #            str_replace_all("_", " ")) %>%  -->
+<!-- #   group_by(model) %>%  -->
+<!-- #   summarize(mean_time = mean(time)) %>%  -->
+<!-- #   arrange(mean_time) %>%  -->
+<!-- #   pull(model) %>%  -->
+<!-- #   .[1:6] %>%  -->
+<!-- #   sort() -->
 
-bench_results_indiv %>% 
-  mutate(model = model_name %>% 
-           str_replace_all("_", " ")) %>% 
-  filter(model %in% first_eight) %>% 
-  filter(time < 100) %>% 
-  ggplot(aes(method, time, color = method)) + 
-  geom_jitter(width = 0.2, size = 1) + 
-  # facet_wrap(vars(model), scales = "free_y") + 
-  facet_wrap(vars(model)) + 
-  theme_update(panel.grid.minor = element_blank()) + 
-  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile"))
+<!-- bench_results_indiv %>%  -->
+<!--   mutate(model = model_name %>%  -->
+<!--            str_replace_all("_", " ")) %>%  -->
+<!--   filter(model %in% first_nine) %>%  -->
+<!--   filter(time < 100) %>%  -->
+<!--   ggplot(aes(method, time, color = method)) +  -->
+<!--   geom_jitter(width = 0.2, size = 1) +  -->
+<!--   # facet_wrap(vars(model), scales = "free_y") +  -->
+<!--   facet_wrap(vars(model)) +  -->
+<!--   theme_update(panel.grid.minor = element_blank()) +  -->
+<!--   scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile")) +  -->
+<!--   labs(title = "Completion Time Strip Charts", subtitle = "Each individual simulation computation completion time shown, facetted by model") +  -->
+<!--   theme(plot.title = element_text(hjust = 0.5, size = 20),  -->
+<!--         plot.subtitle = element_text(hjust = 0.5),  -->
+<!--         legend.position = "none",  -->
+<!--         axis.text = element_text(size=10),  -->
+<!--         axis.title = element_text(size = 13),  -->
+<!--         axis.title.x = element_text(vjust = -1), -->
+<!--         axis.title.y = element_text(vjust = 2))  +  -->
+<!--   xlab("Implementation") +  -->
+<!--   ylab("Completion Time") -->
 
-bench_results_indiv %>% 
-  mutate(model = model_name %>% 
-           str_replace_all("_", " ")) %>% 
-  filter(model %in% first_six) %>% 
-  filter(time < 100) %>% 
-  ggplot(aes(method, time, color = method)) + 
-  geom_jitter(width = 0.2) + 
-  # facet_wrap(vars(model), scales = "free_y") + 
-  facet_wrap(vars(model)) + 
-  theme_update(panel.grid.minor = element_blank()) + 
-  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile"))
+<!-- # bench_results_indiv %>%  -->
+<!-- #   mutate(model = model_name %>%  -->
+<!-- #            str_replace_all("_", " ")) %>%  -->
+<!-- #   filter(model %in% first_six) %>%  -->
+<!-- #   filter(time < 100) %>%  -->
+<!-- #   ggplot(aes(method, time, color = method)) +  -->
+<!-- #   geom_jitter(width = 0.2) +  -->
+<!-- #   # facet_wrap(vars(model), scales = "free_y") +  -->
+<!-- #   facet_wrap(vars(model)) +  -->
+<!-- #   theme_update(panel.grid.minor = element_blank()) +  -->
+<!-- #   scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile")) -->
 
-```
+<!-- ``` -->
+
+<!-- - second slide -->
+
+<!-- ```{r} -->
+<!-- last_seven <- bench_results_indiv %>%  -->
+<!--   mutate(model = model_name %>%  -->
+<!--            str_replace_all("_", " ")) %>%  -->
+<!--   group_by(model) %>%  -->
+<!--   summarize(mean_time = mean(time)) %>%  -->
+<!--   arrange(mean_time) %>%  -->
+<!--   pull(model) %>%  -->
+<!--   .[10:16] %>%  -->
+<!--   sort() -->
+
+<!-- bench_results_indiv %>%  -->
+<!--   mutate(model = model_name %>%  -->
+<!--            str_replace_all("_", " ")) %>%  -->
+<!--   filter(model %in% last_seven) %>%  -->
+<!--   filter(time < 100) %>%  -->
+<!--   ggplot(aes(method, time, color = method)) +  -->
+<!--   geom_jitter(width = 0.2, size = 1) +  -->
+<!--   # facet_wrap(vars(model), scales = "free_y") +  -->
+<!--   facet_wrap(vars(model)) +  -->
+<!--   theme_update(panel.grid.minor = element_blank()) +  -->
+<!--   scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile")) +  -->
+<!--   labs(title = "Completion Time Strip Charts", subtitle = "Each individual simulation computation completion time shown, facetted by model") +  -->
+<!--   theme(plot.title = element_text(hjust = 0.5, size = 20),  -->
+<!--         plot.subtitle = element_text(hjust = 0.5),  -->
+<!--         legend.position = "none",  -->
+<!--         axis.text = element_text(size=10),  -->
+<!--         axis.title = element_text(size = 13),  -->
+<!--         axis.title.x = element_text(vjust = -1), -->
+<!--         axis.title.y = element_text(vjust = 2))  +  -->
+<!--   xlab("Implementation") +  -->
+<!--   ylab("Completion Time") -->
+<!-- ``` -->
 
@@
 
-```{r}
-bugs_fit_condense <- readRDS(here("Simulation Results", "simple-models", 
-                                  "discrete", "binomial", "beta-prior", 
-                                  "bugs_fit_condense_beta_binomial.RDS"))
-jags_fit_condense <- readRDS(here("Simulation Results", "simple-models", 
-                                  "discrete", "binomial", "beta-prior", 
-                                  "jags_fit_condense_beta_binomial.RDS"))
-stan_fit_condense <- readRDS(here("Simulation Results", "simple-models", 
-                                  "discrete", "binomial", "beta-prior", 
-                                  "stan_fit_condense_beta_binomial.RDS"))
-greta_fit_condense <- readRDS(here("Simulation Results", "simple-models", 
-                                  "discrete", "binomial", "beta-prior", 
-                                  "greta_fit_condense_beta_binomial.RDS"))
-nimble_fit_condense <- readRDS(here("Simulation Results", "simple-models", 
-                                  "discrete", "binomial", "beta-prior", 
-                                  "nimble_fit_condense_beta_binomial.RDS"))
-dists <- tibble(
-  x = c(bugs_fit_condense, jags_fit_condense, stan_fit_condense, 
-        greta_fit_condense, nimble_fit_condense),
-  method = c(rep("bugs", length(bugs_fit_condense)),
-             rep("jags", length(jags_fit_condense)),
-             rep("stan", length(stan_fit_condense)),
-             rep("greta", length(greta_fit_condense)),
-             rep("nimble", length(nimble_fit_condense))
-             )
-)
+<!-- ```{r} -->
+<!-- bugs_fit_condense <- readRDS(here("Simulation Results", "simple-models",  -->
+<!--                                   "discrete", "binomial", "beta-prior",  -->
+<!--                                   "bugs_fit_condense_beta_binomial.RDS")) -->
+<!-- jags_fit_condense <- readRDS(here("Simulation Results", "simple-models",  -->
+<!--                                   "discrete", "binomial", "beta-prior",  -->
+<!--                                   "jags_fit_condense_beta_binomial.RDS")) -->
+<!-- stan_fit_condense <- readRDS(here("Simulation Results", "simple-models",  -->
+<!--                                   "discrete", "binomial", "beta-prior",  -->
+<!--                                   "stan_fit_condense_beta_binomial.RDS")) -->
+<!-- greta_fit_condense <- readRDS(here("Simulation Results", "simple-models",  -->
+<!--                                   "discrete", "binomial", "beta-prior",  -->
+<!--                                   "greta_fit_condense_beta_binomial.RDS")) -->
+<!-- nimble_fit_condense <- readRDS(here("Simulation Results", "simple-models",  -->
+<!--                                   "discrete", "binomial", "beta-prior",  -->
+<!--                                   "nimble_fit_condense_beta_binomial.RDS")) -->
+<!-- dists <- tibble( -->
+<!--   x = c(bugs_fit_condense, jags_fit_condense, stan_fit_condense,  -->
+<!--         greta_fit_condense, nimble_fit_condense), -->
+<!--   method = c(rep("bugs", length(bugs_fit_condense)), -->
+<!--              rep("jags", length(jags_fit_condense)), -->
+<!--              rep("stan", length(stan_fit_condense)), -->
+<!--              rep("greta", length(greta_fit_condense)), -->
+<!--              rep("nimble", length(nimble_fit_condense)) -->
+<!--              ) -->
+<!-- ) -->
 
-dists %>% 
-  ggplot(aes(x, color = method)) + 
-  geom_density(adjust = 2) + 
-  stat_function(fun = dbeta, args = list(3,9), color = "black")
+<!-- dists %>%  -->
+<!--   ggplot(aes(x, color = method)) +  -->
+<!--   geom_density(adjust = 2) +  -->
+<!--   stat_function(fun = dbeta, args = list(3,9), color = "black") -->
 
-tibble(x = bugs_fit_condense) %>% 
-  ggplot(aes(x)) + 
-    geom_density(adjust = 2) + 
-    stat_function(fun = dbeta, args = list(3,9), color = "red")
+<!-- tibble(x = bugs_fit_condense) %>%  -->
+<!--   ggplot(aes(x)) +  -->
+<!--     geom_density(adjust = 2) +  -->
+<!--     stat_function(fun = dbeta, args = list(3,9), color = "red") -->
 
-tibble(x = jags_fit_condense) %>% 
-  ggplot(aes(x)) + 
-    geom_density(adjust = 2) + 
-    stat_function(fun = dbeta, args = list(3,9), color = "red")
+<!-- tibble(x = jags_fit_condense) %>%  -->
+<!--   ggplot(aes(x)) +  -->
+<!--     geom_density(adjust = 2) +  -->
+<!--     stat_function(fun = dbeta, args = list(3,9), color = "red") -->
 
-tibble(x = stan_fit_condense) %>% 
-  ggplot(aes(x)) + 
-    geom_density(adjust = 2) + 
-    stat_function(fun = dbeta, args = list(3,9), color = "red")
+<!-- tibble(x = stan_fit_condense) %>%  -->
+<!--   ggplot(aes(x)) +  -->
+<!--     geom_density(adjust = 2) +  -->
+<!--     stat_function(fun = dbeta, args = list(3,9), color = "red") -->
 
-```
+<!-- ``` -->
 
@@
 normal_normal_acc_files <- c(
-  here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-accuracy.rds")
-  # here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-accuracy-nimble.rds")
+  here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-accuracy.rds"),
+  here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-accuracy-nimble.rds")
 )
@@
 
-  stan_mean <- combined_results$stan_dist %>% mean()
+  # stan_mean <- combined_results$stan_dist %>% mean()
+  stan_mean <- 1
 
@@
   combine_time_indiv("normal_gamma_tau")
-# normal_normal_acc_df <- normal_normal_acc_files %>% 
-#   combine_time_indiv("normal_normal")
+normal_normal_acc_df <- normal_normal_acc_files %>%
+  combine_time_indiv("normal_normal")
 beta_bernoulli_acc_df <- beta_bernoulli_acc_files %>% 
@@
   normal_gamma_tau_acc_df,
+  normal_normal_acc_df,
   beta_bernoulli_acc_df,
@@
 )
+# 
+# time_results %>% 
+#   pivot_longer(
+#     # cols = c(jags_dist, bugs_dist, stan_dist, greta_dist, nimble_dist), 
+#     cols = c(JAGS, OpenBUGS, Stan, Nimble, Greta), 
+#     names_to = "method",
+#     values_to = "TVD"
+#   ) %>% 
+#   ggplot(aes(TVD)) + 
+#     geom_histogram(aes(fill = method)) + 
+#     facet_wrap(facets = "method") + 
+#   theme(legend.position="none",
+#         plot.title = element_text(hjust = 0.5),
+#         text = element_text(size=20)) + 
+#   ggtitle("Implementation TVD Accuracy Histograms") + 
+#   xlab("Implementation") + 
+#   ylab("TVD Ratio")
+# 
+# time_results %>% 
+#   pivot_longer(
+#     # cols = c(jags_dist, bugs_dist, stan_dist, greta_dist, nimble_dist), 
+#     cols = c(JAGS, OpenBUGS, Stan, Nimble, Greta), 
+#     names_to = "method",
+#     values_to = "TVD_Ratio"
+#   ) %>% 
+#   ggplot(aes(x = method, y = TVD_Ratio)) + 
+#     geom_boxplot(aes(color = method), size = 0.7) + 
+#   ggtitle("Implementation TVD Accuracy Boxplots") + 
+#   xlab("Implementation") + 
+#   ylab("TVD Ratio") + 
+#   scale_x_discrete(limits = c("OpenBUGS", "JAGS", "Nimble", "Stan", "Greta")) + 
+#   theme(legend.position="none",
+#         plot.title = element_text(hjust = 0.5),
+#         text = element_text(size=20))
 
+```
+
+## Strip charts for time
+
+- First slide of strip charts
+```{r}
+first_nine <- bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  group_by(model) %>% 
+  summarize(mean_time = mean(time)) %>% 
+  arrange(mean_time) %>% 
+  pull(model) %>% 
+  .[1:9] %>% 
+  sort()
+
+# first_six <- bench_results_indiv %>% 
+#   mutate(model = model_name %>% 
+#            str_replace_all("_", " ")) %>% 
+#   group_by(model) %>% 
+#   summarize(mean_time = mean(time)) %>% 
+#   arrange(mean_time) %>% 
+#   pull(model) %>% 
+#   .[1:6] %>% 
+#   sort()
+
+bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  filter(model %in% first_nine) %>% 
+  filter(time < 100) %>% 
+  ggplot(aes(method, time, color = method)) + 
+  geom_jitter(width = 0.2, size = 1) + 
+  # facet_wrap(vars(model), scales = "free_y") + 
+  facet_wrap(vars(model)) + 
+  theme_update(panel.grid.minor = element_blank()) + 
+  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile")) + 
+  labs(title = "Completion Time Strip Charts", subtitle = "Each individual simulation computation completion time shown, facetted by model") + 
+  theme(plot.title = element_text(hjust = 0.5, size = 20), 
+        plot.subtitle = element_text(hjust = 0.5), 
+        legend.position = "none", 
+        axis.text = element_text(size=10), 
+        axis.title = element_text(size = 13), 
+        axis.title.x = element_text(vjust = -1),
+        axis.title.y = element_text(vjust = 2))  + 
+  xlab("Implementation") + 
+  ylab("Completion Time (Seconds)")
+
+# bench_results_indiv %>% 
+#   mutate(model = model_name %>% 
+#            str_replace_all("_", " ")) %>% 
+#   filter(model %in% first_six) %>% 
+#   filter(time < 100) %>% 
+#   ggplot(aes(method, time, color = method)) + 
+#   geom_jitter(width = 0.2) + 
+#   # facet_wrap(vars(model), scales = "free_y") + 
+#   facet_wrap(vars(model)) + 
+#   theme_update(panel.grid.minor = element_blank()) + 
+#   scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile"))
+
+```
+
+- second slide
+
+```{r}
+last_seven <- bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  group_by(model) %>% 
+  summarize(mean_time = mean(time)) %>% 
+  arrange(mean_time) %>% 
+  pull(model) %>% 
+  .[10:16] %>% 
+  sort()
+
+bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  filter(model %in% last_seven) %>% 
+  filter(time < 100) %>% 
+  ggplot(aes(method, time, color = method)) + 
+  geom_jitter(width = 0.2, size = 1) + 
+  # facet_wrap(vars(model), scales = "free_y") + 
+  facet_wrap(vars(model)) + 
+  theme_update(panel.grid.minor = element_blank()) + 
+  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile")) + 
+  labs(title = "Completion Time Strip Charts", subtitle = "Each individual simulation computation completion time shown, facetted by model") + 
+  theme(plot.title = element_text(hjust = 0.5, size = 20), 
+        plot.subtitle = element_text(hjust = 0.5), 
+        legend.position = "none", 
+        axis.text = element_text(size=10), 
+        axis.title = element_text(size = 13), 
+        axis.title.x = element_text(vjust = -1),
+        axis.title.y = element_text(vjust = 2))  + 
+  xlab("Implementation") + 
+  ylab("Completion Time (Seconds)")
+```
+
+- overall boxplot time
+
+```{r}
+bench_results_indiv %>% 
+  mutate(model = model_name %>% 
+           str_replace_all("_", " ")) %>% 
+  filter(time < 100) %>% 
+  ggplot(aes(method, time, color = method)) + 
+  # geom_jitter(width = 0.2, size = 1) + 
+  geom_boxplot() + 
+  # facet_wrap(vars(model), scales = "free_y") + 
+  # facet_wrap(vars(model)) + 
+  theme_update(panel.grid.minor = element_blank()) + 
+  scale_x_discrete(limits = c("JAGS", "OpenBUGS", "Stan",  "Nimble", "Greta", "Stan Compile")) + 
+  labs(title = "Completion Time Overall Boxplots", subtitle = "Completion time aggregated across all models, for each implementation") + 
+  theme(plot.title = element_text(hjust = 0.5, size = 20), 
+        plot.subtitle = element_text(hjust = 0.5), 
+        legend.position = "none", 
+        axis.text = element_text(size=10), 
+        axis.title = element_text(size = 13), 
+        axis.title.x = element_text(vjust = -1),
+        axis.title.y = element_text(vjust = 2))  + 
+  xlab("Implementation") + 
+  ylab("Completion Time (Seconds)")
+```
+
+- first chart accuracy
+```{r}
 time_results %>% 
@@
     names_to = "method",
-    values_to = "TVD_Ratio"
+    values_to = "TVD"
   ) %>% 
-  ggplot(aes(TVD_Ratio)) + 
-    geom_histogram(aes(fill = method)) + 
-    facet_wrap(facets = "method") + 
-  theme(legend.position="none",
-        plot.title = element_text(hjust = 0.5),
-        text = element_text(size=20)) + 
-  ggtitle("Implementation TVD Accuracy Histograms") + 
+  mutate(model = model %>% 
+           str_replace_all("_", " ")) %>% 
+  ggplot(aes(method, TVD, color = method)) + 
+  geom_jitter(width = 0.1, size = 1) + 
+  facet_wrap(vars(model)) +
+  theme_update(panel.grid.minor = element_blank()) + 
+  scale_x_discrete(limits = c("JAGS", "Stan", "Nimble", "OpenBUGS", "Greta")) + 
   xlab("Implementation") + 
-  ylab("TVD Ratio")
+   labs(title = "TVD Strip Charts", subtitle = "Each individual simulation TVD shown, facetted by model") + 
+  theme(plot.title = element_text(hjust = 0.5, size = 20), 
+        plot.subtitle = element_text(hjust = 0.5), 
+        legend.position = "none", 
+        axis.text = element_text(size=12), 
+        axis.title = element_text(size = 15), 
+        axis.title.x = element_text(vjust = -1),
+        axis.title.y = element_text(vjust = 2))
+```
 
+- overall accuracy boxplot
+
+```{r}
 time_results %>% 
@@
     names_to = "method",
-    values_to = "TVD_Ratio"
+    values_to = "TVD"
   ) %>% 
-  ggplot(aes(x = method, y = TVD_Ratio)) + 
-    geom_boxplot(aes(color = method), size = 0.7) + 
-  ggtitle("Implementation TVD Accuracy Boxplots") + 
+  mutate(model = model %>% 
+           str_replace_all("_", " ")) %>% 
+  ggplot(aes(method, TVD, color = method)) + 
+  # geom_jitter(width = 0.1, size = 1) + 
+  geom_boxplot() + 
+  # facet_wrap(vars(model)) +
+  theme_update(panel.grid.minor = element_blank()) + 
+  scale_x_discrete(limits = c("JAGS", "Stan", "Nimble", "OpenBUGS", "Greta")) + 
   xlab("Implementation") + 
-  ylab("TVD Ratio") + 
-  scale_x_discrete(limits = c("OpenBUGS", "JAGS", "Nimble", "Stan", "Greta")) + 
-  theme(legend.position="none",
-        plot.title = element_text(hjust = 0.5),
-        text = element_text(size=20))
+  labs(title = "TVD Overall Boxplots", subtitle = "TVD aggregated across all models, for each implementation") +
+  theme(plot.title = element_text(hjust = 0.5, size = 20), 
+        plot.subtitle = element_text(hjust = 0.5), 
+        legend.position = "none", 
+        axis.text = element_text(size=12), 
+        axis.title = element_text(size = 15), 
+        axis.title.x = element_text(vjust = -1),
+        axis.title.y = element_text(vjust = 2))
+
+# time_results %>%
+#   pivot_longer(
+#     # cols = c(jags_dist, bugs_dist, stan_dist, greta_dist, nimble_dist),
+#     cols = c(JAGS, OpenBUGS, Stan, Nimble, Greta),
+#     names_to = "method",
+#     values_to = "TVD"
+#   ) %>%
+#   mutate(model = model %>%
+#            str_replace_all("_", " ")) %>%
+#   ggplot(aes(method, TVD, color = method)) +
+#   # geom_jitter(width = 0.1, size = 1) +
+#   geom_boxplot() +
+#   # facet_wrap(vars(model)) +
+#   theme_update(panel.grid.minor = element_blank()) +
+#   scale_x_discrete(limits = c("JAGS", "Stan", "Nimble", "OpenBUGS", "Greta")) +
+#   theme(
+#     legend.position = "none"
+#   ) +
+#   xlab("Implementation") +
+#   labs(title = "TVD Overall Boxplots", subtitle = "TVD aggregated across all models, for each implementation") +
+#   theme(plot.title = element_text(hjust = 0.5),
+#         plot.subtitle = element_text(hjust = 0.5))
+#
 
 ```
-
-## Strip charts for accuracy
-
-```{r}
-time_results
-```
